---
title: "TC_genesis_dataset_challenges"
author: "Keith Howen"
date: "2025-08-06"
output: html_document
---

# 💻 set seed for reproducibility
```{r}
set.seed(123)
```



# 💻 Working Directory Setup using `here()`
Use `here()` to *standardize the working directory between R markdown and
R console* as opposed to using `getwd()` (and `setwd()`), at the slight
inconveniences of importing the `here` library, using `here()` to write file
paths, and manually specify paths for files outside the working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports
## Check working directory
```{r}
here()
```

## Import all libraries (incl. custom)
```{r}

# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
library(mvtnorm)
#library(truncnorm)
#library(Matrix)
library(ggplot2)
library(cowplot)
#library(splitstackshape)
#install.packages('reshape2')
library(reshape2)

# for HL test
install.packages('glmtoolbox')
library(glmtoolbox)


```


## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
data.NA.fixed.no.basin = data.NA.fixed[colnames(data.NA.fixed) != "basin"]
data.NA.fixed.no.TC = data.NA.fixed.no.basin[colnames(data.NA.fixed.no.basin) != "TC_genesis"]

X.colnames = colnames(data.NA.fixed.no.TC)

```


# Challenge: scale of variables
```{r}
hist(apply(data.NA.fixed.no.basin, MARGIN=2, FUN=sd),
     breaks=30,
     col='red', freq=FALSE)
```



# Challenge: large n, p
```{r}
str(data.NA.fixed.no.TC)
```


# Challenge: high correlation

Idea: obtain most correlated pairs

```{r}

tc.cor.matrix = cor(data.NA.fixed.no.TC)

# ignore diagonals and upper triangulars
tc.cor.matrix[upper.tri(tc.cor.matrix, diag=TRUE)] = NA

# melt corr matrix into a df
tc.corr.pairs = melt(tc.cor.matrix)

# remove NA values i.e. redundant pairs and diagonals
tc.corr.pairs = tc.corr.pairs[!is.na(tc.corr.pairs$value), ]

tc.corr.pairs


tc.corr.pairs[order(tc.corr.pairs$value, decreasing = TRUE), ]


```


# Univariate signals.

## 1-var logistic model, get all residual deviance

```{r}

# store variable names, drop in deviance (against the null), and coeff est
vars = rep(NaN, length(X.colnames))
deviance.drops = rep(NaN, length(X.colnames))
estimates = rep(NaN, length(X.colnames))

for (i in 1:length(X.colnames)){
  
  # fit a 1-var model
  var = X.colnames[i]
  f = as.formula(paste0("TC_genesis ~", var))
  mod.s = summary(
    glm(f, family="binomial",data=data.NA.fixed.no.basin)
  )
  
  # store var name, deviance drop, coefficient estimate
  vars[i] = var
  deviance.drops[i] = mod.s$null.deviance - mod.s$deviance
  estimates[i] = mod.s$coefficients[2,"Estimate"]
  
}

# order by drop in deviance against null deviance, in decending order.
univar.screening = data.frame(vars=vars, deviance.drops=deviance.drops, estimates=estimates)
univar.screening = univar.screening[order(univar.screening$deviance.drops, decreasing=TRUE),]

univar.screening

```

```{r}

univar.screening

```


```{r}

summary(
    glm(as.formula(paste0("TC_genesis ~", "vertical_wind_shear_between_200_hPa_and_700_hPa", "+", "vertical_wind_shear_between_300_hPa_and_600_hPa")), family="binomial",data=data.NA.fixed.no.basin)
)

cor(data.NA.fixed.no.basin$vertical_wind_shear_between_200_hPa_and_700_hPa,
    data.NA.fixed.no.basin$vertical_wind_shear_between_300_hPa_and_600_hPa)

```



```{r}

plot(data.NA.fixed.no.basin$absolute_vorticity_averaged, data.NA.fixed.no.basin$TC_genesis)

```




# Boxplots
```{r}
boxplot_var_by_TC_genesis <- function(data, col_var, std = TRUE,
                                      cols = c("0" = "#0B7D0B",  # dark green
                                               "1" = "#D62728"), # red
                                      show_trend = TRUE,
                                      median_width = 0.55,
                                      show_legend=TRUE) {
  library(dplyr); library(ggplot2)

  df <- data
  if (std) {
    df[, colnames(df) != "TC_genesis"] <- scale(df[, colnames(df) != "TC_genesis"])
  }
  
  
  # set TC_genesis to be a factor, NOT continuous
  df$TC_genesis <- factor(df$TC_genesis, levels = c(0, 1))

  # group medians
  meds <- df %>%
    group_by(TC_genesis) %>%
    summarise(med = median(.data[[col_var]], na.rm = TRUE), .groups = "drop")

  # for optional trend line between the two medians
  seg <- if (nrow(meds) == 2) data.frame(
    x = meds$TC_genesis, y = meds$med, grp = 1
  ) else NULL

  # col_var is not a literal, so need to use .data[[..]]
  ggplot(df, aes(x = TC_genesis, y = .data[[col_var]])) +
    
    # colored boxes by class
    geom_boxplot(aes(fill = TC_genesis), outlier.alpha = 0.25) +
    
    # optional: show a “trend” line connecting medians
    { if (show_trend && !is.null(seg))
        geom_path(data = seg, aes(x = x, y = y, group = grp),
                  inherit.aes = FALSE, linetype = "solid", linewidth = 0.7,
                  color = "blue",
                  size=0.2,
                  lineend = "round"
                  ) } +
    
    coord_flip() +
    
    scale_fill_manual(values = cols, name = "TC_genesis",
                      labels = c("0 = no genesis", "1 = genesis")) +
    
    scale_color_manual(values = cols, guide = "none") +
    
    labs(x = "TC_genesis", y = col_var) +
    
    theme_minimal(base_size = 13) +
    
    theme(legend.position = ifelse(show_legend, "top", "none"))
}
```


## Geopotential


```{r}
cols = grep('geopotential', colnames(data.NA.fixed.no.basin), value=TRUE)

plist = list()

#plist[[1]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin, cols[1], show_legend=TRUE)

for (i in 1:length(cols)){
  plist[[i]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,
                                                     cols[i],
                                                     show_legend=FALSE)
}

p = plot_grid(plotlist = plist,
          ncol = 3, nrow=ceiling(length(plist)/3),
          rel_widths = c(2, 2, 2),
          rel_heights=c(2,2,2))

p

ggsave(here("output", "tc_challenge_plots", "boxplots_geopotential.png"),
        plot = p, width = 12, height = 10, units = "in", dpi = 300)


```

## Vertical wind shear

```{r}
cols = grep('vertical_wind_shear', colnames(data.NA.fixed.no.basin), value=TRUE)

cols = c(head(cols), tail(cols))

plist = list()

#plist[[1]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin, cols[1], show_legend=TRUE)

for (i in 1:length(cols)){
  plist[[i]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,
                                                     cols[i],
                                                     show_legend=FALSE)
}

p = plot_grid(plotlist = plist,
          ncol = 3, nrow=ceiling(length(plist)/3),
          rel_widths = c(2, 2, 2),
          rel_heights=c(2,2,2))

p

ggsave(here("output", "tc_challenge_plots", "boxplots_vertical_wind_shear.png"),
        plot = p, width = 15, height = 10, units = "in", dpi = 300)


```



## Specific humidity

```{r}
cols = grep('specific_humidity', colnames(data.NA.fixed.no.basin), value=TRUE)

cols = c(head(cols), tail(cols))

plist = list()

#plist[[1]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin, cols[1], show_legend=TRUE)

for (i in 1:length(cols)){
  plist[[i]] = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,
                                                     cols[i],
                                                     show_legend=FALSE)
}

p = plot_grid(plotlist = plist,
          ncol = 3, nrow=ceiling(length(plist)/3),
          rel_widths = c(2, 2, 2),
          rel_heights=c(2,2,2))
p

ggsave(here("output", "tc_challenge_plots", "boxplots_specific_humidity.png"),
        plot = p, width = 12, height = 10, units = "in", dpi = 300)



```

