---
title: "analyze_MH_multiple_chains"
output: html_document
date: "2025-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Load all local and ext libraries 
```{r}
if (!requireNamespace("here", quietly=TRUE)) install.packages("here")
library(here) # to override here(), simply go to File -> New Project
here()
library(parallel)
library(future.apply)
library(progressr)

source(here("lib", "gibbs_sampler_random_search.R"))
source(here("lib", "load_and_preprocess_TC.R"))

#SEED=123
#set.seed(SEED) # NOTE: NO further set.seed() needs to be invoked.
```
```{r}
list.files(here("output", "mh_runs_4_lambda_run_2"))
```


```{r}

names(
  readRDS(here("output", "mh_runs_4_lambda_run_2", "mh_lambda_0.25_chain_3_sparse.rds"))
)

readRDS(here("output", "mh_runs_4_lambda_run_2", "mh_lambda_0.25_chain_1_full.rds"))$summary

```


# Initialize


```{r}

# extract 3 things: lambda value, initid, init desc
chainids = str_match( list.files(here("output", "mh_runs_4_lambda_run_2")),
  "([\\d]+\\.+[\\d]+)_chain_(\\d)_([^\\d]+).rds"
  )

# ignore all mismatches and discard raw file names
chainids = chainids[apply(!is.na(chainids), 1, all), -1]

# store lookup for init val descriptions
init.dicts = unique(chainids[,-1])
init.dicts = as.list(setNames(init.dicts[,2], init.dicts[,1]))
chainids = chainids[,-3]

# initialize data structure to store MH samples across all lambda and init vals
MH_samples = list()
for (i in 1:nrow(chainids)){
  lambda = chainids[i,][1]
  initid = chainids[i,][2]
  
  if (! lambda %in% names(MH_samples)){
    MH_samples[[lambda]] = list()
  } 
  
  MH_samples[[lambda]][[initid]] = list()
}
```


# Read samples

```{r}

for (i in 1:nrow(chainids)){
  
  lambda = chainids[i,][1]
  initid = chainids[i,][2]
  initdesc = init.dicts[[initid]]
  
  fp = sprintf("mh_lambda_%0.2f_chain_%d_%s.rds",
          as.double(lambda),
          as.integer(initid),
          initdesc)
  
  
  MH_samples[[lambda]][[initid]] =
    readRDS(here("output", "mh_runs_4_lambda_run_2", fp))$rs

  
}
```


# Perform chi^2 convergence test

### Function
```{r}
stationarity_chisq <- function(
  S,                      # numeric vector (one long chain) OR list of numeric vectors (q chains)
  k0 = 0,                 # extra burn-in to drop per chain
  q  = NULL,              # number of sub-sequences (if NULL and S is a vector, chosen automatically)
  r  = NULL,              # number of intervals (if NULL, chosen automatically)
  breaks = c("quantile", "equalwidth"),  # how to make r bins
  min_expected = 5,       # target >= 5 expected counts/cell
  simulate_p = FALSE,     # if some cells < 5, use Monte Carlo p-value
  B = 10000               # replicates for Monte Carlo p-value (chisq.test)
) {
  breaks <- match.arg(breaks)

  # --- helper: coerce input to list of q sub-sequences, each length K ---
  to_blocks <- function(S, k0, q) {
    if (is.list(S)) {
      # q chains provided
      chains <- lapply(S, function(x) {
        stopifnot(is.numeric(x))
        if (k0 >= length(x)) stop("k0 >= chain length.")
        x[(k0 + 1):length(x)]
      })
      K <- min(sapply(chains, length))
      if (K < 10) stop("Not enough post-burn draws per chain for splitting.")
      blocks <- lapply(chains, function(x) x[(length(x) - K + 1):length(x)])
      list(blocks = blocks, q = length(blocks), K = K)
    } else {
      # single long chain: split into q blocks after burn-in
      x <- as.numeric(S)
      if (k0 >= length(x)) stop("k0 >= chain length.")
      x <- x[(k0 + 1):length(x)]
      if (is.null(q)) {
        # aim for at least ~100 per block, but ≥ 2 blocks
        q <- max(2L, min(10L, floor(length(x) / 100)))
      }
      K <- floor(length(x) / q)
      if (K < 10) stop("Not enough draws per block; decrease q or k0.")
      # take the last q*K draws so blocks are equal length
      x <- tail(x, q * K)
      blocks <- split(x, rep(seq_len(q), each = K))
      list(blocks = blocks, q = q, K = K)
    }
  }

  blk <- to_blocks(S, k0 = k0, q = q)
  blocks <- blk$blocks; q <- blk$q; K <- blk$K

  # --- choose r if missing to satisfy ~>= min_expected expected per cell ---
  N_total <- q * K
  if (is.null(r)) {
    r <- max(2L, min(10L, floor(N_total / (min_expected * q))))
  }

  # --- define bin edges over pooled S ---
  pooled <- unlist(blocks, use.names = FALSE)
  if (breaks == "quantile") {
    # quantile bins give near-balanced row totals -> helps the 5-per-cell rule
    probs <- seq(0, 1, length.out = r + 1)
    # jitter duplicates in extremes to avoid identical breaks
    qs <- as.numeric(stats::quantile(pooled, probs, type = 7, na.rm = TRUE))
    eps <- .Machine$double.eps^0.25
    qs <- unique(qs)  # guard: if r is too large for unique values, reduce r
    if (length(qs) < r + 1) {
      r <- length(qs) - 1L
      if (r < 2) stop("Too few unique values in S to form >=2 bins.")
    }
    brks <- if (length(qs) == r + 1) qs else stats::quantile(pooled, seq(0, 1, length.out = r + 1))
    brks[1]  <- brks[1]  - eps
    brks[r+1] <- brks[r+1] + eps
  } else {
    brks <- seq(min(pooled), max(pooled), length.out = r + 1)
  }

  # --- build r x q contingency table ---
  binlab <- function(x) cut(x, breaks = brks, include.lowest = TRUE, right = TRUE)
  tab <- sapply(blocks, function(b) table(binlab(b)))
  # ensure matrix with rownames (intervals) and colnames (block index)
  if (is.null(dim(tab))) tab <- matrix(tab, nrow = r)
  rownames(tab) <- levels(binlab(pooled))
  colnames(tab) <- paste0("block_", seq_len(q))

  # --- chi-squared test of independence ---
  # check expected counts
  exp_counts <- suppressWarnings(chisq.test(tab)$expected)
  small <- any(exp_counts < min_expected)

  test <- suppressWarnings(
    chisq.test(tab, simulate.p.value = simulate_p || small, B = B)
  )

  list(
    p_value = unname(test$p.value),
    statistic = unname(test$statistic),
    df = unname(test$parameter),
    small_expected = small,
    method = if (simulate_p || small) "Chi-squared with simulated p-value" else "Pearson Chi-squared",
    r = r, q = q, K = K, k0 = k0,
    breaks = brks,
    table = tab
  )
}

```


```{r}
stationarity_chisq_lambda = function(MH_samples,
                                     extra_burn_in=2000,
                                     breaks="quantile",
                                     r=NULL,
                                     min_expected=5){
  library(stringr)
  

  lambda_str = readline(prompt=
                          paste("Enter lambda value from",
                                str_c(names(MH_samples), collapse = ", "),
                                ": \n"))
  lambda_str = as.character(lambda_str)
  
  # BIC traces across the (four) initial values
  BIC_list = lapply(MH_samples[[as.character(lambda_str)]],
                    function(S){S$bic_trace})
  
  cat("BIC list successfully created.\n")
  
  # perform the chi squared test
  stationarity_chisq(BIC_list,
                     breaks=breaks,
                     k0 = extra_burn_in,
                     r=r,
                     min_expected=min_expected)
}
```


### For lambda=0.1
```{r}
chisq_0.1 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=5000,
                    breaks='quantile',
                    r=10,
                    min_expected=5)
chisq_0.1
```


### For lambda=0.14
```{r}
chisq_0.14 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=5000,
                    breaks='quantile',
                    r=10,
                    min_expected=5)
chisq_0.14
```


### For lambda=0.18
```{r}
chisq_0.18 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=5000,
                    breaks='quantile',
                    r=10,
                    min_expected=5)
chisq_0.18
```


### For lambda=0.20
```{r}
chisq_0.20 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=5000,
                    breaks='quantile',
                    r=10,
                    min_expected=5)
chisq_0.20
```



### For lambda=0.25

```{r}
chisq_0.25 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=5000,
                    breaks='quantile',
                    r=7,
                    min_expected=5)
chisq_0.25
```


### For lambda=0.5

```{r}

chisq_0.5 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=2000,
                    breaks='quantile')

chisq_0.5
```


### For lambda=0.75
```{r}
chisq_0.75 = stationarity_chisq_lambda(MH_samples,
                    extra_burn_in=2000,
                    breaks='quantile')

chisq_0.75

```


# Perform GR diagnostics test
### Function
```{r}
GR_diagnostics_v2 <- function(results,
                           split_chains = TRUE,
                           drop_rule = c("per_chain","global","none"),
                           rhat_cut = 1.05) {
  stopifnot(is.list(results), length(results) >= 2)
  if (!requireNamespace("coda", quietly = TRUE)) install.packages("coda")
  drop_rule <- match.arg(drop_rule)

  # 1) Collect raw indicator matrices per chain
  mats <- lapply(results, function(S) {
    stopifnot(is.matrix(S))
    S
  })

  # 2) Split each chain into two halves (BDA3 recommendation)
  if (split_chains) {
    mats <- unlist(lapply(mats, function(S) {
      n <- nrow(S); h <- floor(n/2)
      list(S[1:h, , drop = FALSE], S[(h+1):n, , drop = FALSE])
    }), recursive = FALSE)
  }

  p <- ncol(mats[[1]])
  vnames <- colnames(mats[[1]])

  # 3) Choose variables to keep (avoid degeneracy)
  keep <- rep(TRUE, p); why_drop <- rep(NA_character_, p)
  if (drop_rule != "none") {
    if (drop_rule == "per_chain") {
      wvar <- sapply(mats, function(S) apply(S, 2, var))
      keep <- apply(wvar > 0, 1, all)
      why_drop[!keep] <- "degenerate_in_some_chain"
    } else if (drop_rule == "global") {
      pooled <- do.call(rbind, mats)
      keep <- apply(pooled, 2, var) > 0
      why_drop[!keep] <- "degenerate_globally"
    }
  }
  if (!any(keep)) stop("No non-degenerate variables to diagnose.")

  # 4) Build coda mcmc.list from subchains
  mlist <- coda::mcmc.list(lapply(mats, function(S) coda::mcmc(S[, keep, drop = FALSE])))

  # 5) Univariate GR and ESS (ESS summed across subchains)
  gr   <- coda::gelman.diag(mlist, multivariate = FALSE)$psrf
  ess  <- sapply(seq_len(sum(keep)), function(j)
           sum(sapply(mats, function(S) coda::effectiveSize(S[, j]))))

  out <- data.frame(
    variable   = vnames[keep],
    Rhat_point = gr[, "Point est."],
    Rhat_upper = gr[, "Upper C.I."],
    ESS_sum    = ess,
    converged  = gr[, "Point est."] < rhat_cut
  )
  rownames(out) <- NULL

  list(
    table   = out[order(out$Rhat_point, decreasing = TRUE), ],
    dropped = data.frame(variable = vnames[!keep], reason = why_drop[!keep]),
    params  = list(split_chains = split_chains, drop_rule = drop_rule, rhat_cut = rhat_cut)
  )
}
```


### GR diag for lambda=0.10

```{r}

XTRA_BURN_IN = 100

GR_diag_0.10 = GR_diagnostics_v2(
  lapply(MH_samples[["0.10"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.10
```

```{r}

saveRDS(GR_diag_0.10,
        here("output", "mh_runs_4_lambda_run_2_diagostic", "GR_diag_0.10.rds"))




```



### GR diag for lambda=0.14

```{r}

XTRA_BURN_IN = 100

GR_diag_0.14 = GR_diagnostics_v2(
  lapply(MH_samples[["0.14"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.14
```


### GR diag for lambda=0.18

```{r}

XTRA_BURN_IN = 100

GR_diag_0.18 = GR_diagnostics_v2(
  lapply(MH_samples[["0.18"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.18
```




### GR diag for lambda=0.20

```{r}

XTRA_BURN_IN = 100

GR_diag_0.20 = GR_diagnostics_v2(
  lapply(MH_samples[["0.20"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.20
```


### GR diag for lambda=0.25


```{r}

XTRA_BURN_IN = 100

GR_diag_0.25 = GR_diagnostics_v2(
  lapply(MH_samples[["0.25"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.25
```


### GR diag for lambda=0.5
```{r}

XTRA_BURN_IN = 800

GR_diag_0.50 = GR_diagnostics_v2(
  lapply(MH_samples[["0.50"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.50
```


### GR diag for lambda=0.75
```{r}

XTRA_BURN_IN = 800

GR_diag_0.75 = GR_diagnostics_v2(
  lapply(MH_samples[["0.75"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_0.75
```


### GR diag for lambda=1
```{r}

XTRA_BURN_IN = 800

GR_diag_1 = GR_diagnostics_v2(
  lapply(MH_samples[["1.00"]], function(S){S$samples[-c(1:XTRA_BURN_IN),]}),
  rhat_cut = 1.1
)

GR_diag_1
```


# Summary

### Acceptance rate

```{r}

for (lam in names(MH_samples)){
  for (initid in names(MH_samples[[lam]])){
    acc.rate=MH_samples[[lam]][[initid]]$acc.rate
    cat(sprintf("For lambda=%s with initid=%s, acc.rate=%0.3f\n",
                lam,initid,acc.rate))
  }
}

#MH_samples[["0.25"]][["1"]]$acc.rate
```

### BIC Trace

```{r}

#want_legend=TRUE

outdir <- here::here("output", "mh_runs_4_lambda_run_2_plots")
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

png(file.path(outdir, "bic_traces_grid.png"),
    width = 12, height = 6, units = "in", res = 300)

# 1. Find global min and max across all bic_traces (scaled by 0.5)
all_bics <- unlist(lapply(MH_samples, function(lam_list) {
  lapply(lam_list, function(chain) chain$bic_trace * 0.5)
}))
bic_min <- min(all_bics)
bic_max <- max(all_bics)

# 2. Adjust layout: extra row for legend
layout(matrix(c(1:(length(MH_samples)), rep(length(MH_samples)+1, 4)),
              ncol = 4, byrow = TRUE))

par(mar = c(4,4,1,0) + 0.08)

# plot each lambda
for (lam in names(MH_samples)) {
  trace1 <- MH_samples[[lam]][["1"]]$bic_trace * 0.5
  
  plot(trace1, type='l', col=1,
       xlab="Iteration", ylab="BIC",
       ylim = c(bic_min, bic_max - 3800))
  
  for (i in 2:length(MH_samples[[lam]])) {
    lines(MH_samples[[lam]][[as.character(i)]]$bic_trace * 0.5,
          col=i)
  }
  
  title(sprintf("BIC trace plots for lambda=%s", lam))
}

# 3. Shared legend at bottom
par(mar=c(0,0,0,0))
plot.new()
legend("bottom",
       legend = unlist(init.dicts),
       col = 1:length(init.dicts),
       lty = 1,
       cex = 1.5,
       horiz = FALSE)

dev.off()




```




### Find BIC Minimizer.
```{r}
# Find the best model across all (lambda, chain) runs
# Assumes each MH_samples[[lambda]][[chain]] has:
#   $samples   : matrix [iters x p] of 0/1 (burn-in already removed)
#   $bic_trace : numeric vector length = iters  (NOTE not yet burn-in adjusted)
#
# data_for_BIC: data.frame(X, y) if we need to compute BICs on the fly
# top_k       : how many unique best models to report
# delta       : ΔBIC threshold for the elite set

find_best_overall <- function(MH_samples, data_for_BIC = NULL,
                              top_k = 5, delta = 2, burn_in=1000) {

  # helper: get BICs for a chain (use stored trace if present)
  get_chain_bics <- function(chain_obj) {
    if (!is.null(chain_obj$bic_trace)) return(chain_obj$bic_trace)
    stopifnot(!is.null(data_for_BIC))
    # lazily compute if not provided
    S <- chain_obj$samples
    apply(S, 1, function(v) compute.BIC.auto(data_for_BIC, v))
  }

  rows <- list()

  for (lam in names(MH_samples)) {
    for (ch in names(MH_samples[[lam]])) {
      obj  <- MH_samples[[lam]][[ch]]
      S    <- obj$samples
      bics <- get_chain_bics(obj)
      
      bics <- bics[ (burn_in+1) :length(bics)]

      # per-iteration records for this chain
      if (is.null(colnames(S))) colnames(S) <- paste0("X", seq_len(ncol(S)))
      keys <- apply(S, 1, paste0, collapse = "")  # hash the model
      rows[[length(rows) + 1]] <- data.frame(
        lambda = lam,
        chain  = ch,
        iter   = seq_along(bics),
        BIC    = bics,
        key    = keys,
        p_active = rowSums(S),
        stringsAsFactors = FALSE
      )
    }
  }

  all_df <- do.call(rbind, rows)

  # Reduce to unique models: keep the *best* BIC seen for each key
  best_by_key <- do.call(rbind, lapply(split(all_df, all_df$key), function(d) {
    d[which.min(d$BIC), ][1, ]  # the row with min BIC for this model
  }))

  # Order by BIC
  best_by_key <- best_by_key[order(best_by_key$BIC), ]

  # The single best model
  best_row <- best_by_key[1, ]
  # Recover variable names for the best model (from any chain where it appeared)
  # Find one sample row that matches its key:
  pick <- NULL
  for (lam in names(MH_samples)) {
    for (ch in names(MH_samples[[lam]])) {
      S <- MH_samples[[lam]][[ch]]$samples
      if (is.null(colnames(S))) colnames(S) <- paste0("X", seq_len(ncol(S)))
      k <- apply(S, 1, paste0, collapse = "")
      w <- which(k == best_row$key)
      if (length(w)) { pick <- list(S = S[w[1], ], names = colnames(S)); break }
    }
    if (!is.null(pick)) break
  }
  best_vars <- pick$names[as.logical(pick$S)]

  # Top-k unique models table
  top_k_tbl <- head(best_by_key, top_k)
  # Attach variable count convenience
  top_k_tbl$k <- top_k_tbl$p_active

  # Elite set (ΔBIC <= delta)
  minB <- best_row$BIC
  elite_tbl <- subset(best_by_key, BIC <= minB + delta)
  elite_tbl$k <- elite_tbl$p_active

  list(
    best = list(
      lambda   = best_row$lambda,
      chain    = best_row$chain,
      iter     = best_row$iter,
      BIC      = as.numeric(best_row$BIC),
      k        = as.integer(best_row$p_active),
      key      = best_row$key,
      variables = best_vars
    ),
    top_k  = top_k_tbl[, c("lambda","chain","iter","BIC","k","key")],
    elite  = elite_tbl[,  c("lambda","chain","iter","BIC","k","key")]
  )
}


# If your objects already contain $bic_trace, you don’t need data_for_BIC:
res <- find_best_overall(MH_samples)

# If not, pass your data.frame(X, y) so BIC can be computed:
# res <- find_best_overall(MH_samples, data_for_BIC = data.frame(X, y))

res$best$BIC
res$best$variables           # names in the overall best model
res$top_k                    # top-5 unique models
res$elite                    # ΔBIC ≤ 2 set (edit `delta` if you like)

```



```{r}

# BIC minimizer from lambda=0.1
min(
  unlist(
  lapply(MH_samples[["0.10"]], function(S){min(S$bic_trace)} )
)
)


```

#### Save BIC minimizer(s)
```{r}
here()

saveRDS(res,
        here("output", "mh_runs_4_lambda_run_2", "BIC_minimizers_lam_0.1_to_1.rds"))
```

#### Find for updated lambda
```{r}


MH_samples_lam_gt_1 =
  MH_samples[
    names(MH_samples)[sapply(names(MH_samples), function(x) as.double(x) > 1)]
  ]


res_lam_from_1.5_to_3 <- find_best_overall(MH_samples_lam_gt_1)

saveRDS(res_lam_from_1.5_to_3,
        here("output", "mh_runs_4_lambda_run_2", "BIC_minimizers_lam_1.50_to_3.rds"))


```


```{r}

res_lam_from_1.5_to_3$best

```


### FIND BIC MINIMIZER FOR ALL.

```{r}


res_lam_all_0.1_to_3 <- find_best_overall(MH_samples)

saveRDS(res_lam_all_0.1_to_3,
        here("output", "mh_runs_4_lambda_run_2", "res_lam_all_0.1_to_3.rds"))


```


```{r}

```


```{r}
res_lam_all_0.1_to_3$best
```


### Distinct prop


```{r}


for (f in list.files(here("output", "mh_runs_4_lambda_run_2"))){
  if (f != "metadata.rds"){
    print(f)
    print(readRDS(here("output", "mh_runs_4_lambda_run_2", f))$summary$distinct_prop)
  }
}



```



# Plots

```{r}
library(ggplot2)
outdir <- here::here("output", "mh_runs_4_lambda_run_2_plots")
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

# build the data once
inc_df <- inclusion_summary(mcmc.samples.0.1.combined, top_k = 65)
```


```{r, width_in = 14, height_in = 10}

# make the plot (your function should return a ggplot object)
p_inc <- plot_inclusion(inc_df) +
  theme(
    text = element_text(size = 11),
    axis.text.y = element_text(size = 8),
    plot.margin = margin(10, 20, 10, 10)  # a bit more room on the right
  )

# dynamic sizing: ~0.25 in per row works well for dense caterpillars
n_rows  <- nrow(inc_df)
height_in <- max(12, 0.25 * n_rows + 2)   # min 12in; pad a little
width_in  <- 12

# 1) Thesis-quality vector PDF (embed fonts, crisp at any zoom)
pdf_file <- file.path(outdir, "inclusion_top65_lambda0.10.pdf")
ggsave(pdf_file, plot = p_inc, device = cairo_pdf,
       width = width_in, height = height_in, units = "in")

# 2) Optional PNG for quick sharing
png_file <- file.path(outdir, "inclusion_top65_lambda0.10.png")
ggsave(png_file, plot = p_inc, dpi = 300,
       width = width_in, height = height_in, units = "in", bg='white')


```

# BIC Minimizer agree with top rel freqs?!

```{r}


top.50.vars.relfreq = names(head(sort(rel.freq.inference(mcmc.samples.0.1.combined)$rel.freqs.non.basin,
     decreasing=TRUE), n=52))



intersect(top.50.vars.relfreq, res$best$variables)

```


```{r}
res$best$variables   
```





# Next Steps/interpretation


With λ = 0.10 chains mix well (R̂ ≈ 1 for all non-degenerate vars), so it’s a solid setting for inference on inclusion frequencies.

The global BIC winner came from λ = 0.75 (chain 3), which is totally fine.
- We will just report the overall best/elite models pooled across all λ, and use λ = 0.10 samples to discuss stability.

Hmm... we can try to further narrow down variable selection to take the intersection of best var set from:
- BIC minimizer (52 variables)
- top 50 with highest rel freq 

Then obtain out of sample predicive performance. 


