---
title: "analyze_results_RCP_n_2000_p_160"
author: "Keith Howen"
date: "2025-07-26"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))

#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")

#library(splitstackshape)

#install.packages("glmnet")
library(glmnet)

```

# Read stuff

## Read synthetic datasets

```{r}
results_n_2000_p_160 = readRDS(
  here("output",
       "simulation_factorial_analysis_n_2000_p_160_outputs",
       "results_n_2000_p_160.rds"))

X.no.corr = results_n_2000_p_160[['0']][['moderate']]$X.syn
y.no.corr = results_n_2000_p_160[['0']][['moderate']]$y.syn

X.high.corr = results_n_2000_p_160[['0.9']][['moderate']]$X.syn
y.high.corr = results_n_2000_p_160[['0.9']][['moderate']]$y.syn
```


## Simulation results from RCP

```{r}


RCP_OUTPUT_FILEPATH = here("output",
                           "simulation_factorial_analysis_n_2000_p_160_outputs",
                           "RCP outputs (more chains)")

RCP_OUTPUT_FILEPATH = file.path(RCP_OUTPUT_FILEPATH)

list.files(RCP_OUTPUT_FILEPATH)

RCP_OUTPUT_METADATA_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                         "metadata.rds")

RCP_OUTPUT_NO_CORR_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                        "gibbs_3chains_results_no_corr.rds")

RCP_OUTPUT_HIGH_CORR_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                          "gibbs_3chains_results_high_corr.rds")

# to access a file, we can do say
# readRDS(RCP_OUTPUT_METADATA_FILEPATH)
```
## metadata for the previous RCP run

```{r}
readRDS(RCP_OUTPUT_METADATA_FILEPATH)
```


## MCMC samples

```{r}
names(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH))

print("No correlation metadata")
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3)

dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time1)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time2)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time3)


print("High correlation metadata")
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3)

dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time1)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time2)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time3)

# read samples

samples.1.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1
samples.2.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2
samples.3.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3


samples.1.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1
samples.2.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2
samples.3.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3


```


# Analyze moderate s/n, no corr

## Relative frequency 

```{r}

samples.1.rel.freq = rel.freq.inference(samples.1.no.corr, CUTOFF=0.5)
sort(samples.1.rel.freq$rel.freqs.non.basin, decreasing=TRUE)
samples.1.rel.freq$selected.variables

```
## BIC Minimizer ("X1" "X2" "X3" "X5")

```{r}
BIC.minimizer.no.corr.1 =  suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.no.corr, y.no.corr)), samples.1.no.corr)
  )
```



#### Compare to BIC of true model.

```{r}
compute.BIC(
  data.frame(cbind(X.no.corr, y.no.corr)),
  as.integer(colnames(X.no.corr) %in% c("X1","X2","X3","X4"))
)
```



## Empirical mode ("X1"  "X2"  "X3"  "X5"  "X94")

```{r}
colnames(X.no.corr)[
  as.logical(find_MAP(samples.1.no.corr))
]
```


# Analyze moderate s/n, high corr

## Relative frequency
```{r}
sort(
  rel.freq.inference(samples.1.high.corr, CUTOFF=0.5)$rel.freqs.non.basin,
  decreasing=TRUE
)


```

## BIC minimizer ("X1"   "X2"   "X3"   "X4"   "X105")

```{r}
suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.high.corr, y.high.corr)), samples.1.high.corr)
  )
```
#### Compare to BIC of true model.

```{r}
compute.BIC(
  data.frame(cbind(X.high.corr, y.high.corr)),
  as.integer(colnames(X.high.corr) %in% c("X1","X2","X3","X4"))
)
```


## Empirical mode ("X1"   "X3"   "X105" "X142" "X146")

```{r}
colnames(X.high.corr)[
  as.logical(find_MAP(samples.1.high.corr))
]
```

## P-chart
```{r}
p_chart_slowest_converging_vars(samples.1.high.corr,
                                metric = "D")
```


```{r}
p_chart_slowest_converging_vars(samples.1.high.corr,
                                metric = "M")
```


```{r}
p_chart_slowest_converging_vars(samples.1.high.corr,
                                metric = "MCSE")
```

# Hauck-Donner effect check

## Check distribution of true probabiilties


```{r}
names(results_n_2000_p_160[['0']][['moderate']])
```
```{r}

true_prob_dist = function(X,
                          beta,
                          title="Distribution of True Probabilities"){
  # Suppose you have
  X_true = as.matrix(X[,1:length(beta)])
  probs <- plogis(as.matrix(X_true) %*% beta)
  hist(probs, breaks = 30,
       main = title, xlab = "p_i = sigmoid(X beta_true)")
  abline(v = c(0.1, 0.9), col = "red", lty = 2) # Danger zone boundaries
}

true_prob_dist(results_n_2000_p_160[['0']][['moderate']]$X.syn,
               results_n_2000_p_160[['0']][['moderate']]$beta)

```
## Check distribution of fitted probabilities in the fitted most efficient model


Most efficient model: model that only contains X1,..,X4, the ground truths.

```{r}
hist(results_n_2000_p_160[['0']][['moderate']]$mod$fitted.values)
```

# Curious: using glmnet() LASSO on both datasets.

## Function

```{r}

lasso.variable.selection = function(X, y,
                                    how='lambda.1se'){
  
  # default is LASSO regression (alpha=1), so we good here
  lasso.fit = glmnet(X, y, family = "binomial")
  
  # note best lambda is not chosen yet -- one way is via cross validation.
  cvfit = cv.glmnet(X, y)
  
  # now, get model coefficients at that value of lambda
  coefs.best.lambda = coef(cvfit, s=how)
  
  # extract non-zero coefficients
  
  vars.non.zero.coefs = names(
      coefs.best.lambda[,how][
        coefs.best.lambda[,how] != 0
        ]
    )
    

  coefs.best.lambda.nonzero =
    coefs.best.lambda[vars.non.zero.coefs, ]
  
  # return everything!
  return(
    list(
      lasso.fit=lasso.fit,
      cvfit=cvfit,
      coefs.best.lambda.nonzero=coefs.best.lambda.nonzero
    )
  )
  
}

```


## Moderate s/n, no corr

```{r}
# default is LASSO regression. 
lasso.no.corr = lasso.variable.selection(X.no.corr, y.no.corr)
lasso.no.corr$coefs.best.lambda.nonzero
```


```{r}
lasso.variable.selection(X.no.corr,
                         y.no.corr,
                         how='lambda.min')$coefs.best.lambda.nonzero
```



## Moderate s/n, high corr

```{r}
# default is LASSO regression. 
lasso.high.corr = lasso.variable.selection(X.high.corr, y.high.corr)
lasso.high.corr$coefs.best.lambda.nonzero
```
```{r}
lasso.variable.selection(X.high.corr,
                         y.high.corr,
                         how='lambda.min')$coefs.best.lambda.nonzero
```



## Generate latex code.
```{r}
lasso_latex_code_string <- function(coefs, caption = "", label = "") {
  # Remove intercept (assumes it's the first element)
  if("(Intercept)" %in% names(coefs)) coefs <- coefs[-1]
  
  df <- as.data.frame(t(coefs))
  
  colnames(df) <- names(coefs)
  
  latex_code <- knitr::kable(
    df, format = "latex", booktabs = TRUE, digits = 4,
    caption = caption, label = label
  )
  
  paste(latex_code, collapse = "\n") # returns a string
}

# Example usage:
latex_string_no_corr <- lasso_latex_code_string(
  lasso.no.corr$coefs.best.lambda.nonzero,
  caption = "LASSO (corr=0)",
  label = "tab:lasso_no_corr"
)

cat(latex_string_no_corr)

latex_string_high_corr <- lasso_latex_code_string(
  lasso.high.corr$coefs.best.lambda.nonzero,
  caption = "LASSO (corr=0.9)",
  label = "tab:lasso_high_corr"
)

cat(latex_string_high_corr)


```


# ⚠ Convergence checks, pool samples if possible

```{r}

dim(
  rbind(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1,
      readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2,
      readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3)
)

```


## Find BIC minimizer for no corr (by pooling)

```{r}
suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.no.corr, y.no.corr)),
                       rbind(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1,
      readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2,
      readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3))
  )
```


## Find BIC minimizer for high corr (by pooling)

```{r}
suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.high.corr, y.high.corr)),
                       rbind(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1,
      readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2,
      readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3))
  )
```

## Empirical mode for no corr (by pooling)

```{r}
colnames(X.no.corr)[
  as.logical(
    find_MAP(
      rbind(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1,
        readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2,
        readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3)
    )
  )
]
```



## Empirical mode for high corr (by pooling)

```{r}
colnames(X.high.corr)[
  as.logical(
    find_MAP(
      rbind(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1,
        readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2,
        readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3)
    )
  )
]
```


# Dimension of models explored overtime
```{r}

summary(
  apply(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1, 1, sum) / 160
)


summary(
  apply(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2, 1, sum) / 160
)

summary(
  apply(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3, 1, sum) / 160
)

```

```{r}

summary(
  apply(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1, 1, sum) / 160
)


summary(
  apply(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2, 1, sum) / 160
)

summary(
  apply(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3, 1, sum) / 160
)

```


# Format to latex renderable code.

```{r}
mcmc.samples.RCP = list()

mcmc.samples.RCP[['0']] = list()
mcmc.samples.RCP[['0']][['moderate']] = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1


mcmc.samples.RCP[['0.9']] = list()
mcmc.samples.RCP[['0.9']][['moderate']] = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1


```


```{r}
summary_table_RCP <- make_summary_from_mcmc_samples(
  mcmc.samples.RCP,
  ground_truth_vars = paste0("X", 1:4),
  results_n_2000_p_160)
```


```{r}
summary_table_RCP
```


```{r}
format_sim_study_tables(summary_table_RCP)
```

# Run a SECOND random search based on BIC minimizer.

## No correlation
```{r}

model.1strun.no.corr = c("X1","X2","X3","X5")
v0.run2.no.corr = rep(1, length(model.1strun.no.corr))

samples.2ndrun.1.no.corr = suppressWarnings(
        Gibbs.random.search(v0.run2.no.corr,
                            data.frame(X.no.corr[,model.1strun.no.corr],
                                       y.no.corr),
                            burnin=300,
                            iter=2300,
                            penalty='BIC')
      )

suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.no.corr[,model.1strun.no.corr], y.no.corr)),
                     samples.2ndrun.1.no.corr)
)


rel.freq.inference(samples.2ndrun.1.no.corr, CUTOFF=0.5)


```




## High correlation
```{r}

model.1strun.high.corr = c("X1","X2","X3","X4","X105")
v0.run2.high.corr = rep(1, length(model.1strun.high.corr))

samples.2ndrun.1.high.corr = suppressWarnings(
        Gibbs.random.search(v0.run2.high.corr,
                            data.frame(X.high.corr[,model.1strun.high.corr],
                                       y.high.corr),
                            burnin=300,
                            iter=2300,
                            penalty='BIC')
      )

suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.high.corr[,model.1strun.high.corr],
                                      y.high.corr)),
                     samples.2ndrun.1.high.corr)
)

rel.freq.inference(samples.2ndrun.1.high.corr, CUTOFF=0.5)


```

