---
title: "TC_sim_factorial_analysis_generate_data_2_large_n_small_p"
author: "Keith Howen"
date: "2025-07-19"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))

#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
#library(splitstackshape)
```

## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
```




# Remarks on modularizing synthetic data generation.

Instead of storing each design matrix as variables `X.c*`, as well as 
the synthetic dataset results as variables `result.c*`, 

# Defining parameters & Inits for the study.

NOTE in favor of the sparsity problem we also choose X1,..,X4 as
the ground truths.

```{r}
p=160
n=2000

# X1 to X2 are CORRELATED ground truths
p_corr_gt = 2

# X3,X4 are UNCORRELATED ground truths
p_uncorr_gt = 2

# The rest are UNCORRELATED REDUNDANTs
p_uncorr_red = p - p_corr_gt - p_uncorr_gt

# Initializing a 'dict' like structure to store results for each
# of the 15 factor combinations
# this is because we cannot directly do multi-level indexing -- 
# we have to initialize each value as a list first, then overwrite it.
rhos <- c('-0.9', '-0.5', '0', '0.5', '0.9')
signal_strengths <- c('weak', 'moderate', 'strong')

results <- list()
for (rho in rhos) {
  
  results[[rho]] <- list()
  
  for (s in signal_strengths) {
    results[[rho]][[s]] <- "empty"  
  }
  
}


Sigmas <- list()
for (rho in rhos) {
  
  Sigmas[[rho]] <- list()
  
  for (s in signal_strengths) {
    Sigmas[[rho]][[s]] <- "empty"  
  }
  
}


X_data <- list()
for (rho in rhos) {
  
  X_data[[rho]] <- list()
  
  for (s in signal_strengths) {
    X_data[[rho]][[s]] <- "empty"  
  }
  
}

results

```

# Tests

## Automate construction of correlation matrix

```{r}
dim(construct_block_cov(p, 0))
```
# Only 2 combinations.

To assess scalability of RS, we just need to assess scenario where s/n
is moderate.

# ✓ Comb 1: rho=0, moderate s/n
Remember, we also need to check the distribution of y. 

```{r}

# generate the design matrix with specified ground truth correlation
Sigmas[['0']][['moderate']] = construct_block_cov(p, 0)
X_data[['0']][['moderate']] = generate_synthetic_vars(Sigmas[['0']][['moderate']],
                                                    SIZE=n)

dim(X_data[['0']][['moderate']])
sum(Sigmas[['0']][['moderate']]) - sum(diag(Sigmas[['0']][['moderate']]))

# we can use the same resuulting beta values from n=2000, p=16
# for moderate s/n and no correlation
# recall it is c(3, 0.35, -3, -0.18)

results[['0']][['moderate']] = check_significance(
                                  c(3, 0.3, -3, -0.3),
                                  X_data[['0']][['moderate']])


```


# ✓ Comb 2: rho=0.9, moderate s/n

```{r}

# generate the design matrix with specified ground truth correlation
Sigmas[['0.9']][['moderate']] = construct_block_cov(p, 0.9)
X_data[['0.9']][['moderate']] = generate_synthetic_vars(Sigmas[['0.9']][['moderate']],
                                                    SIZE=n)

dim(X_data[['0.9']][['moderate']])
#sum(Sigmas[['0.9']][['moderate']]) - sum(diag(Sigmas[['0.9']][['moderate']]))
Sigmas[['0.9']][['moderate']][1:10, 1:10]

# we can use the same resuulting beta values from n=2000, p=16
# for moderate s/n and high correlation
# recall it is c(3, 0.6, -3, -0.29)

results[['0.9']][['moderate']] = check_significance(
                                  c(3, 0.37, -3, -0.25),
                                  X_data[['0.9']][['moderate']])


```


# Save the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(results,
        here("output",
             "simulation_factorial_analysis_outputs",
             "results_n_2000_p_160.rds"))
```





