---
title: "TC_sim_factorial_analysis_RS_2_and_3"
author: "Keith Howen"
date: "2025-07-22"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))

source(here("lib", "probit_gibbs.R"))
source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")

#library(splitstackshape)
```

## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
```


# BASE DIRECTORY TO SAVE TO
```{r}
BASE_DIR = "simulation_factorial_analysis_n_2000_p_160_outputs"
```



# Read the 2 synthetic datasets (n=2000, p=160)


```{r}
results_n_2000_p_160 = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "results_n_2000_p_160.rds"))

names(results_n_2000_p_160[['0']][['moderate']])
names(results_n_2000_p_160[['0.9']][['moderate']])


round(cor(results_n_2000_p_160[['0.9']][['moderate']]$X.syn[,1:5]), 5)

```

# MCMC RS: run both in parallel (moderate s/n and no & high corr)

## Iter 1

```{r}


X.no.corr = results_n_2000_p_160[['0']][['moderate']]$X.syn
y.no.corr = results_n_2000_p_160[['0']][['moderate']]$y.syn

X.high.corr = results_n_2000_p_160[['0']][['moderate']]$X.syn
y.high.corr = results_n_2000_p_160[['0']][['moderate']]$y.syn


if (!require(future)) install.packages("future", dependencies=TRUE)
library(future)

# 1st attempt failed. 
#Error in loadNamespace(j <- i[[1L]], c(lib.loc, .libPaths()), versionCheck = vI[[j]]) : 
#  namespace 'parallelly' 1.43.0 is being loaded, but >= 1.44.0 is required
#Calls: <Anonymous> ... namespaceImportFrom -> asNamespace -> loadNamespace
#Execution halted
#ERROR: lazy loading failed for package 'future'
#* removing 'C:/Users/Keith/AppData/Local/R/win-library/4.3/future'
#Warning in install.packages :
#  installation of package ‘future’ had non-zero exit status
# SOLUTION:
# 1. get version number
# packageVersion("parallelly") 
# 2. make sure to have RTools installed, and click "Tools" -> "Check for
# Package Updates" and scroll to "Parallelly"
# install.packages("parallelly") 

plan(multisession, workers = 2) # or however many cores you want

f1 <- future({
  suppressWarnings(
    Gibbs.random.search(
      rep(1, ncol(X.no.corr)),
      data.frame(cbind(X.no.corr, y.no.corr)),
      burnin = 200,
      iter = 1000,
      penalty = 'BIC'
    )
  )
})

f2 <- future({
  suppressWarnings(
    Gibbs.random.search(
      rep(1, ncol(X.high.corr)),
      data.frame(cbind(X.high.corr, y.high.corr)),
      burnin = 200,
      iter = 1000,
      penalty = 'BIC'
    )
  )
})

gibbs.RS.study2.moderate.no.corr <- value(f1)
gibbs.RS.study2.moderate.high.corr <- value(f2)

saveRDS(gibbs.RS.study2.moderate.no.corr,
        here("output",
             BASE_DIR,
             "gibbs_samples_n_2000_p_160_no_corr.rds"))

saveRDS(gibbs.RS.study2.moderate.high.corr,
        here("output",
             BASE_DIR,
             "gibbs_samples_n_2000_p_160_high_corr.rds"))


```


### Inferences

#### Relative freqs

```{r}
# relative frequencies
rel.freq.inference(gibbs.RS.study2.moderate.no.corr, CUTOFF=0.5)
rel.freq.inference(gibbs.RS.study2.moderate.high.corr, CUTOFF=0.5)


saveRDS(rel.freq.inference(gibbs.RS.study2.moderate.no.corr, CUTOFF=0.5),
        here("output",
             BASE_DIR,
             "gibbs.RS.c1.rel.freqs.rds"))


saveRDS(rel.freq.inference(gibbs.RS.study2.moderate.high.corr, CUTOFF=0.5),
        here("output",
             BASE_DIR,
             "gibbs.RS.c2.rel.freqs.rds"))


```


#### Empirical Mode

```{r}

# find MAP
saveRDS(colnames(X.no.corr)[
  as.logical(find_MAP(gibbs.RS.study2.moderate.no.corr))
],
        here("output",
             BASE_DIR,
             "gibbs.RS.c1.best.Xs.rds"))


saveRDS(colnames(X.high.corr)[
  as.logical(find_MAP(gibbs.RS.study2.moderate.high.corr))
],
        here("output",
             BASE_DIR,
             "gibbs.RS.c2.best.Xs.rds"))


print("Empirical Mode for Moderate s/n, no correlation below")
readRDS(here("output", BASE_DIR, "gibbs.RS.c1.best.Xs.rds"))

print("Empirical Mode for Moderate s/n, high correlation below")
readRDS(here("output", BASE_DIR, "gibbs.RS.c2.best.Xs.rds"))

```


#### BIC Minimizers
```{r}

# find BIC minimizers

saveRDS(suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.no.corr, y.no.corr)), gibbs.RS.study2.moderate.no.corr)
  ),
        here("output",
             BASE_DIR,
             "gibbs.RS.c1.best.Xs.BIC.rds"))


saveRDS(suppressWarnings(
  find_BIC_minimizer(data.frame(cbind(X.high.corr, y.high.corr)), gibbs.RS.study2.moderate.high.corr)
),
        here("output",
             BASE_DIR,
             "gibbs.RS.c2.best.Xs.BIC.rds"))


print("BIC Minimizer for Moderate s/n, no correlation below")
readRDS(here("output", BASE_DIR, "gibbs.RS.c1.best.Xs.BIC.rds"))

print("BIC Minimizer for Moderate s/n, high correlation below")
readRDS(here("output", BASE_DIR, "gibbs.RS.c2.best.Xs.BIC.rds"))

```


#### P-charts

```{r}
# p charts
# Suppose 'samples' is your matrix (K x p) of 0/1 indicator samples
Gibbs.RS.p.chart(gibbs.RS.study2.moderate.no.corr)

# Suppose 'samples' is your matrix (K x p) of 0/1 indicator samples
Gibbs.RS.p.chart(gibbs.RS.study2.moderate.high.corr)

```

#### Heuristic convergence metrics
```{r}
# note these are sorted descendingly by default by (for each variable k)
# D = |p_k^(T/2) - p_k^(T)|
cm_no_corr = convergence_metrics(gibbs.RS.study2.moderate.no.corr)
cm_high_corr = convergence_metrics(gibbs.RS.study2.moderate.high.corr)
```


```{r}
cm_no_corr
```


```{r}
cm_high_corr %>% arrange(desc(M))
```

### P-charts for slowest converging variables

#### Based on D
```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.no.corr,
                                metric = "D")
```

```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.high.corr,
                                metric = "D")
```

#### Based on M
```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.no.corr,
                                metric = "M")
```

```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.high.corr,
                                metric = "M")
```

#### Based on MCSE
```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.no.corr,
                                metric = "MCSE")
```

```{r}
p_chart_slowest_converging_vars(gibbs.RS.study2.moderate.high.corr,
                                metric = "MCSE")
```



### Save Summary Statistics

# Format into latex table

## Functions

```{r}

make_summary_table = function(path_base=here("output", "simulation_factorial_analysis_outputs"),
                              meta=tibble(
                                  combination = 1:15,
                                  correlation = rep(c("0", "+0.5", "+0.9", "-0.5", "-0.9"), 3),
                                  s_n = rep(c("high", "moderate", "weak"), each = 5)
                                )){
  
  # Ground truth variables
  ground_truth_vars <- paste0("X", 1:4)
  
  # Initialize empty list to collect rows
  all_results <- vector("list", length = nrow(meta))
  
  for (i in 1:nrow(meta)) {
    combo_id <- paste0("c", i)
  
    # Read relative frequencies
    rel_freq_obj <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".rel.freqs.rds")))
    rel_freqs <- rel_freq_obj$rel.freqs.non.basin
  
    # Read empirical mode (most common model in MCMC)
    empirical_mode <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".best.Xs.rds")))
  
    # Read BIC-selected model
    bic_model_obj <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".best.Xs.BIC.rds")))
    bic_model <- names(which(bic_model_obj$best.model == 1))
    min_bic <- bic_model_obj$min.BIC
  
    # Logical check: were all GT variables recovered?
    recovered_all_gt_emp <- all(ground_truth_vars %in% empirical_mode)
    recovered_all_gt_bic <- all(ground_truth_vars %in% bic_model)
  
    # Combine into one row (as tibble)
    row <- tibble(
      Combination = i,
      Correlation = meta$correlation[i],
      Signal_to_Noise = meta$s_n[i],
      Empirical_Mode = paste(empirical_mode, collapse = ", "),
      BIC_Model = paste(bic_model, collapse = ", "),
      Min_BIC = min_bic,
      Recovered_GT_Empirical = recovered_all_gt_emp,
      Recovered_GT_BIC = recovered_all_gt_bic
    ) %>%
      bind_cols(as_tibble(t(rel_freqs)))  # Add relative freq columns
  
    all_results[[i]] <- row
  }

  # Final summary table
  return(bind_rows(all_results))
}

# install.packages(c("dplyr","knitr","kableExtra"))
# a little helper to turn one SNR‐subset into a LaTeX table
make_snr_table <- function(df, snr_level) {
  sub <- df %>%
    filter(Signal_to_Noise == snr_level) %>%
    select(
      Correlation,
      `v1 freq` = X1,
      `v2 freq` = X2,
      `v3 freq` = X3,
      `v4 freq` = X4,
      `Empirical mode`    = Empirical_Mode,
      `BIC model`         = BIC_Model,
      `Min. BIC`          = Min_BIC
    )
  
  cap <- paste0("Summary statistics for \\textbf{", snr_level,
                "} signal‐to‐noise across correlation levels")
  
  sub %>%
    kable(
      format       = "latex",
      booktabs     = TRUE,
      caption      = cap,
      label        = paste0("tab:", snr_level),
      align        = c("l", rep("c", ncol(sub)-1)),
      escape       = FALSE
    ) %>%
    kable_styling(latex_options = c("hold_position","scale_down")) %>%
    column_spec(1, bold = TRUE)
}

# generate result tables for latex report

make_sim_study_tables = function(summary_table_filtered){
  # now loop over each SNR level and cat() the tables
  snrs <- unique(summary_table_filtered$Signal_to_Noise)
  for (snr in snrs) {
    cat("\n\n", make_snr_table(summary_table_filtered, snr), "\n\n")
  }
}

```



## Make the summary table
```{r}
summary_table_n_2000_p_160 = make_summary_table(here("output", BASE_DIR),
                   tibble(
                     combination = 1:2,
                     correlation = c("0", "+0.9"),
                     s_n = c("moderate", "moderate")
 ))


SUMMARY_TABLE_COLS_TO_EXCLUDE = c(sapply(5:16, function(i) paste0("X",i)), "Combination")
summary_table_n_2000_p_160_filtered = summary_table[,!colnames(summary_table) %in% SUMMARY_TABLE_COLS_TO_EXCLUDE]

```


```{r}
summary_table_n_2000_p_160
```


## Make the latex-formatted summary table
```{r}
make_sim_study_tables(summary_table_n_2000_p_160)
```


# Brainstorm code to run in RCP

```{r}

```








