---
title: "reanalyze_results_RCP_n_2000_p_160"
author: "Keith Howen"
date: "2025-08-25"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))

#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("cowplot")
#library(cowplot)

#install.packages("mvnTest")

#library(splitstackshape)

#install.packages("glmnet")
library(glmnet)

```

# Read stuff

## Read synthetic datasets

```{r}
results_n_2000_p_160 = readRDS(
  here("output",
       "simulation_factorial_analysis_n_2000_p_160_outputs",
       "results_n_2000_p_160.rds"))

X.no.corr = results_n_2000_p_160[['0']][['moderate']]$X.syn
y.no.corr = results_n_2000_p_160[['0']][['moderate']]$y.syn

X.high.corr = results_n_2000_p_160[['0.9']][['moderate']]$X.syn
y.high.corr = results_n_2000_p_160[['0.9']][['moderate']]$y.syn
```


## Simulation results from RCP

```{r}


RCP_OUTPUT_FILEPATH = here("output",
                           "simulation_factorial_analysis_n_2000_p_160_outputs",
                           "RCP outputs (more chains)")

RCP_OUTPUT_FILEPATH = file.path(RCP_OUTPUT_FILEPATH)

list.files(RCP_OUTPUT_FILEPATH)

RCP_OUTPUT_METADATA_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                         "metadata.rds")

RCP_OUTPUT_NO_CORR_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                        "gibbs_3chains_results_no_corr.rds")

RCP_OUTPUT_HIGH_CORR_FILEPATH = file.path(RCP_OUTPUT_FILEPATH,
                                          "gibbs_3chains_results_high_corr.rds")

# to access a file, we can do say
# readRDS(RCP_OUTPUT_METADATA_FILEPATH)
```
## metadata for the previous RCP run

```{r}
readRDS(RCP_OUTPUT_METADATA_FILEPATH)
```


## MCMC samples

```{r}
names(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH))

print("No correlation metadata")
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3)

dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time1)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time2)
dim(readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$time3)


print("High correlation metadata")
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3)

dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time1)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time2)
dim(readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$time3)

# read samples

samples.1.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init1
samples.2.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init2
samples.3.no.corr = readRDS(RCP_OUTPUT_NO_CORR_FILEPATH)$init3


samples.1.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init1
samples.2.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init2
samples.3.high.corr = readRDS(RCP_OUTPUT_HIGH_CORR_FILEPATH)$init3


```

# (plot) Heatmap of model inclusion


```{r}

# 3) Heatmap of model picks (thin if very long)
plot_grid(
  
  
  plot_model_heatmap(samples.1.no.corr,
                   thin = 10,
                   order = "by_mean",
                   exclude_unimportant_vars=155,
                   put_gts_on_top=FALSE,
                   title = "Model path heatmap (uncorrelated GTs)"),
  
  plot_model_heatmap(samples.1.high.corr,
                   thin = 10,
                   order = "by_mean",
                   exclude_unimportant_vars=152,
                   put_gts_on_top=FALSE,
                   title = "Model path heatmap (highly correlated GTs)"), 
  
  
  nrow=2
  
)





```


# (plot) Trace of number of variables in the sampled models

HORIXONTAL LINE: NUMBER OF VARIABLES IN BIC MINIMIZER


```{r}
# high corr case number of variables in every sampled model
plot_num_vars_trace(samples.1.no.corr,
                    window = 50,
                    title = "Trace: number of variables in every sampled model",
                    dimension_BIC_minim = 4)
```

```{r}
# high corr case number of variables in every sampled model
plot_num_vars_trace(samples.1.high.corr,
                    window = 50,
                    title = "Trace: number of variables in every sampled model")
```

# (plot) CI for rel freqs


```{r}
# 1) Inclusion summary
inc_df <- inclusion_summary(samples.1.high.corr, top_k = 12, gt_idx = gt_idx, ensure_gt = TRUE)
plot_inclusion(inc_df)
```
```{r}
# 1) Inclusion summary
inc_df <- inclusion_summary(samples.1.no.corr, top_k = 12, gt_idx = gt_idx, ensure_gt = TRUE)
plot_inclusion(inc_df)
```



# Spurious ass. for X105, X110, X116

```{r}

inv_spurious_vars = function(X, y, suspect=c("X105", "X116", "X110", "X157")){
  true_vars <- c("X1","X3","X2","X4")  

  df <- data.frame(X, y)
  f_true   <- as.formula(paste("y ~", paste(true_vars, collapse = " + ")))
  f_aug    <- as.formula(paste("y ~", paste(c(true_vars, suspect), collapse = " + ")))
  
  m_true <- glm(f_true, df, family = binomial())
  m_aug  <- glm(f_aug , df, family = binomial())
  
  print( summary(m_true) )
  print( summary(m_aug) )
  
  # p-value for X105 | GTs
  print(
    anova(m_true, m_aug, test = "LRT") 
  )
         
  
  # ΔBIC (>0 favors aug model)
  BIC(m_true) - BIC(m_aug)                  
  
}

inv_spurious_vars(X.high.corr, y.high.corr)

```

## Hmm.. regenerate datasets?

Ok this time X1,X2,X3,X4,X105 have lower BIC than X1,..,X4. 
IS THIS BY CHANCE? Likely so!

```{r}

# Optional helper: EBIC difference when adding exactly 1 variable
# Uses lchoose to avoid overflow: ΔEBIC = ΔBIC + 2γ[logC(p,k_true) - logC(p,k_true+1)]
.ebic_delta_1 <- function(dBIC, p, k_true, gamma = 0.5){
  dBIC + 2*gamma*(lchoose(p, k_true) - lchoose(p, k_true + 1L))
}

check_spurious_LR_BIC <- function(beta_gt= c(3, 0.37, -3, -0.25), X,
                                  suspect = c("X105", "X110", "X116", "157"),
                                  nrep = 100, seed_offset = 1,
                                  alpha = 0.05, compute_ebic = TRUE, gamma = 0.5) {
  # beta_gt: numeric vector of true betas for GT vars (no intercept)
  # X      : design matrix with named columns ("X1","X2",...)
  # suspect: variable name to test (e.g., "X105")
  # nrep   : number of simulated response sets
  # alpha  : LR test level for significance flag
  # compute_ebic/gamma: also compute ΔEBIC with parameter gamma

  stopifnot(is.matrix(X) || is.data.frame(X))
  gt_vars <- paste0("X", seq_along(beta_gt))
  p_total <- ncol(X)
  k_true  <- length(gt_vars)

  res <- vector("list", nrep)

  for (i in 1:nrep) {
    regen <- check_significance(beta_gt, X, SEED = seed_offset + i)
    df <- data.frame(regen$X.syn, y = regen$y.syn)

    f_true <- as.formula(paste("y ~", paste(gt_vars, collapse = " + ")))
    f_aug  <- as.formula(paste("y ~", paste(c(gt_vars, suspect), collapse = " + ")))

    m_true <- glm(f_true, df, family = binomial())
    m_aug  <- glm(f_aug , df, family = binomial())

    a <- anova(m_true, m_aug, test = "LRT")
    ddev <- as.numeric(a$Deviance[2L])
    pval <- as.numeric(a$`Pr(>Chi)`[2L])

    dBIC <- BIC(m_true) - BIC(m_aug)       # >0 favors augmented model
    if (compute_ebic) {
      dEBIC <- .ebic_delta_1(dBIC, p = p_total, k_true = k_true, gamma = gamma)
    } else {
      dEBIC <- NA_real_
    }

    res[[i]] <- data.frame(
      rep = i,
      dDeviance = ddev,
      pval = pval,
      dBIC = dBIC,
      dEBIC = dEBIC,
      lr_sig = (pval < alpha),
      bic_favors = (dBIC > 0),
      ebic_favors = if (compute_ebic) (dEBIC > 0) else NA
    )
  }

  out <- do.call(rbind, res)
  rownames(out) <- NULL
  out
}


res <- check_spurious_LR_BIC(
  X = X.high.corr,
  suspect = "X105",
  nrep = 500,
  compute_ebic = TRUE, gamma = 0.5
)

# Quick summaries
mean(res$lr_sig)          # empirical Type I error (LR @ alpha)
mean(res$bic_favors)      # P(ΔBIC > 0)
mean(res$ebic_favors, na.rm = TRUE)  # P(ΔEBIC > 0), if computed
summary(res$dDeviance); summary(res$dBIC); summary(res$dEBIC)


hist(res$dBIC)
```

```{r}
# Histogram with color-coded bins
h <- hist(res$dBIC, breaks = 30, plot = FALSE)

cols <- ifelse(h$mids < 0, "darkseagreen3", "indianred2")

plot(h, col = cols, border = "white",
     main = "ΔBIC: True model vs True + X105",
     xlab = "ΔBIC (BIC_true - BIC_augmented)")

abline(v = 0, lty = 2, col = "black")
legend("topright", fill = c("darkseagreen3", "indianred2"),
       legend = c("ΔBIC < 0 → True model favored",
                  "ΔBIC > 0 → True model + X105 favored"),
       border = NA)

```



# ✓ Output top 5 BIC minimizers

## no corr case

```{r}
BIC_minim_no_corr = find_BIC_minimizer(data.frame(X.no.corr, y.no.corr),
                   samples.1.no.corr)

BIC_minim_no_corr[["models_by_bic"]][["1"]]
BIC_minim_no_corr[["models_by_bic"]][["2"]]
BIC_minim_no_corr[["models_by_bic"]][["3"]]
BIC_minim_no_corr[["models_by_bic"]][["4"]]
BIC_minim_no_corr[["models_by_bic"]][["5"]]

```

## high corr case

```{r}
BIC_minim_high_corr = find_BIC_minimizer(data.frame(X.high.corr, y.high.corr),
                   samples.1.high.corr)

BIC_minim_high_corr[["models_by_bic"]][["1"]]
BIC_minim_high_corr[["models_by_bic"]][["2"]]
BIC_minim_high_corr[["models_by_bic"]][["3"]]
BIC_minim_high_corr[["models_by_bic"]][["4"]]
BIC_minim_high_corr[["models_by_bic"]][["5"]]

```


```{r}

compute.BIC(data.frame(X.high.corr, y.high.corr),
            as.integer(colnames(X.high.corr) %in% c("X1","X2","X3","X4")))


samples.1.high.corr[,]


```






# Read beta values

```{r}

names(
  results_n_2000_p_160[['0']][['moderate']]
  )


```


```{r}
results_n_2000_p_160[['0']][['moderate']]$beta
results_n_2000_p_160[['0.9']][['moderate']]$beta
```
How do we think we want to present it as a table in the appendix?


# (plot) P-charts

```{r}
p_chart_slowest_converging_vars(samples.1.no.corr, "MCSE")
```


```{r}
p_chart_slowest_converging_vars(samples.1.high.corr, "MCSE")
```







# Case-by-case investigation of spurious variables.

What other variables were in the model where X105 = 1?

```{r}

other_variables_alongside_spurious_vars = function(samples, var="X105"){
  
  # obtain sampled models where var is in the models
  samples.filtered = samples[samples[,"X105"] == 1, ]
  
  # obtain the exact variable sets
  var.sets = apply(samples[samples[,"X105"] == 1, ],
      1,
      function(row) which(row != 0))
  
  return(
    list(
      samples.filtered = samples.filtered,
      var.sets = var.sets
    )
  )
  
}



other_variables_alongside_spurious_vars(samples.1.high.corr)









```



