---
title: "analyze_gibbs_multiple_lambda"
author: "Keith Howen"
date: "2025-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```


```{r}

list.files(
  here("output", "gibbs_runs_multi_lambda", "gibbs_lambda_0.75_chain_1_full.rds")
)

```

```{r}

names(
    readRDS(
    here("output", "gibbs_runs_multi_lambda", "gibbs_lambda_0.75_chain_1_full.rds")
  )
)



plot(
  readRDS(
    here("output", "gibbs_runs_multi_lambda", "gibbs_lambda_0.75_chain_1_full.rds")
  )$rs$bic_trace,
  type='l'
)

```


