---
title: "analyze_MH_lambda_for_writeup"
author: "Keith Howen"
date: "2025-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports
## Check working directory
```{r}
library(here)
here()
```

# Import all libraries (incl. custom)

```{r}

source(here("lib", "load_and_preprocess_TC.R"))
source(here("lib", "gibbs_sampler_random_search.R"))
#library(mvtnorm)
#library(truncnorm)
#library(Matrix)
library(ggplot2)
#library(splitstackshape)
#install.packages('reshape2')
library(reshape2)
library(glmnet)
#install.packages('caret')
library(caret)


# for HL test
#install.packages('glmtoolbox')
library(glmtoolbox)

#install.packages('PRROC')
library(PRROC)

library(cowplot)

library(diptest)

library(purrr)


library(kableExtra)

#install.packages("xtable")
library(xtable)

library(glmnet)


#install.packages("ResourceSelection")
library(ResourceSelection)  # for hoslem.test

library(glmtoolbox)

library(predtools)
library(magrittr)


library(energy)

library(posterior)

library(stringr)

library(precrec)

library(ggtext) 


library(xtable)

```



# Import MH lambda results

```{r}
library(here)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyr)
library(stringr)

# Directory containing all MH chain runs
mh_dir <- here("output", "mh_runs_4_lambda_run_2")

# Get all MH chain result files
mh_files <- list.files(mh_dir, pattern = "^mh_lambda_.*_chain_.*\\.rds$", full.names = TRUE)

# Function to extract the *minimum* BIC per chain
extract_min_bic <- function(path) {
  obj <- readRDS(path)
  s <- obj$summary
  bic_min <- min(obj$rs$bic_trace * 0.5)
  tibble(
    lambda = s$lambda,
    chain  = factor(s$chain),
    init   = s$init,
    min_bic = bic_min
  )
}

# Combine all into one table
bic_summary <- map_dfr(mh_files, extract_min_bic)

# For plotting aesthetics
chain_labels <- c("1: full", "2: empty", "3: sparse", "4: random")
chain_colors <- c("#E66101", "#5E3C99", "#4DAC26", "#FFD700")

bic_summary <- bic_summary %>%
  mutate(chain = factor(chain, levels = 1:4, labels = chain_labels)) %>%
  arrange(lambda, chain)

# Compute minimum BIC across all 4 chains for each lambda
bic_min_by_lambda <- bic_summary %>%
  group_by(lambda) %>%
  summarise(global_min_bic = min(min_bic))

# Merge back for annotation
bic_summary <- left_join(bic_summary, bic_min_by_lambda, by = "lambda")

# Plot stacked bar chart
ggplot(bic_summary, aes(x = lambda, y = min_bic, fill = chain)) +
  geom_bar(stat = "identity", position = "stack", width = 0.6) +
  geom_text(
    data = bic_min_by_lambda,
    aes(x = lambda, y = global_min_bic * 1.02,
        label = sprintf("min BIC = %.1f", global_min_bic)),
    inherit.aes = FALSE, hjust = 0, size = 3.5
  ) +
  scale_fill_manual(values = chain_colors, name = "Chain") +
  scale_x_continuous(breaks = sort(unique(bic_summary$lambda))) +
  coord_flip() +
  labs(
    x = expression(lambda),
    y = "Minimum BIC per chain",
    title = "Minimum BIC across chains for each λ",
    subtitle = "Bars show chain-wise minima; text shows global minimum per λ"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

```
```{r}
xtable(bic_summary %>% arrange(global_min_bic, min_bic) %>% select(-init) %>% mutate(min_bic_diff_to_global = min_bic - global_min_bic))
```



## Acceptance rate, distinct models for all lambdas
```{r}
names(mh_lambda_0.10_chain_1)

mh_lambda_0.10_chain_3$summary
```

```{r}
library(here)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)

# Directory where all MH runs live
mh_dir <- here("output", "mh_runs_4_lambda_run_2")

# Identify all chain result files
mh_files <- list.files(mh_dir, pattern = "^mh_lambda_.*\\.rds$", full.names = TRUE)

# Helper to extract summary metadata from each file
extract_summary <- function(path) {
  obj <- readRDS(path)
  s <- obj$summary
  tibble(
    lambda = s$lambda,
    chain  = s$chain,
    init   = s$init,
    acc    = s$acc,
    distinct_prop = s$distinct_prop,
    mean_active   = s$mean_active
  )
}

# Read and combine all summaries
mh_summary <- map_dfr(mh_files, extract_summary)

# Optional: order chain IDs consistently
mh_summary <- mh_summary %>%
  mutate(chain = factor(chain, levels = 1:4,
                        labels = c("1: full", "2: empty", "3: sparse", "4: random"))) %>%
  arrange(lambda, chain)

# Pivot wider to make 2-way table (lambda × chain)
distinct_table <- mh_summary %>%
  select(lambda, chain, distinct_prop) %>%
  pivot_wider(names_from = chain, values_from = distinct_prop)

# Optionally include mean_active in parentheses
distinct_table_with_dim <- mh_summary %>%
  mutate(entry = sprintf("%.3f (%.0f)", distinct_prop, mean_active)) %>%
  select(lambda, chain, entry) %>%
  pivot_wider(names_from = chain, values_from = entry)

# Show result
distinct_table_with_dim

```

```{r}
xtable(distinct_table_with_dim)
```


## Read best BIC minimizers for ALL lambda

```{r}

best_BIC_minimizer_all_lambda = readRDS(here("output", "mh_runs_4_lambda_run_2", "res_lam_all_0.1_to_3.rds"))

```




```{r}
names(best_BIC_minimizer_all_lambda)
```

```{r}
best_BIC_minimizer_all_lambda$elite
# variable names are in varnames_elites s.t. 
# varnames_elites[[i]] correspond to i-th row of best_BIC_minimizer_all_lambda$elite
```

```{r}
print(best_BIC_minimizer_all_lambda$elite,
      n=40)
```
```{r}
xtable(best_BIC_minimizer_all_lambda$elite %>% select(-key))
```


### Elite sets

```{r}
# around 38 models with BIC within 2 of the best
best_BIC_minimizer_all_lambda$elite
```




#### Simplest model amongst elite sets

We have a model with BIC only <2 higher with only 50 variables.

```{r}

best_simplest_model_bin = (best_BIC_minimizer_all_lambda$elite %>% filter(k==50))$key 
best_simplest_model_bin = str_split(best_simplest_model_bin, "")[[1]]
best_simplest_model_bin = sapply(
  best_simplest_model_bin, function(x)as.logical(as.numeric(x))
)

```


```{r}
best_simplest_model_varnames = colnames(X)[best_simplest_model_bin]
```


```{r}
# variable names that the best and the simplest elite agree with
best_varnames_best_and_simplest = intersect(
  best_BIC_minimizer_all_lambda$best$variables,
  best_simplest_model_varnames
)
```



```{r}
# Vars in the elite simplest model that are not in the best
setdiff(best_simplest_model_varnames, best_varnames_best_and_simplest)
```


```{r}
# Vars in the best model that are not in the simplest
setdiff(best_BIC_minimizer_all_lambda$best$variables, best_varnames_best_and_simplest)
```

#### "Core variables": Intersection of elite sets

```{r}
# 1. map each binary string to variable names
varnames_elites = sapply(
  best_BIC_minimizer_all_lambda$elite$key,
  
  function(bin_key){
    # split "1100" as c("1","1","0","0"), say
    bin_key = str_split(bin_key, "")[[1]]
    # convert "1" to TRUE, "0" to FALSE
    bin_key = sapply(
      bin_key, function(x)as.logical(as.numeric(x))
    )
    # map logical cols to actual colnames
    colnames(X)[bin_key]
  }, 
  
  USE.NAMES=FALSE
)



# 2. obtain the intersection 
# one way: apply intersect() repeteadly
core_vars_among_elites = Reduce(intersect,varnames_elites)

core_vars_among_elites
```
### Summarize elite models to latex
```{r}

summarize_elite_models = function(varnames_elites){
    # --- helpers: parse family + levels ---------------------------------------
  
  # normalize family names into readable labels (optional tweak as you like)
  pretty_family <- function(fam) {
    fam <- gsub("_", " ", fam, fixed = TRUE)
    fam <- gsub("vorticity relative", "relative vorticity", fam, fixed = TRUE)
    fam <- gsub("vertical wind shear between", "vertical wind shear", fam, fixed = TRUE)
    fam <- gsub("equiv pot temperature diff between surface and 500 hPa",
                "equivalent potential temperature diff (surface–500 hPa)", fam, fixed = TRUE)
    fam <- gsub("Coriolis parameter absolute", "Coriolis parameter (absolute)", fam, fixed = TRUE)
    fam
  }
  
  # extract "family" (physical variable class)
  extract_family <- function(x) {
    # patterns:
    # 1) vertical_wind_shear_between_<a>_hPa_and_<b>_hPa
    if (str_detect(x, "^vertical_wind_shear_between_")) return("vertical_wind_shear")
    # 2) <something>_at_<level>_hPa  -> family = part before _at_
    if (str_detect(x, "_at_\\d+_hPa$")) return(str_replace(x, "_at_\\d+_hPa$", ""))
    # 3) everything else: as is (e.g., sea_surface_temperature, Coriolis_parameter, wind_u)
    x
  }
  
  # extract a compact “level” label
  extract_level <- function(x) {
    # between_<a>_hPa_and_<b>_hPa
    m_between <- str_match(x, "between_(\\d+)_hPa_and_(\\d+)_hPa$")
    if (!is.na(m_between[1,1])) {
      a <- as.integer(m_between[1,2]); b <- as.integer(m_between[1,3])
      return(sprintf("%d–%d hPa", min(a,b), max(a,b)))
    }
    # _at_<a>_hPa
    m_at <- str_match(x, "_at_(\\d+)_hPa$")
    if (!is.na(m_at[1,1])) return(sprintf("%s hPa", m_at[1,2]))
    # no explicit level
    return(NA_character_)
  }
  
  # --- reshape all elites into long df --------------------------------------
  
  elite_long <- imap_dfr(
    varnames_elites,
    ~ tibble(
        model_id = .y,
        variable = .x
      )
  ) %>%
    mutate(
      family_raw = vapply(variable, extract_family, character(1)),
      family     = pretty_family(family_raw),
      level      = vapply(variable, extract_level, character(1))
    )
  
  # --- % presence per family across elite models ----------------------------
  
  # flag presence (per model) if ANY level of a family appears
  present_per_model_family <- elite_long %>%
    distinct(model_id, family) %>%
    mutate(present = 1L)
  
  n_models <- length(varnames_elites)
  
  presence_tbl <- present_per_model_family %>%
    count(family, wt = present, name = "n_models_with_family") %>%
    mutate(present_pct = 100 * n_models_with_family / n_models)
  
  # --- representative levels per family -------------------------------------
  
  # collect unique levels seen for each family; compact the label
  levels_tbl <- elite_long %>%
    distinct(family, level) %>%
    filter(!is.na(level)) %>%
    # split into single-level and between-level for nice ordering
    mutate(
      lvl_num_lo = as.integer(str_extract(level, "^\\d+")),
      lvl_num_hi = as.integer(str_extract(level, "(\\d+) hPa$"))
    ) %>%
    group_by(family) %>%
    summarise(
      levels_sorted = {
        # order: single levels asc, then ranges by lower bound
        levs <- na.omit(level)
        single <- levs[grepl("^\\d+ hPa$", levs)]
        rng    <- levs[grepl("^\\d+–\\d+ hPa$", levs)]
        single <- single[order(as.integer(str_extract(single, "^\\d+")))]
        rng <- rng[order(as.integer(str_extract(rng, "^\\d+")))]
        paste(c(single, rng), collapse = ", ")
      },
      .groups = "drop"
    )
  
  # some families truly have no pressure level (e.g., SST, Coriolis)
  no_level_fams <- setdiff(elite_long$family, levels_tbl$family)
  no_level_tbl  <- tibble(family = no_level_fams, levels_sorted = "—")
  
  # --- final compact summary table ------------------------------------------
  
  elite_summary <- presence_tbl %>%
    bind_rows() %>%
    left_join(bind_rows(levels_tbl, no_level_tbl), by = "family") %>%
    # nicer column order/labels
    transmute(
      `Physical variable`                 = family,
      `Representative pressure levels`    = levels_sorted,
      `Present in (% of elite models)`    = round(present_pct, 1)
    ) %>%
    arrange(desc(`Present in (% of elite models)`),
            `Physical variable`)
  
  elite_summary
}

```


```{r}
summary_elite_models = summarize_elite_models(varnames_elites)
```

```{r}
print(
  xtable(
   summary_elite_models
    ),
  type = "latex",
  #add.to.row = rows_to_color,
  floating=TRUE
)
```


### Fit a logreg model on best variables

```{r}

best.model.MH = glm(y ~ .,
      family=binomial,
      data=data.frame(
        scale(X[, best_BIC_minimizer_all_lambda$best$variables]),y)
      )

summary(best.model.MH)
```
#### Plot coeffs/ log odds ratio

```{r, fig.width=20}


best.model.log.ORs = summary(best.model.MH)$coefficients[,'Estimate']


best.model.ORs.df = data.frame(
  vars = names(best.model.log.ORs),
  coeffs = best.model.log.ORs
)

ggplot(data=best.model.ORs.df,
       aes(x=reorder(vars, -coeffs), y=coeffs)) + 
      geom_bar(stat="identity") + 
      coord_flip()


```

#### HL test for best model

```{r}

hltest(best.model.MH, verbose=TRUE)


```


## Read samples from lambda=0.1
```{r}
mh_lambda_0.10_chain_1 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.10_chain_1_full.rds"))

mh_lambda_0.10_chain_2 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.10_chain_2_empty.rds"))

mh_lambda_0.10_chain_3 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.10_chain_3_sparse.rds"))

mh_lambda_0.10_chain_4 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.10_chain_4_random.rds"))

mh_lambda_0.10_samples = rbind(mh_lambda_0.10_chain_1$rs$samples,
                       mh_lambda_0.10_chain_2$rs$samples,
                       mh_lambda_0.10_chain_3$rs$samples,
                       mh_lambda_0.10_chain_4$rs$samples)

```


### Find correlation between indicators

```{r}

corr_indicators = cor(mh_lambda_0.10_chain_1$rs$samples)
diag(corr_indicators) = NA


max(corr_indicators, na.rm = TRUE)
min(corr_indicators, na.rm = TRUE)


```

```{r}

corr_indicators = cor(mh_lambda_0.10_chain_2$rs$samples)
diag(corr_indicators) = NA


max(corr_indicators, na.rm = TRUE)
min(corr_indicators, na.rm = TRUE)


```

```{r}

corr_indicators = cor(mh_lambda_0.10_chain_3$rs$samples)
diag(corr_indicators) = NA


max(corr_indicators, na.rm = TRUE)
min(corr_indicators, na.rm = TRUE)


```

### Convergence

#### GR
```{r}
list.files(here("output", "mh_runs_4_lambda_run_2_diagostic"))
```
```{r}
GR_diag_0.10 = readRDS(here("output",
                            "mh_runs_4_lambda_run_2_diagostic",
                            "GR_diag_0.10.rds"))
```

```{r}
GR_diag_0.10
```
#### P-chart

```{r}
p_chart_slowest_converging_vars(mh_lambda_0.10_chain_1$rs$samples,
                                metric='MCSE')
```

#### Chisq (auto)


```{r}
stationarity_chisq <- function(
  S,                      # numeric vector (one long chain) OR list of numeric vectors (q chains)
  k0 = 0,                 # extra burn-in to drop per chain
  q  = NULL,              # number of sub-sequences (if NULL and S is a vector, chosen automatically)
  r  = NULL,              # number of intervals (if NULL, chosen automatically)
  breaks = c("quantile", "equalwidth"),  # how to make r bins
  min_expected = 5,       # target >= 5 expected counts/cell
  simulate_p = FALSE,     # if some cells < 5, use Monte Carlo p-value
  B = 10000               # replicates for Monte Carlo p-value (chisq.test)
) {
  breaks <- match.arg(breaks)

  # --- helper: coerce input to list of q sub-sequences, each length K ---
  to_blocks <- function(S, k0, q) {
    if (is.list(S)) {
      # q chains provided
      chains <- lapply(S, function(x) {
        stopifnot(is.numeric(x))
        if (k0 >= length(x)) stop("k0 >= chain length.")
        x[(k0 + 1):length(x)]
      })
      K <- min(sapply(chains, length))
      if (K < 10) stop("Not enough post-burn draws per chain for splitting.")
      blocks <- lapply(chains, function(x) x[(length(x) - K + 1):length(x)])
      list(blocks = blocks, q = length(blocks), K = K)
    } else {
      # single long chain: split into q blocks after burn-in
      x <- as.numeric(S)
      if (k0 >= length(x)) stop("k0 >= chain length.")
      x <- x[(k0 + 1):length(x)]
      if (is.null(q)) {
        # aim for at least ~100 per block, but ≥ 2 blocks
        q <- max(2L, min(10L, floor(length(x) / 100)))
      }
      K <- floor(length(x) / q)
      if (K < 10) stop("Not enough draws per block; decrease q or k0.")
      # take the last q*K draws so blocks are equal length
      x <- tail(x, q * K)
      blocks <- split(x, rep(seq_len(q), each = K))
      list(blocks = blocks, q = q, K = K)
    }
  }

  blk <- to_blocks(S, k0 = k0, q = q)
  blocks <- blk$blocks; q <- blk$q; K <- blk$K

  # --- choose r if missing to satisfy ~>= min_expected expected per cell ---
  N_total <- q * K
  if (is.null(r)) {
    r <- max(2L, min(10L, floor(N_total / (min_expected * q))))
  }

  # --- define bin edges over pooled S ---
  pooled <- unlist(blocks, use.names = FALSE)
  if (breaks == "quantile") {
    # quantile bins give near-balanced row totals -> helps the 5-per-cell rule
    probs <- seq(0, 1, length.out = r + 1)
    # jitter duplicates in extremes to avoid identical breaks
    qs <- as.numeric(stats::quantile(pooled, probs, type = 7, na.rm = TRUE))
    eps <- .Machine$double.eps^0.25
    qs <- unique(qs)  # guard: if r is too large for unique values, reduce r
    if (length(qs) < r + 1) {
      r <- length(qs) - 1L
      if (r < 2) stop("Too few unique values in S to form >=2 bins.")
    }
    brks <- if (length(qs) == r + 1) qs else stats::quantile(pooled, seq(0, 1, length.out = r + 1))
    brks[1]  <- brks[1]  - eps
    brks[r+1] <- brks[r+1] + eps
  } else {
    brks <- seq(min(pooled), max(pooled), length.out = r + 1)
  }

  # --- build r x q contingency table ---
  binlab <- function(x) cut(x, breaks = brks, include.lowest = TRUE, right = TRUE)
  tab <- sapply(blocks, function(b) table(binlab(b)))
  # ensure matrix with rownames (intervals) and colnames (block index)
  if (is.null(dim(tab))) tab <- matrix(tab, nrow = r)
  rownames(tab) <- levels(binlab(pooled))
  colnames(tab) <- paste0("block_", seq_len(q))

  # --- chi-squared test of independence ---
  # check expected counts
  exp_counts <- suppressWarnings(chisq.test(tab)$expected)
  small <- any(exp_counts < min_expected)

  test <- suppressWarnings(
    chisq.test(tab, simulate.p.value = simulate_p || small, B = B)
  )

  list(
    p_value = unname(test$p.value),
    statistic = unname(test$statistic),
    df = unname(test$parameter),
    small_expected = small,
    method = if (simulate_p || small) "Chi-squared with simulated p-value" else "Pearson Chi-squared",
    r = r, q = q, K = K, k0 = k0,
    breaks = brks,
    table = tab
  )
}
```


```{r}
start_time = Sys.time()

#S_list <- lapply(trimmed, function(S) {
#  apply(S, 1, function(v) compute.BIC(data.frame(X, y), v))
#})



S_list_MH = list()
i=1
S_list_MH[[i]] = mh_lambda_0.10_chain_1$rs$bic_trace; i = i + 1

S_list_MH[[i]] = mh_lambda_0.10_chain_2$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.10_chain_3$rs$bic_trace; i = i + 1

S_list_MH[[i]] = mh_lambda_0.10_chain_4$rs$bic_trace; i = i + 1

# CONSIDER SAVING S_List!!!

# no need burn-in -- already calibrated
out <- stationarity_chisq(S_list_MH,
                          k0 = 20,
                          breaks='equalwidth',
                          r=2,
                          simulate_p = FALSE)

end_time = Sys.time()
cat("Overall time taken (minutes): ",
      as.numeric(end_time - start_time, units = "mins"),
      "\n")
out

```

#### Chisq (manual)


```{r}
# 1. get all BIC values
BIC_MH_EXTRA_BURNIN = 2
S_list_MH = list()
i=1
S_list_MH[[i]] = mh_lambda_0.10_chain_1$rs$bic_trace[-c(1:BIC_MH_EXTRA_BURNIN)]; i = i + 1
S_list_MH[[i]] = mh_lambda_0.10_chain_2$rs$bic_trace[-c(1:BIC_MH_EXTRA_BURNIN)]; i = i + 1
S_list_MH[[i]] = mh_lambda_0.10_chain_3$rs$bic_trace[-c(1:BIC_MH_EXTRA_BURNIN)]; i = i + 1
S_list_MH[[i]] = mh_lambda_0.10_chain_4$rs$bic_trace[-c(1:BIC_MH_EXTRA_BURNIN)]; i = i + 1

# 2. create q bins from the range of the BICs in all chains
S_list_MH_combined = unlist(S_list_MH)

want_equal_width = FALSE

if (want_equal_width){
  S_list_MH_combined_min = min(S_list_MH_combined)
  S_list_MH_combined_max = max(S_list_MH_combined)
  range = S_list_MH_combined_max - S_list_MH_combined_min
  breaks = seq(S_list_MH_combined_min, S_list_MH_combined_max, length.out=3)
} else{
  breaks = quantile(S_list_MH_combined, probs=seq(0,1,by=0.25), na.rm=TRUE)
}



# 3. create frequency table
BIC_MH_table = matrix(NA, nrow=length(breaks)-1, ncol=length(S_list_MH))

for (i in 1:length(S_list_MH)){
  BIC_list = S_list_MH[[i]]
  BIC_binned = cut(BIC_list, breaks=breaks, include.lowest=TRUE, right=TRUE)
  
  BIC_MH_table[,i] = as.vector(table(BIC_binned))
}


BIC_MH_table


chisq.test(BIC_MH_table)


```

#### Raftery diag

```{r}
library(coda)

?raftery.diag


raftery.diag(
  as.mcmc(S_list_MH[[1]])
)
```




## Read samples from lambda=0.75

```{r}
mh_lambda_0.75_chain_3 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.75_chain_3_sparse.rds"))

mh_lambda_0.75_chain_1 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.75_chain_1_full.rds"))

mh_lambda_0.75_chain_2 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.75_chain_2_empty.rds"))

mh_lambda_0.75_chain_4 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.75_chain_4_random.rds"))

```

```{r}
mh_lambda_0.75_chains = list(
  mh_lambda_0.75_chain_1,
  mh_lambda_0.75_chain_2,
  mh_lambda_0.75_chain_3,
  mh_lambda_0.75_chain_4
)
```

### Raftery diag

```{r}
library(coda)

?raftery.diag


raftery.diag(as.mcmc(unlist(readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      "mh_lambda_0.10_chain_2_empty.rds"))$rs$bic_trace)))
```

### Find correlation between indicators
```{r}

corr_indicators = cor(mh_lambda_0.75_chain_3$rs$samples)
diag(corr_indicators) = NA


melt(corr_indicators)[
  order(melt(corr_indicators)$value),
]


```


```{r}

corr_indicators = cor(mh_lambda_0.75_chain_1$rs$samples)
diag(corr_indicators) = NA


melt(corr_indicators)[
  order(melt(corr_indicators)$value),
]


```

```{r}
# samples is your post–burn-in 0/1 matrix (iters x p)
pair_stats <- function(samples, v1, v2) {
  x <- samples[, v1]; y <- samples[, v2]
  tab <- table(x, y)  # 2x2 co-inclusion table
  p1  <- mean(x); p2 <- mean(y)
  p11 <- mean(x & y)
  phi <- cor(x, y)               # same as Pearson on binaries
  jac <- if (p11 == 0 && (p1+p2-p11)==0) NA_real_ else p11 / (p1 + p2 - p11) # Jaccard
  list(p1 = p1, p2 = p2, p11 = p11, phi = phi, jaccard = jac, table = tab)
}

# find top K most negative pairs
top_anticor_pairs <- function(samples, K = 25) {
  C <- cor(samples); diag(C) <- NA
  o <- order(C, na.last = NA)  # increasing → most negative first
  idx <- arrayInd(o[seq_len(K)], dim(C))
  data.frame(
    var1 = colnames(samples)[idx[,1]],
    var2 = colnames(samples)[idx[,2]],
    phi  = C[ o[seq_len(K)] ],
    stringsAsFactors = FALSE
  )
}

# Example:
S <- mh_lambda_0.75_chain_3$rs$samples
neg <- top_anticor_pairs(S, K = 20)
neg
pair_stats(S, "divergence_at_1000_hPa", "vertical_velocity_at_925_hPa")

```

### GR
```{r}
GR_diagnostics_v2 <- function(results,
                           split_chains = TRUE,
                           drop_rule = c("per_chain","global","none"),
                           rhat_cut = 1.05) {
  stopifnot(is.list(results), length(results) >= 2)
  if (!requireNamespace("coda", quietly = TRUE)) install.packages("coda")
  drop_rule <- match.arg(drop_rule)

  # 1) Collect raw indicator matrices per chain
  mats <- lapply(results, function(S) {
    stopifnot(is.matrix(S))
    S
  })

  # 2) Split each chain into two halves (BDA3 recommendation)
  if (split_chains) {
    mats <- unlist(lapply(mats, function(S) {
      n <- nrow(S); h <- floor(n/2)
      list(S[1:h, , drop = FALSE], S[(h+1):n, , drop = FALSE])
    }), recursive = FALSE)
  }

  p <- ncol(mats[[1]])
  vnames <- colnames(mats[[1]])

  # 3) Choose variables to keep (avoid degeneracy)
  keep <- rep(TRUE, p); why_drop <- rep(NA_character_, p)
  if (drop_rule != "none") {
    if (drop_rule == "per_chain") {
      wvar <- sapply(mats, function(S) apply(S, 2, var))
      keep <- apply(wvar > 0, 1, all)
      why_drop[!keep] <- "degenerate_in_some_chain"
    } else if (drop_rule == "global") {
      pooled <- do.call(rbind, mats)
      keep <- apply(pooled, 2, var) > 0
      why_drop[!keep] <- "degenerate_globally"
    }
  }
  if (!any(keep)) stop("No non-degenerate variables to diagnose.")

  # 4) Build coda mcmc.list from subchains
  mlist <- coda::mcmc.list(lapply(mats, function(S) coda::mcmc(S[, keep, drop = FALSE])))

  # 5) Univariate GR and ESS (ESS summed across subchains)
  gr   <- coda::gelman.diag(mlist, multivariate = FALSE)$psrf
  ess  <- sapply(seq_len(sum(keep)), function(j)
           sum(sapply(mats, function(S) coda::effectiveSize(S[, j]))))

  out <- data.frame(
    variable   = vnames[keep],
    Rhat_point = gr[, "Point est."],
    Rhat_upper = gr[, "Upper C.I."],
    ESS_sum    = ess,
    converged  = gr[, "Point est."] < rhat_cut
  )
  rownames(out) <- NULL

  list(
    table   = out[order(out$Rhat_point, decreasing = TRUE), ],
    dropped = data.frame(variable = vnames[!keep], reason = why_drop[!keep]),
    params  = list(split_chains = split_chains, drop_rule = drop_rule, rhat_cut = rhat_cut)
  )
}
```

```{r}


GR_lam_0.75 = GR_diagnostics_v2(list(mh_lambda_0.75_chain_1$rs$samples,
                                     mh_lambda_0.75_chain_2$rs$samples,
                                     mh_lambda_0.75_chain_3$rs$samples,
                                     mh_lambda_0.75_chain_4$rs$samples),
                                                          rhat_cut = 1.1)
```

```{r}
GR_lam_0.75
```


### Chisq 
```{r}
S_list_MH = list()
i=1

S_list_MH[[i]] = mh_lambda_0.75_chain_1$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_2$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_3$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_4$rs$bic_trace; i = i + 1

par(mfrow=c(2,2))
for (k in 1:length(S_list_MH)){
  hist(S_list_MH[[k]], breaks=65, col = k+1)
}


# CONSIDER SAVING S_List!!!

# no need burn-in -- already calibrated
out <- stationarity_chisq(S_list_MH,
                          k0 = 5000,
                          breaks='equalwidth',
                          r=6,
                          simulate_p = FALSE)

out
```

```{r}
BIC_all <- unlist(S_list_MH)
best    <- min(BIC_all)
elite_ind_blocks <- lapply(S_list_MH, function(b) 1*(b - best <= 2))
tab_elite <- table(stack(elite_ind_blocks)$values, stack(elite_ind_blocks)$ind)
chisq.test(tab_elite, simulate.p.value = FALSE)
```



# Rel-freqs
## Lambda-wise rel freqs

### Function
```{r}
plot_rel_freqs_pooled_lam = function(lam,
                                     want_asy_ci=TRUE,
                                     alpha=0.1,
                                     epsilon=0.005,
                                     want_errorbar=TRUE){


  mh_lambda_chain_1 = readRDS(here("output",
                                        "mh_runs_4_lambda_run_2",
                                        sprintf("mh_lambda_%.2f_chain_1_full.rds", lam)
                                        )
                                   )
  
  mh_lambda_chain_2 = readRDS(here("output",
                                        "mh_runs_4_lambda_run_2",
                                        sprintf("mh_lambda_%.2f_chain_2_empty.rds", lam)
                                        )
                                   )
  
  mh_lambda_chain_3 = readRDS(here("output",
                                      "mh_runs_4_lambda_run_2",
                                      sprintf("mh_lambda_%.2f_chain_3_sparse.rds", lam)
                                      )
                                   )
  
  mh_lambda_chain_4 = readRDS(here("output",
                                        "mh_runs_4_lambda_run_2",
                                        sprintf("mh_lambda_%.2f_chain_4_random.rds", lam)
                                        )
                                   )
  
  if (want_asy_ci){
     inclusion_summary_asy_CI = inclusion_summary_asy_CI(
                    rbind(mh_lambda_chain_1$rs$samples,
                          mh_lambda_chain_2$rs$samples,
                          mh_lambda_chain_3$rs$samples,
                          mh_lambda_chain_4$rs$samples),
                  top_k=149,
                  alpha=alpha,
                  epsilon=epsilon)
  
      p_gt_0.5 = plot_inclusion(
        inclusion_summary_asy_CI %>% filter(p_hat >= 0.5),
        title="Inclusion probabilities\n(w/ Asymptotic conservative CI)",
        errorbar=want_errorbar
      )
      
      p_le_0.5 = plot_inclusion(
        inclusion_summary_asy_CI %>% filter(p_hat < 0.5),
        title="Inclusion probabilities\n(w/ Asymptotic conservative CI)",
        errorbar=want_errorbar
      )
      
      return(
        list(p_gt_0.5 = p_gt_0.5, p_le_0.5 = p_le_0.5)
      )

  }
  
  incl_summary = inclusion_summary(
                    rbind(mh_lambda_chain_1$rs$samples,
                          mh_lambda_chain_2$rs$samples,
                          mh_lambda_chain_3$rs$samples,
                          mh_lambda_chain_4$rs$samples),
                  top_k=149)
  
  p_gt_0.5 = plot_inclusion(incl_summary %>% filter(p_hat >= 0.5),
        errorbar=want_errorbar)
  
  p_le_0.5 = plot_inclusion(incl_summary %>% filter(p_hat < 0.5),
        errorbar=want_errorbar)
  
  return(
    list(p_gt_0.5 = p_gt_0.5, p_le_0.5 = p_le_0.5)
  )
  

}
```



### lam=0.5

```{r}
plot_rel_freqs_pooled_lam(0.5, epsilon=1e-30)
```



### lam=0.1
```{r}
plot_inclusion(
  inclusion_summary(mh_lambda_0.10_chain_1$rs$samples,
                  top_k=149)
)


```


#### "Cluster" Relative Frequencies

```{r}

dim(mh_lambda_0.10_samples)

mh_lambda_0.10_rel_freqs = rel.freq.inference(mh_lambda_0.10_samples)$rel.freqs.non.basin
mh_lambda_0.10_rel_freqs = sort(mh_lambda_0.10_rel_freqs, decreasing=TRUE)

plot(mh_lambda_0.10_rel_freqs, type='o')


#par(mfrow=c(2,1))
plot(mh_lambda_0.10_rel_freqs[0:50], type='o')
```

```{r}
plot(mh_lambda_0.10_rel_freqs[51:100], type='o')
```

```{r}
plot(mh_lambda_0.10_rel_freqs[101:149], type='o')
```
```{r}
find_elbows <- function(f, K = 1, x = NULL) {
  stopifnot(is.numeric(f), length(f) >= 10, K >= 0, K == as.integer(K))
  n <- length(f)
  if (is.null(x)) x <- seq_len(n)

  # cumulative sums
  c1  <- c(0, cumsum(1 + 0 * x))        # count
  cx  <- c(0, cumsum(x))
  cxx <- c(0, cumsum(x * x))
  cy  <- c(0, cumsum(f))
  cxy <- c(0, cumsum(x * f))
  cyy <- c(0, cumsum(f * f))

  sse_seg <- function(i, j) {
    m   <- j - i + 1L
    if (m == 1L) return(0)              # single-point segment: zero residual
    sx  <-  cx[j+1] -  cx[i]
    sxx <- cxx[j+1] - cxx[i]
    sy  <-   cy[j+1] -  cy[i]
    sxy <-  cxy[j+1] - cxy[i]
    syy <-   cyy[j+1] -  cyy[i]

    XtX <- matrix(c(m, sx, sx, sxx), 2, 2)
    detXtX <- XtX[1,1] * XtX[2,2] - XtX[1,2] * XtX[2,1]
    if (detXtX <= 1e-12) return(0)      # numerical guard (shouldn’t happen with m>=2)
    beta <- solve(XtX, c(sy, sxy))
    rss  <- syy - sum(beta * c(sy, sxy))
    max(rss, 0)
  }

  Kseg <- K + 1L
  dp   <- matrix(Inf, nrow = Kseg, ncol = n)
  prev <- matrix(NA_integer_, nrow = Kseg, ncol = n)

  for (j in 1:n) dp[1, j] <- sse_seg(1L, j)

  for (k in 2:Kseg) {
    for (j in k:n) {
      best <- Inf; arg <- NA_integer_
      for (t in (k-1):(j-1)) {
        val <- dp[k-1, t] + sse_seg(t+1L, j)
        if (val < best) { best <- val; arg <- t }
      }
      dp[k, j]   <- best
      prev[k, j] <- arg
    }
  }

  breaks <- integer(K)
  j <- n
  for (k in Kseg:2) { t <- prev[k, j]; breaks[k-1] <- t; j <- t }
  breaks <- sort(breaks)

  k_params <- 2L * Kseg
  bic <- n * log(dp[Kseg, n] / n) + k_params * log(n)

  list(
    breaks   = breaks,
    clusters = cut(seq_len(n), c(0, breaks, n), labels = FALSE),
    bic      = bic
  )
}


# ----- tiny plotting helper ---------------------------------------------------
plot_elbows <- function(f, elbows, xlab = "Index (ranked)", ylab = "Rel. freq") {
  n <- length(f); x <- seq_len(n)
  plot(x, f, type = "o", pch = 1, xlab = xlab, ylab = ylab, col=elbows$clusters)
  abline(v = elbows$breaks + 0.5, lty = 2, col = "gray50")
}



```

```{r}

# f should be your relative frequencies, sorted decreasing
# e.g., f <- sort(mh_lambda_0.10_rel_freqs, decreasing = TRUE)
f <- sort(mh_lambda_0.10_rel_freqs, decreasing = TRUE)
el1 <- find_elbows(f, K = 1)   # one elbow → two clusters
el2 <- find_elbows(f, K = 2)   # two elbows → three clusters

el1$breaks   # index of the elbow
el2$breaks   # two indices (e.g., c(4, 21))

plot_elbows(f, el2)

```

```{r, fig.width=12, fig.height=7}

el = find_elbows(f, K = 6)
plot_elbows(f, el)

for (brk in el$breaks){
  plot(brk + -2:2,
       mh_lambda_0.10_rel_freqs[brk + -2:2],
       type='o')
  abline(v=brk)
  #axis(side=1, at=(brk + -2:2), las=2)
}



```

```{r}
incl.summary = inclusion_summary(mh_lambda_0.10_samples,
                  top_k=149)
```

```{r}
incl.summary = cbind(incl.summary, el$clusters)
incl.summary = incl.summary %>% rename("clusters" = "el$clusters" )
```

#### Relative frequencies clustered

```{r, fig.width=12, fig.height=10}
plot_inclusion(incl.summary %>% filter(p_hat > 0.5), elbow_clusters=TRUE)
```


```{r, fig.width=7, fig.height=5}}
# Fix the cluster level order once
lvl <- sort(unique(incl.summary$clusters))
pal <- mk_cluster_palette(lvl, base = "Set2")  # or "Set2", "Paired", etc.

#options(repr.plot.width = 12, repr.plot.height =13)

p1 = plot_inclusion_v2(incl.summary %>% dplyr::filter(p_hat > 0.5),
               color_by = "clusters",
               cluster_levels = lvl,
               cluster_palette = pal,
               title = "Inclusion probabilities > 0.5 (w/ ESS 95% CI)") + theme(aspect.ratio = 1.2) 


p2 = plot_inclusion_v2(incl.summary %>% dplyr::filter(p_hat <= 0.5),
               color_by = "clusters",
               cluster_levels = lvl,
               cluster_palette = pal,
               title = "Inclusion probabilities <= 0.5 (w/ ESS 95% CI)") + theme(aspect.ratio = 1)   


ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_clustered_ge_0.5.png"),
       p1, width = 10, height = 9, units = "in", bg='white',
       scale=1.3 # Scale all plot elements (text, points, spacing) with the canvas
       )

ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_clustered_le_0.5.png"),
       p2, width = 10, height = 9, units = "in", bg='white',
       scale=1.5 # Scale all plot elements (text, points, spacing) with the canvas
       )



```


```{r}
p3 = plot_inclusion_v2(incl.summary %>% dplyr::filter(p_hat <= 0.5 & clusters == 6),
               color_by = "clusters",
               cluster_levels = lvl,
               cluster_palette = pal,
               title = "Inclusion probabilities <= 0.5 (w/ ESS 95% CI)") + theme(aspect.ratio = 1)   

ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_clustered_le_0.5_clust6.png"),
       p3, width = 10, height = 9, units = "in", bg='white',
       scale=1.3 # Scale all plot elements (text, points, spacing) with the canvas
       )

```



```{r}
p4 = plot_inclusion_v2(incl.summary %>% dplyr::filter(p_hat <= 0.5 & clusters == 7),
               color_by = "clusters",
               cluster_levels = lvl,
               cluster_palette = pal,
               title = "Inclusion probabilities <= 0.5 (w/ ESS 95% CI)") + theme(aspect.ratio = 1)   

ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_clustered_le_0.5_clust7.png"),
       p4, width = 10, height = 9, units = "in", bg='white',
       scale=1.7 # Scale all plot elements (text, points, spacing) with the canvas
       )

```

### lam=0.75

```{r}
inclusion_summary_lam_0.75 = inclusion_summary(
    rbind(mh_lambda_0.75_chain_1$rs$samples,
          mh_lambda_0.75_chain_2$rs$samples,
          mh_lambda_0.75_chain_3$rs$samples,
          mh_lambda_0.75_chain_4$rs$samples),
                  top_k=149)
```



```{r}
plot_rel_freqs_pooled_lam(0.75, want_asy_ci = FALSE)
```

```{r}
plot_rel_freqs_pooled_lam(0.75, want_asy_ci = FALSE, want_errorbar = FALSE)
```


#### Core variables

```{r}

plot_inclusion(
  inclusion_summary_lam_0.75 %>%
    filter(variable %in% core_vars_among_elites & p_hat >= 0.5)
)


plot_inclusion(
  inclusion_summary_lam_0.75 %>%
    filter(variable %in% core_vars_among_elites & p_hat < 0.5)
)


xtable(
  inclusion_summary_lam_0.75 %>%
    filter(variable %in% core_vars_among_elites) %>%
    select(-is_gt)
)

```

#### For latex
```{r}
xtable(
  inclusion_summary_lam_0.75 %>%
    select(variable, p_hat, ess, se, lower, upper, -is_gt)
)
```


```{r}
print(
  (inclusion_summary_lam_0.75 %>%
         filter(variable %in% core_vars_among_elites) %>%
         filter(grepl("cloud", variable))
       ),
      
      n=50)
```




## Rel freqs for core elites

```{r}
incl.summary.v2 <- incl.summary
incl.summary.v2$is_core_var = incl.summary.v2$variable %in% core_vars_among_elites
```



```{r}

plot_incl_core_elite_gt_0.5 = plot_inclusion_custom_fill(incl.summary.v2 %>% filter(is_core_var == TRUE & p_hat >= 0.5),
                           #col_fill="is_core_var",
                            title = "Inclusion probabilities for core elite sets with p >= 0.5 (w/ ESS 95% CI)") 


plot_incl_core_elite_le_0.5 = plot_inclusion_custom_fill(incl.summary.v2 %>% filter(is_core_var == TRUE & p_hat < 0.5),
                           #col_fill="is_core_var",
                              title = "Inclusion probabilities for core elite sets with p < 0.5 (w/ ESS 95% CI)") 


#ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_core_elite_gt_0.5.png"),
#       plot_incl_core_elite_gt_0.5,
#       width = 10, height = 9, units = "in", bg='white',
#       scale=1.7 # Scale all plot elements (text, points, spacing) with the canvas
#       )


#ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "inclusion_core_elite_le_0.5.png"),
#       plot_incl_core_elite_le_0.5,
#       width = 10, height = 9, units = "in", bg='white',
#       scale=1.7 # Scale all plot elements (text, points, spacing) with the canvas
#       )



```

#### Vars with 0.5 rel freq NOT in core?

```{r}
incl.summary.v2 %>% filter(is_core_var == FALSE & p_hat >= 0.5)
```

```{r}

intersect(
  (incl.summary.v2 %>% filter(is_core_var == FALSE & p_hat >= 0.5))$variable,
  varnames_elites[[1]]
)

```


### LaTeX code for thesis
```{r}


rows_to_color <- list(pos = list(which(incl.summary.v2$is_core_var == TRUE)),
                      command = "\\rowcolor{white}" 
                      )



print(
  xtable(
    incl.summary.v2 %>%
      filter(is_core_var == TRUE) %>%
      select(variable, p_hat, ess)
    ),
  type = "latex",
  #add.to.row = rows_to_color,
  floating=TRUE
)

```




## BIC for core vars

```{r}


compute.BIC.auto(data.frame(X,y),
                 colnames(X) %in% (incl.summary.v2 %>% filter(is_core_var == TRUE))$variable)

```

```{r}
compute.BIC.auto(data.frame(X,y),
                 colnames(X) %in% (incl.summary.v2 %>% filter(is_core_var == TRUE & p_hat > 0.5))$variable)
```
## Importance sampling (0.1 to 0.75)
```{r}
# ---- Importance-sampled inclusion probs: lambda_s -> lambda_t ----------------
# samples_list: list of 0/1 matrices, each (n_draws x p), from lambda_s
# bic_list    : list of numeric vectors, each length n_draws, BIC per draw
# lambda_s    : source lambda (e.g., 0.10)
# lambda_t    : target lambda (e.g., 0.75)
# burn        : drop this many initial draws in each chain
# thin        : keep every 'thin'-th draw (optional)
importance_relfreq <- function(samples_list, bic_list,
                               lambda_s = 0.10, lambda_t = 0.75,
                               burn = 0L, thin = 1L) {
  stopifnot(length(samples_list) == length(bic_list))

  V <- list()
  bic_all <- list()

  for (k in seq_along(samples_list)) {
    S <- samples_list[[k]]
    B <- bic_list[[k]]

    # --- ensure lengths match ---
    n <- min(nrow(S), length(B))
    S <- S[seq_len(n), , drop = FALSE]
    B <- B[seq_len(n)]

    # --- burn & thin ---
    if (burn > 0) {
      S <- S[(burn + 1):nrow(S), , drop = FALSE]
      B <- B[(burn + 1):length(B)]
    }
    if (thin > 1) {
      keep <- seq(1, nrow(S), by = thin)
      S <- S[keep, , drop = FALSE]
      B <- B[keep]
    }

    V[[k]] <- S
    bic_all[[k]] <- B
  }

  V <- do.call(rbind, V)
  bic <- unlist(bic_all)
  stopifnot(nrow(V) == length(bic))

  # --- compute weights safely ---
  delta <- (lambda_t - lambda_s)
  lw <- -delta * bic
  lw <- lw - max(lw)   # stabilize
  w <- exp(lw)

  wsum <- sum(w)
  ess <- (wsum^2) / sum(w^2)

  # --- weighted inclusion probs ---
  num <- as.numeric(crossprod(w, V))   # safe & efficient
  phat <- num / wsum

  se <- sqrt(pmax(phat * (1 - phat), 0) / ess)
  ci_lo <- pmax(0, phat - 1.96 * se)
  ci_hi <- pmin(1, phat + 1.96 * se)

  data.frame(
    var = colnames(V),
    p_hat = phat,
    se = se,
    ci_lo = ci_lo,
    ci_hi = ci_hi,
    ESS = ess,
    row.names = NULL
  )
}


# ---------- Example with your objects -----------------------------------------
# Each is 40000 x 149 of {0,1} indicators at λ = 0.10:
S10 <- list(
  mh_lambda_0.10_chain_1$rs$samples,
  mh_lambda_0.10_chain_2$rs$samples,
  mh_lambda_0.10_chain_3$rs$samples,
  mh_lambda_0.10_chain_4$rs$samples
)

# Matching BIC vectors per draw (replace with your stored BIC traces):
B10 <- list(
  mh_lambda_0.10_chain_1$rs$bic_trace,  # or whatever field stores BIC per iteration
  mh_lambda_0.10_chain_2$rs$bic_trace,
  mh_lambda_0.10_chain_3$rs$bic_trace,
  mh_lambda_0.10_chain_4$rs$bic_trace
)

is_tab_075 <- importance_relfreq(S10, B10, lambda_s = 0.10, lambda_t = 0.75,
                                 burn = 2000, thin = 1)

# Sort by importance:
is_tab_075 <- is_tab_075[order(-is_tab_075$p_hat), ]
head(is_tab_075)

```


```{r}
#plot_inclusion(is_tab_075)

ggplot(is_tab_075 %>% filter(p_hat < 1), aes(x = reorder(var, p_hat), y = p_hat)) + 
  geom_col(width = 0.7, alpha = 0.2, color = NA) +
  coord_flip()
```



# Methodologies: IN SUMMARY

*Methodology: method-wise*
- MH on granular lambda values from 0.05 to 3, each 4 x 40k chains with overdispersed inits (no need to tune lambda, since lower lambda)


*Methdology for variable importance: rel freq inference*
1. For finding BIC minimizers, find the minimizer across ALL chains.
2. For relative frequency inference, pick largest lambda such that chain has converged based on GR diagnostics (happens to be at lambda=0.1) [justify asymptotic property of rel freqs for variables in true model -- it is lambda agnostic]
3. Cluster the relative frequencies based on observable kinks (done)



*Methdology to for DGP reconstruction: best model*


1. We have ~38 models s.t. BIC(model) - BIC(best model) <= 2. Call them 'elite' models on-par to the best (invoke large n & $\Delta BIC(v_1,v_2)$ as a large sample approximation of $BF(v_1,v_2)$).

2. Differences in variable sets from the best model (52 vars) and the simplest model (50 vars) below:

```{r}

best_simplest_model_bin = (best_BIC_minimizer_all_lambda$elite %>% filter(k==50))$key 
best_simplest_model_bin = str_split(best_simplest_model_bin, "")[[1]]
best_simplest_model_bin = sapply(
  best_simplest_model_bin, function(x)as.logical(as.numeric(x))
)

best_simplest_model_varnames = colnames(X)[best_simplest_model_bin]

# variable names that the best and the simplest elite agree with
cat("\nvariable names that the best and the simplest elite agree with\n")
best_varnames_best_and_simplest = intersect(
  best_BIC_minimizer_all_lambda$best$variables,
  best_simplest_model_varnames
)
best_varnames_best_and_simplest

# Vars in the elite simplest model that are not in the best
cat("\nVars in the elite simplest model that are not in the best\n")
setdiff(best_simplest_model_varnames, best_varnames_best_and_simplest)


# Vars in the best model that are not in the simplest
cat("\nVars in the best model that are not in the simplest\n")
setdiff(best_BIC_minimizer_all_lambda$best$variables, best_varnames_best_and_simplest)

```
We do NOT know which variables are redundant or useless in both models. Which brings us to 3:

3. (Under consideration) consider both models, and validate them in some way (e.g. using AUC-ROC and AUC-PR with 5 fold cross validation -- obtain out of sample predicted probas in each test fold, use them to make the AUC-ROC and AUC-PR)

4. From 3, take whatever model with the highest AUC-ROC (or AUC-PR ?? ) and treat it as the best model for reconstructing the DGP.

5. Finally, perform a weighted ```glm()``` to the chosen model (hope intercept is adjusted), then perform HL test. Once fit is adequate, present the coeffs in the weighted glm as our final model.



*Alternative methodology for variable importance: rel freq inference ON CORE ELITE SETS.*

1. Among all 38 elite models, find the largest variable set that consistently 
appears in all 38 elite models (i.e. ```Reduce(intersect, 38 elite models```). Call this
the core elite sets.

2. Plot their variable importance of core elite sets.

3. Then treat all variables with rel freq >= 0.5 as the one that are 'important' (not necessarily influential in the DGP).

4. Do you still want to perform the knee elbow clustering for the rel freqs??


*Defenses for why we consider our results robust* 
1. Procedure already run on standardized dataset
2. Outliers (leverage via mahalanobis distance) already considered. Variable sets with or without them are still the same 

# Validate models



## CV log loss

### best model VS best simplest

```{r}
# choose between 50- and 52-var elite (and any others with ΔBIC ≤ 2)
cv_logloss <- function(X, y, vars, K=5, seed=123){
  set.seed(seed)
  folds <- sample(rep(1:K, length.out=length(y)))
  ll <- 0; n <- 0
  
  ll_across_folds = c()
  n_across_folds = c()
  
  for(k in 1:K){
    tr <- folds!=k; te <- !tr
    fit <- glm(y[tr] ~ ., binomial, data=data.frame(scale(X[tr, vars, drop=FALSE]), y=y[tr]))
    p   <- predict(fit, newdata=data.frame(scale(X[te, vars, drop=FALSE])), type="response")
    # clip probs
    p <- pmin(pmax(p, 1e-12), 1-1e-12)
    
    ll_fold = - sum(y[te]*log(p) + (1-y[te])*log(1-p))
    
    ll_across_folds = c(ll_across_folds,
                        ll_fold)
    n_across_folds = c(n_across_folds, sum(te))
    
    ll <- ll + ll_fold
    n  <- n + sum(te)
  }
  
  return(
    list(
      ll_across_folds = ll_across_folds,
      n_across_folds = n_across_folds,
      avg_ll = ll/n
    )
  )
}

# then compute PR-AUC secondarily (e.g., PRROC::pr.curve)
cv_logloss_best_model = cv_logloss(X, y, best_BIC_minimizer_all_lambda$best$variables)
cv_logloss_best_simplest_model = cv_logloss(X, y, best_simplest_model_varnames)

avg_cv_logloss_best_model = cv_logloss_best_model$avg_ll
avg_cv_logloss_best_simplest_model = cv_logloss_best_simplest_model$avg_ll

```


```{r}
cat(sprintf("log loss for best model: %.4f\nlog loss for best simplest: %.4f\n",
            avg_cv_logloss_best_model,
            avg_cv_logloss_best_simplest_model))


```

### CV log loss across all elite models

```{r}

avg_cv_logloss_elite_models = lapply(1:length(varnames_elites),
                        function(i){
                          cv_ll = cv_logloss(X, y, varnames_elites[[i]])
                          cat("Elite ", i, " done!\n")
                          return(cv_ll)
                      }
              )

```
```{r}
saveRDS(avg_cv_logloss_elite_models,
        here("output",
             "validation_outputs",
             "avg_cv_logloss_elite_models.rds"))



avg_cv_logloss_elite_models_df = data.frame(
  avg_ll = unlist(lapply(avg_cv_logloss_elite_models, function(x) x$avg_ll)),
  ll_across_folds = unlist(lapply(avg_cv_logloss_elite_models, function(x) x$ll_across_folds)),
  n_across_folds = unlist(lapply(avg_cv_logloss_elite_models, function(x) x$n_across_folds)),
)


#saveRDS(avg_cv_logloss_elite_models_df,
#        here("output",
#             "validation_outputs",
#             "avg_cv_logloss_elite_models_df.rds"))

```


```{r}
which.min(avg_cv_logloss_elite_models_df$avg_ll)
```



## CV AUC-PR


### Function: compute AUC-PR

```{r}
# install.packages("precrec")  # if needed
library(precrec)
library(ggplot2)

# y: 0/1 vector (1 = positive), p: predicted probabilities for class 1
pr_stats <- function(y, p, title = "Precision–Recall") {
  mm <- evalmod(scores = p, labels = y)       # computes ROC + PR by default
  au <- auc(mm)                               # data frame of AUCs
  aupr <- subset(au, curvetypes == "PRC")$aucs

  prev <- mean(y == 1)                        # baseline precision
  g <- autoplot(mm, "PRC") +
    labs(title = sprintf("%s (AUC-PR = %.5f)", title, aupr),
         x = "Recall", y = "Precision") +
    geom_hline(yintercept = prev, linetype = 2, alpha = .6) +
    annotate("text", x = .05, y = prev + .02,
             label = sprintf("Baseline = prevalence = %.5f", prev), hjust = 0)

  list(aupr = as.numeric(aupr), plot = g)
}

```

### Function: compute avg CV AUC-PR

```{r}

# compute out of sample AUC-PRs across folds using a 
# pre specified variable set 'vars'
cv_aucpr <- function(X, y, vars, K=5, seed=123, verbose=FALSE){
  
  # 0. fix the fold splitting
  set.seed(seed)
  folds <- sample(rep(1:K, length.out=length(y)))
  
  aucpr_across_folds = c()
  
  for(k in 1:K){
    
    # 1. create training fold and held out fold
    tr <- folds!=k; te <- !tr
    
    if (verbose) cat(sprintf("Proportion of 1s in y[tr] = %.2f\n", sum(y[tr])/length(y[tr]) ))
    
    # 2. fit on training
    fit <- glm(y[tr] ~ ., binomial, data=data.frame(scale(X[tr, vars, drop=FALSE]), y=y[tr]))
    
    if (verbose) print(summary(fit))
    
    # 3. predict on held out 
    
        tryCatch(
        {
            p   <- predict(fit,
                           newdata=data.frame(scale(X[te, vars, drop=FALSE])),
                           type="response")
            # clip probs
            p <- pmin(pmax(p, 1e-12), 1-1e-12)
            
            
            # 4. obtain out of sample AUC-PR
            out <- pr_stats(y[te], p, title="")
            
            # 5. record aucpr
            cat(sprintf("Fold %d done, AUC-PR=%.3f\n", k, out$aupr))
            aucpr_across_folds = c(aucpr_across_folds, out$aupr)
            
            
        },
        error = function(cond) {
          cat(sprintf("Fold %d problematic, see msg below: \n", k))
          message(conditionMessage(cond))
          aucpr_across_folds = c(aucpr_across_folds, NA)
        },
        
        finally = invisible() 

    )
    
    

  }
  
  return(
    list(
      aucpr_across_folds = aucpr_across_folds,
      avg_aucpr = mean(aucpr_across_folds, na.rm=TRUE)
    )
  )
}

# compute AUC PR on training dataset using a var set
compute_auc_pr = function(var_name){
  fit <- glm(y ~ .,
           family = binomial,
           data = data.frame(X[,var_name],y))
  p <- predict(fit, type = "response")
  out <- pr_stats(y, p, title="")
  out$aupr
}



# plot PR curve on training dataset using a var set
plot_PR = function(var_name, title="AUC-PR: Best Model"){
  fit <- glm(y ~ .,
           family = binomial,
           data = data.frame(X[,var_name],y))
  p <- predict(fit, type = "response")
  out <- pr_stats(y, p, title)
  out$aupr
  out$plot
}


#plot_grid(
#  plot_PR(best_BIC_minimizer_all_lambda$best$variables, "AUC-PR: Best Model"),
#  plot_PR(best_simplest_model_varnames, "AUC-PR: Best Simplest Model")
#)



```

### across all elite models

```{r}
cv_auc_pr_elite_models = lapply(1:length(varnames_elites),
                        function(i){
                          cat(sprintf("Begin Elite %d\n", i))
                          cv_pr = cv_aucpr(X, y, varnames_elites[[i]])
                          cat("Average AUC-PR for Elite ", i, "=", cv_pr$avg_aucpr, "\n\n")
                          return(cv_pr)
                      }
              )
```


```{r}
cv_auc_pr_elite_models_avgAucPr = sapply(1:length(cv_auc_pr_elite_models),
       function(i) cv_auc_pr_elite_models[[i]]$avg_aucpr)

which.max(cv_auc_pr_elite_models_avgAucPr)

#varnames_elites[[which.max(cv_auc_pr_elite_models_avgAucPr)]]



```



## Finding the best model by validation

```{r}
which.min(avg_cv_logloss_elite_models_df$avg_ll)
```



```{r}

auc_prs = c()
avg_lls = c()

for(i in 1:length(auc_pr_elite_models)){
  auc_prs = c(auc_prs, cv_auc_pr_elite_models_avgAucPr[i])
  avg_lls = c(avg_lls, avg_cv_logloss_elite_models[[i]]$avg_ll)
}


elite_vals_df = data.frame(auc_pr = auc_prs,
                           avg_ll = avg_lls,
                           elite_idx = 1:length(auc_pr_elite_models))

```


### latex: min avgll and max aucpr

```{r}
xtable(
  elite_vals_df %>% arrange(avg_ll, desc(auc_pr)),
  digits = c(0,4,4,0)
  
)
```



```{r}
elite_vals_df %>% arrange(avg_ll)
```

```{r}
varnames_elites[[25]]
```

## Elite model 25 is the winner!

```{r}
varnames_elites[[25]]
```


## Best elite model

```{r}

best_elite_model = glm(y ~ .,
    data=data.frame(scale(X)[,varnames_elites[[25]]], y),
    family=binomial
    )


summary(best_elite_model)


```

```{r}

setdiff(varnames_elites[[25]],
        core_vars_among_elites)


```
wind_u no prob
potential_vorticity no prob
mean_sea_level_pressure
surface_dew


### Summarize in latex
```{r}
library(broom)

# Extract tidy coefficients
coef_tbl <- broom::tidy(best_elite_model) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    coef_formatted = sprintf("%.2f (%.2f)", estimate, std.error)
  ) %>%
  arrange(desc(abs(estimate))) %>%
  select(variable = term, coefficient = coef_formatted)


xtable(coef_tbl)

# Display neatly
knitr::kable(coef_tbl,
             caption = "Final elite model: Coefficients (Estimate with SE in parentheses)",
             align = c("l", "r"))

```


## Correlation ~ OOS performance

### Examine correlations in best elite model

```{r}
melt(round(cor(X[,varnames_elites[[25]]]), 2)) %>%
  filter(value != 1) %>%
  arrange(desc(value)) %>% filter(Var1 == "geopotential_at_925_hPa")

```

### reduction in cv log loss

#### For vertical_wind_shear_between_300_hPa_and_1000_hPa

```{r}
for (shear_var in grep("vertical_wind_shear", varnames_elites[[25]], value=TRUE)){
  
  best_elite_set_no_shear_var = varnames_elites[[25]][! varnames_elites[[25]] %in%
                                      c(shear_var)]
  
  cv_log_loss_anomaly_shear = cv_logloss(X, y,
             best_elite_set_no_shear_var
  ) 
  
  cat(sprintf("OOS log-loss after removing %s: %.3f, with BIC reduction: %.3f \n",
              shear_var,
              cv_log_loss_anomaly_shear$avg_ll - min(avg_lls),
              compute.BIC.auto(data.frame(X,y), 
                 colnames(X) %in% best_elite_set_no_shear_var) - best_BIC_minimizer_all_lambda$best$BIC))
  
  
  
}
```



```{r}
cv_log_loss_anomaly_shear = cv_logloss(X, y,
           varnames_elites[[25]][! varnames_elites[[25]] %in% c("vertical_wind_shear_between_300_hPa_and_1000_hPa")]
) 
```



```{r}
cv_log_loss_anomaly_shear$avg_ll - min(avg_lls)
```

```{r}
cv_aucpr_anomaly_shear = cv_aucpr(X, y,
           varnames_elites[[25]][! varnames_elites[[25]] %in% c("vertical_wind_shear_between_300_hPa_and_1000_hPa")]
) 
```

```{r}
cv_aucpr_anomaly_shear$avg_aucpr - max(auc_prs, na.rm=TRUE)
```


```{r}
compute.BIC.auto(data.frame(X,y), 
                 colnames(X) %in% varnames_elites[[25]][! varnames_elites[[25]] %in% c("vertical_wind_shear_between_300_hPa_and_1000_hPa")]) - best_BIC_minimizer_all_lambda$best$BIC
```

```{r}
compute.BIC.auto(data.frame(X,y), 
                 colnames(X) %in% varnames_elites[[1]][! varnames_elites[[1]] %in% c("vertical_wind_shear_between_300_hPa_and_1000_hPa")]) - best_BIC_minimizer_all_lambda$best$BIC
```



#### geopotential_at_925_hPa and mean_sea_level_pressure

```{r}

cv_logloss(X, y,
           varnames_elites[[25]][! varnames_elites[[25]] %in% c("geopotential_at_925_hPa")]
) 


cv_logloss(X, y,
           varnames_elites[[25]][! varnames_elites[[25]] %in% c("mean_sea_level_pressure")]
)

```

### Reduction in BIC

```{r}

compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]][! varnames_elites[[25]] %in%
                                                    c("geopotential_at_925_hPa")]
) - best_BIC_minimizer_all_lambda$best$BIC


compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]][! varnames_elites[[25]] %in%
                                                    c("geopotential_at_925_hPa")]
) - compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]]
)

```

```{r}

compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]][! varnames_elites[[25]] %in%
                                                    c("mean_sea_level_pressure")]
) - best_BIC_minimizer_all_lambda$best$BIC


compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]][! varnames_elites[[25]] %in%
                                                    c("mean_sea_level_pressure")]
) - compute.BIC.auto(data.frame(X,y),
           colnames(X) %in% varnames_elites[[25]]
)

```


```{r}
head(
  elite_vals_df %>% arrange(avg_ll),
  n=1
)

head(
  elite_vals_df %>% arrange(desc(auc_pr)),
  n=1
)

```



## Validate with LASSO

Ensure that both RS and LASSO, in the cross validation, uses the same folds

```{r}
make_folds <- function(y, k = 5, seed = 123) {
  set.seed(seed)
  # caret returns a list of indices per fold; we’ll convert to a foldid vector
  idx_list <- createFolds(y, k = k, list = TRUE, returnTrain = FALSE)
  foldid <- integer(length(y))
  for (i in seq_along(idx_list)) foldid[idx_list[[i]]] <- i
  foldid
}

binom_deviance_per_obs <- function(y_true, p_hat, eps = 1e-15) {
  p_hat <- pmin(pmax(p_hat, eps), 1 - eps)
  # deviance = -2 * sum(y log p + (1-y) log (1-p))
  -2 * mean(y_true * log(p_hat) + (1 - y_true) * log(1 - p_hat))
}


cv_deviance_rs <- function(X, y, vars, foldid) {
  k <- max(foldid)
  scores <- numeric(k)
  for (f in 1:k) {
    test_idx  <- which(foldid == f)
    train_idx <- setdiff(seq_along(y), test_idx)

    df_tr <- data.frame(scale(X)[train_idx, vars, drop=FALSE], y = y[train_idx])
    df_te <- data.frame(scale(X)[test_idx,  vars, drop=FALSE], y = y[test_idx])

    fit <- glm(y ~ ., data = df_tr, family = binomial)
    p_hat <- predict(fit, newdata = df_te, type = "response")
    scores[f] <- binom_deviance_per_obs(df_te$y, p_hat)
  }
  list(mean = mean(scores), sd = sd(scores), per_fold = scores, num_vars=length(vars))
}

## 4) CV for LASSO on the same folds (uses *all* predictors by design)
cv_deviance_lasso <- function(X, y, foldid) {
  # glmnet expects matrix X and vector y; foldid aligns with rows of X
  fit_cv <- cv.glmnet(
    x = as.matrix(X), y = y,
    family = "binomial",
    type.measure = "deviance",
    foldid = foldid,
    keep = TRUE # keep per-fold predictions if you want to inspect later
  )
  # glmnet's cvm is mean deviance per observation
  list(mean = as.numeric(min(fit_cv$cvm)),
       sd   = as.numeric(fit_cv$cvsd[which.min(fit_cv$cvm)]),
       model = fit_cv)
}

```


```{r}

# 1) folds
foldid <- make_folds(y, k = 5, seed = 123)

# 2) RS model CV deviance
rs_out <- cv_deviance_rs(X, y, vars = varnames_elites[[25]], foldid = foldid)

# 3) LASSO CV deviance (on all X)
lasso_out <- cv_deviance_lasso(X, y, foldid = foldid)


```

```{r}
cat(sprintf("RS (fixed vars) CV deviance:   mean=%.4f (sd=%.4f)\n", rs_out$mean, rs_out$sd))
cat(sprintf("LASSO (glmnet) CV deviance:    mean=%.4f (sd=%.4f)\n", lasso_out$mean, lasso_out$sd))
```



```{r}
plot_OOS_deviances = function(rs_out, lasso_out, foldid){
  # Results
  desc_rs = sprintf("RS (fixed var set),\nWith %d Variables\non full model", rs_out$num_vars)
  desc_lasso = sprintf("LASSO (best lambda=%.4f),\nWith 81 Variables\non full model",
                       lasso_out$model$lambda.1se)
  
  methods <- c(desc_rs, desc_lasso)
  means   <- c(rs_out$mean, lasso_out$mean)
  sds     <- c(rs_out$sd, lasso_out$sd)
  nfold   <- length(unique(foldid))
  df      <- nfold - 1
  
  # Critical t value for 95% CI
  tval <- qt(0.975, df)
  
  # Compute CI half-widths
  errors <- tval * sds / sqrt(nfold)
  
  # Put into a data.frame for plotting
  df_plot <- data.frame(
    method = methods,
    mean   = means,
    lower  = means - errors,
    upper  = means + errors
  )
  
  df_plot$mean = round(df_plot$mean, 3)
  
  #vars <- c("50 variables\npicked", "81 variables\npicked")  # text annotations
  #df_plot$vars <- vars
  
  
  
  
  ###
  
  palette <- c(desc_lasso = "#2563EB",  # blue
               desc_rs = "#DC2626") # red
  
  ggplot(df_plot, aes(x = method, y = mean)) +
    
    # bars (colored)
    geom_col(aes(fill = method), width = 0.5, alpha = 0.35, color='grey') +
    
    # error bars (always black)
    geom_errorbar(aes(ymin = lower, ymax = upper), 
                  width = 0.12, linewidth = 1, color = "black") +
    
    # points (colored)
    geom_point(aes(color = method), size = 2.5) +
    
    # mean labels (colored)
    geom_text(aes(label = mean, color = method), 
              vjust = 0.7, hjust = -0.5, size = 4.5, fontface = "bold") +
    
    # manual colors
    scale_fill_manual(values = palette, guide = "none") +
    scale_color_manual(values = palette, guide = "none") +
    
    labs(
      title = "Comparison of RS vs LASSO",
      y = "CV Deviance (mean ± 95% CI)", x = NULL
    ) +
    
    coord_cartesian(ylim = c(0, max(df_plot$upper) + 0.05)) +
    
    theme_minimal(base_size = 14) +
    theme(
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(margin = margin(t = 6))
    ) 
}


```


```{r}
plot_OOS_deviances(rs_out, lasso_out, foldid)
```

## Validate core vars?
```{r}
cv_logloss(X, y, vars=core_vars_among_elites)
```

```{r}
cv_aucpr(X, y, vars=core_vars_among_elites)
```



# Coef distributions of core elites across elite models
```{r}
length(varnames_elites)

varnames_elites[[1]]
```



```{r}
core_vars_among_elites
```
## Obtain elite coefs

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)

# Suppose you have:
# - varnames_elites: list of length = number of elite models
#   (each entry is a character vector of variable names in that elite model)
# - X: your predictor matrix (already standardized)
# - y: response vector

# -----------------------
# 1. Fit glm to each elite model
# -----------------------
elite_coefs <- lapply(seq_along(varnames_elites), function(i) {
  vars <- varnames_elites[[i]]
  
  df <- data.frame(y = y, scale(X)[, vars, drop = FALSE])
  fit <- glm(y ~ ., family = binomial, data = df)
  
  broom::tidy(fit) %>%
    filter(term != "(Intercept)") %>%
    mutate(model_id = i)
})

elite_coefs_df <- bind_rows(elite_coefs)

# -----------------------
# 2. Filter to only core vars
# -----------------------
core_coefs_df <- elite_coefs_df %>%
  filter(term %in% core_vars_among_elites)

# -----------------------
# 3. Visualization (see below)
# -----------------------

```
## Boxplots
```{r}
ggplot(core_coefs_df %>% filter(estimate >= 0),
       aes(x = reorder(term, estimate, FUN=var, decreasing=TRUE), y = estimate)) +
  geom_boxplot(outlier.alpha = 0.3, fill='red') +
  theme_minimal() +
  coord_flip() +
  labs(title = "Coefficient variability across elite models (core vars)",
       x = "Variable", y = "Coefficient estimate")

ggplot(core_coefs_df %>% filter(estimate < 0),
       aes(x = reorder(term,estimate, FUN=var, decreasing=TRUE), y = estimate)) +
  geom_boxplot(outlier.alpha = 0.3, fill='red') +
  theme_minimal() +
  coord_flip() +
  labs(title = "Coefficient variability across elite models (core vars)",
       x = "Variable", y = "Coefficient estimate")
```
## Density plot with uncertainty

```{r}
library(dplyr)
library(ggplot2)

# Optional: unimodality score in uncertainty
if (!requireNamespace("diptest", quietly = TRUE)) install.packages("diptest")

# ----------------------------- stats + sign class -----------------------------
# core_coefs_df: columns term, estimate
term_span <- core_coefs_df %>%
  group_by(term) %>%
  summarise(min_est = min(estimate, na.rm=TRUE),
            max_est = max(estimate, na.rm=TRUE),
            .groups="drop") %>%
  mutate(sign_group = dplyr::case_when(
    min_est >  0 ~ "positive-only",
    max_est <  0 ~ "negative-only",
    TRUE         ~ "mixed"
  ))

term_stats <- core_coefs_df %>%
  group_by(term) %>%
  summarise(
    n     = n(),
    mean  = mean(estimate, na.rm=TRUE),
    sd    = sd(estimate, na.rm=TRUE),
    iqr   = IQR(estimate, na.rm=TRUE),
    mad   = mad(estimate, constant = 1, na.rm=TRUE),
    dip_p = tryCatch(diptest::dip.test(estimate)$p.value, error = function(e) NA_real_),
    .groups = "drop"
  ) %>%
  mutate(
    spread_score = scales::rescale(iqr, to = c(0, 1), from = range(iqr, na.rm=TRUE)),
    mode_score   = 1 - pmin(pmax(dip_p, 0), 1),         # low p = more multi-modal
    uncertainty  = 0.7*spread_score + 0.3*mode_score    # tweak weights if desired
  )

coef_df <- core_coefs_df %>%
  left_join(term_span,  by="term") %>%
  left_join(term_stats, by="term")


#

```


```{r}
plot_coef_dens <- function(df, title, show_zero = FALSE, ncol = 4, var_text_size=7){
  if (!nrow(df)) return(NULL)
  ggplot(df, aes(x = estimate, group = term, color = uncertainty, fill = uncertainty)) +
    geom_density(alpha = 0.22, linewidth = 0.6) +
    facet_wrap(~ reorder(term, uncertainty), 
               scales = "free", ncol = ncol, 
               labeller = labeller(term = label_wrap_gen(width = 20))) +
    scale_color_gradient(low = "#00A676", high = "#D7263D") +
    scale_fill_gradient( low = "#00A676", high = "#D7263D") +
    { if (show_zero) geom_vline(xintercept = 0, linetype = 2, color = "grey55") } +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid      = element_blank(),
      #legend.position = "none",
      axis.title      = element_blank(),
      axis.text.y     = element_blank(),
      axis.ticks      = element_blank(),
      strip.text      = element_text(size = var_text_size),            # smaller strip titles
      panel.spacing   = unit(0.3, "lines"),                # tighter grid
      plot.title      = element_text(size=14, face="bold"),
      plot.subtitle   = element_text(size=10)
    ) +
    labs(
      title    = title,
      subtitle = "Color encodes uncertainty = 0.7*scaled IQR + 0.3*(1 - dip-test p-value)"
    ) + theme(axis.text.x = element_text(angle = 60, hjust = 1)) # Adjust angle and justification
}


# ----------------------------- three panels -----------------------------------
p_pos  <- coef_df %>% filter(sign_group == "positive-only") %>%
  plot_coef_dens("Coefficient distributions across elite models (β > 0)")

p_neg  <- coef_df %>% filter(sign_group == "negative-only") %>%
  plot_coef_dens("Coefficient distributions across elite models (β < 0)")

p_mix  <- coef_df %>% filter(sign_group == "mixed") %>%
  plot_coef_dens("Coefficient distributions across elite models (mixed: crosses zero)", show_zero = TRUE, var_text_size=10)

# print whatever exists
p_pos; p_neg; p_mix

# optional: save
ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "coef_pos.png"),
       p_pos, width=12, height=8, dpi=300, bg = 'white')
ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "coef_neg.png"),
       p_neg, width=12, height=8, dpi=300, bg = 'white')
ggsave(here("output", "mh_runs_4_lambda_run_2_plots", "coef_mix.png"), 
       p_mix, width=12, height=8, dpi=300, bg = 'white')
```


# Presenting final model
## Weighted LR and HL Test
```{r}

weighted_lr_and_HL_test <- function(vars, weight = TRUE, want_p_adj = FALSE, g = 10) {
  stopifnot(exists("X"), exists("y"))  # expect X, y in env
  
  library(ResourceSelection)
  library(dplyr)
  
  # 1) data for chosen subset
  df <- data.frame(X[, vars, drop = FALSE], y = y)
  
  # 2) class weights (minority gets larger weight)
  if (weight) {
    wts <- ifelse(df$y == 1, sum(df$y == 0) / sum(df$y == 1), 1)
    fit <- glm(y ~ ., data = df, family = binomial, weights = wts)
  } else {
    fit <- glm(y ~ ., data = df, family = binomial)
  }
  #print(summary(fit))
  
  # fitted probs from the (possibly weighted) fit
  p_hat <- as.numeric(predict(fit, type = "response"))
  
  # 3) optional intercept adjustment (recalibration)
  if (want_p_adj) {
    # linear predictor from the fitted model
    linpred <- as.numeric(predict(fit, type = "link"))
    
    # re-estimate only the intercept via an offset model
    # (use *unweighted* here to calibrate to the empirical prevalence)
    adj_fit <- glm(y ~ offset(linpred), data = df, family = binomial)
    
    # swap the intercept by NAME
    beta_adj <- coef(fit)
    beta_adj[1] <- coef(adj_fit)[1]
    
    # recompute eta using the *same* design matrix as the original fit
    MM <- model.matrix(fit)
    #eta_adj <- drop(MM %*% beta_adj)
    eta_adj <- MM %*% beta_adj
    print(eta_adj)
    p_hat   <- plogis(eta_adj)                 # = 1 / (1 + exp(-eta_adj))
    
    #print(p_hat)
  }
  
  # 4) numerical clipping for safety
  eps   <- 1e-15
  p_hat <- pmin(pmax(p_hat, eps), 1 - eps)
  
  # 5) HL test (note: with very large n, HL often rejects even well-behaved models)
  hl <- hoslem.test(df$y, p_hat, g = g)
  print(hl)
  
  list(
    fit     = fit,
    p_hat   = p_hat,
    hl      = hl,
    coef    = coef(fit),
    coef_adj= if (want_p_adj) beta_adj else NULL
  )
}

prior_corrected_logit <- function(vars, tau, weight = TRUE) {
  library(ResourceSelection)

  df <- data.frame(X[, vars, drop = FALSE], y = y)
  ybar <- mean(df$y)

  # Fit logistic model
  if (weight) {
    wts <- ifelse(df$y == 1, sum(df$y == 0)/sum(df$y == 1), 1)
    fit <- glm(y ~ ., data = df, family = binomial, weights = wts)
  } else {
    fit <- glm(y ~ ., data = df, family = binomial)
  }

  beta_hat <- coef(fit)
  beta0_hat <- beta_hat[1]

  # Apply King & Zeng prior correction to intercept
  beta0_corr <- beta0_hat - log(((1 - tau) / tau) * (ybar / (1 - ybar)))
  beta_corr  <- beta_hat
  beta_corr[1] <- beta0_corr

  # Recompute predicted probabilities
  eta_corr <- as.matrix(cbind(1, df[, vars])) %*% beta_corr
  p_hat_corr <- plogis(eta_corr)

  # Hosmer–Lemeshow test
  hl <- hoslem.test(df$y, p_hat_corr)

  list(
    fit = fit,
    beta_hat = beta_hat,
    beta_corr = beta_corr,
    intercept_shift = beta0_corr - beta0_hat,
    hl = hl,
    p_hat_corr = p_hat_corr
  )
}


```

```{r}
final_model_prior_corrected = prior_corrected_logit(varnames_elites[[25]],
                                                    0.089,
                                                    weight=FALSE)

final_model_prior_corrected$hl
```


## Calibration plot

```{r}
final_model = weighted_lr_and_HL_test(varnames_elites[[25]],
                                      want_p_adj = FALSE,
                                      weight=FALSE,
                                      g=3)
```

```{r}



calib_df = data.frame(predict.glm(final_model$fit, type = 'response'), y)
calib_df = calib_df %>% rename(all_of(c(pred = colnames(calib_df)[1])))


calib_plot = calibration_plot(data = calib_df,
                 obs = "y", pred = "pred",
                 title = "Calibration plot for the best model",
                 y_lim = c(0, 0.6), x_lim=c(0, 0.6),
                 data_summary = TRUE,
                 nTiles=10)
calib_plot$calibration_plot +
  geom_point(color = "blue", size=6, alpha=0.3) + 
  theme(text = element_text(size = 23))
```
```{r}
calib_plot$data_summary
```

```{r}
xtable(calib_plot$data_summary,
       digits = c(0, 0, 6, 6, 0, 6, 6, 6))
```



```{r}

calib_plot_from_model = function(fit, nTiles=10){
  calib_df = data.frame(predict.glm(fit, type = 'response'), y)
  calib_df = calib_df %>% rename(all_of(c(pred = colnames(calib_df)[1])))
  
  
  calib_plot = calibration_plot(data = calib_df,
                   obs = "y", pred = "pred",
                   title = "Calibration plot",
                   y_lim = c(0, 1), x_lim=c(0, 1),
                   data_summary = TRUE,
                   nTiles=nTiles)
  
  calib_plot$calibration_plot + geom_point(color = "blue")
  
}

```



# Reassessing convergence

## Rationale

## lam=0.75


### BIC boxplot grouped by chain id
```{r}

distribution_bic_traces = function(){
  
  K = length(mh_lambda_0.10_chain_1$rs$bic_trace)
  
  bic_traces_lam_0.1 = c(
    mh_lambda_0.10_chain_1$rs$bic_trace,
    mh_lambda_0.10_chain_2$rs$bic_trace,
    mh_lambda_0.10_chain_3$rs$bic_trace,
    mh_lambda_0.10_chain_4$rs$bic_trace
  )
  chain_desc = c("1: full", "2: empty", "3: sparse", "4: random")
  
  bic_traces_lam_0.1_ggplot = data.frame(
    bic_traces = bic_traces_lam_0.1,
    chain_id = factor(rep(chain_desc, each=K), levels=chain_desc),
    iteration = rep(1:K, times=4)
  )
  
  bic_traces_lam_0.75 = c(
    mh_lambda_0.75_chain_1$rs$bic_trace,
    mh_lambda_0.75_chain_2$rs$bic_trace,
    mh_lambda_0.75_chain_3$rs$bic_trace,
    mh_lambda_0.75_chain_4$rs$bic_trace
  )
  
  bic_traces_lam_0.75_ggplot = data.frame(
    bic_traces = bic_traces_lam_0.75,
    chain_id = factor(rep(chain_desc, each=K), levels=chain_desc),
    iteration = rep(1:K, times=4)
  )
  p_0.1 = ggplot(bic_traces_lam_0.1_ggplot, aes(x = bic_traces, y=chain_id)) +
    geom_boxplot(position = position_dodge(width = 0.8)) +
    labs(title = "Distribution of BIC traces by chain for lam=0.1", x = "BIC", y = "Chain id") +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
  
  p_0.75 = ggplot(bic_traces_lam_0.75_ggplot, aes(x = bic_traces, y=chain_id)) +
    geom_boxplot(position = position_dodge(width = 0.8)) +
    labs(title = "Distribution of BIC traces by chain for lam=0.75", x = "BIC", y = "Chain id") +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
    geom_vline(xintercept = 9590,
              color = "red", linetype = "dashed")
  
  list(p_0.1 = p_0.1,
       p_0.75 = p_0.75,
       bic_traces_lam_0.75_ggplot = bic_traces_lam_0.75_ggplot,
       bic_traces_lam_0.1_ggplot = bic_traces_lam_0.1_ggplot)
  
}
distribution_bic_traces = distribution_bic_traces()
distribution_bic_traces$p_0.1
distribution_bic_traces$p_0.75

```


### ANOVA: BIC ~ chain_id
```{r}

df = distribution_bic_traces$bic_traces_lam_0.75_ggplot

df = df[df$bic_traces < 10700,]

anova(
  lm(bic_traces ~ iteration + chain_id,
     data = df)
)

anova(
  lm(bic_traces ~ chain_id,
     data = df)
)

```


### Normalized R hat 

```{r}
library(posterior)
BIC_by_chain <- split(df$bic_traces, df$chain_id)
# build draws_array: iterations x chains x 1
K   <- min(sapply(BIC_by_chain, length))
arr <- array(unlist(lapply(BIC_by_chain, \(z) z[1:K])), dim=c(K, length(BIC_by_chain), 1))
draws <- as_draws_array(arr)
rhat  <- rhat(draws)     # rank-normalized R-hat
rhat

```
```{r}
rhat_basic(draws)
```


## Normalized R hat for all lambda

```{r}
list.files(here("output", "mh_runs_4_lambda_run_2" ))
```


```{r}
library(here)
library(posterior)


files <- list.files(
  here("output", "mh_runs_4_lambda_run_2"),
  pattern = "^mh_lambda_[0-9\\.]+_chain_[1-4]_", full.names = TRUE
)

lambdas <- sort(unique(as.numeric(sub(".*mh_lambda_([0-9\\.]+)_chain_.*", "\\1",
                                      basename(files)))))

rhat_by_lambda <- numeric(length(lambdas))

for (i in seq_along(lambdas)) {
  lam <- lambdas[i]
  s = sprintf("mh_lambda_%.2f_chain_", lam)
  chain_files <- grep(paste0(s), files, value = TRUE)

  if (length(chain_files) != 4) {
    stop(sprintf("λ=%s: expected 4 chain files, found %d", lam, length(chain_files)))
  }

  # Read BIC traces; coerce to numeric and validate
  bic_traces <- lapply(chain_files, function(f) {
    obj <- readRDS(f)
    bt  <- obj$rs$bic_trace
    if (is.null(bt)) stop(sprintf("No $rs$bic_trace in file: %s", f))
    as.numeric(bt)
  })

  # lengths as a numeric vector (vapply is strict)
  lens <- vapply(bic_traces, length, integer(1))
  K    <- min(lens, na.rm = TRUE)

  # build iterations x chains matrix
  bic_mat <- do.call(cbind, lapply(bic_traces, `[`, seq_len(K)))

  # to draws_array and R-hat (rank-normalized)
  arr   <- array(bic_mat, dim = c(K, ncol(bic_mat), 1))
  draws <- as_draws_array(arr)
  rhat_by_lambda[i] <- rhat(draws)
}

rhat_df = data.frame(lambda = lambdas, rhat = rhat_by_lambda)

print(rhat_df)

# Optional: visualize
library(ggplot2)
ggplot(rhat_df, aes(x = lambda, y = rhat)) +
  geom_point(size = 3) + geom_line() +
  #geom_hline(yintercept = 1.01, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 1.01, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Rank-normalized R-hat by λ",
       y = "R-hat (rank-normalized)", x = expression(lambda))

```

```{r}

S_list_MH = list()
i=1

S_list_MH[[i]] = mh_lambda_0.75_chain_1$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_2$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_3$rs$bic_trace; i = i + 1
S_list_MH[[i]] = mh_lambda_0.75_chain_4$rs$bic_trace; i = i + 1

best_bic = best_BIC_minimizer_all_lambda$best$BIC

lapply(1:4,
       function(j){
         which( S_list_MH[[j]] - best_bic <= 2 )
       })

```


```{r}

vi_trace_plot = function(bottom_ess=1){
  
  inclusion_order_by_ess = inclusion_summary_lam_0.75 %>%
                                  filter(ess > 0) %>%
                                  arrange(ess)
  
  c = inclusion_order_by_ess$ess[bottom_ess]
  
  
  plot(mh_lambda_0.75_chain_1$rs$samples[,c],
     type='l', col=1)
  
  lines(mh_lambda_0.75_chain_2$rs$samples[,c],
       col=2)
  lines(mh_lambda_0.75_chain_3$rs$samples[,c],
       col=3)
  lines(mh_lambda_0.75_chain_4$rs$samples[,c],
       col=4)
}



inclusion_summary_lam_0.75 %>% filter(ess > 0) %>% arrange(ess)
```

```{r}

vi_trace_plot = function(bottom_ess=1){
  
  inclusion_order_by_ess = inclusion_summary_lam_0.75 %>%
                                  filter(ess > 0) %>%
                                  arrange(ess)
  
  c = inclusion_order_by_ess$variable[bottom_ess]
  bottom_ess_value = inclusion_order_by_ess$ess[bottom_ess]
  
  par(mfrow=c(2,2))
  
  xlab="Iteration"
  
  plot(mh_lambda_0.75_chain_1$rs$samples[,c],
     type='o', col=1, ylim=c(0,1), xlab=xlab,
     main=sprintf("Indicators on chain 1 (full) for %s\nhaving pooled ESS=%.2f (%d-th lowest pooled ESS)", c, bottom_ess_value, bottom_ess),
     ylab="Chain 1")
  
  plot(mh_lambda_0.75_chain_2$rs$samples[,c],
       col=1, type='o', ylim=c(0,1), xlab=xlab,
       main=sprintf("Indicators on chain 2 (empty) for %s\nhaving pooled ESS=%.2f (%d-th lowest pooled ESS)", c, bottom_ess_value, bottom_ess),
       ylab="Chain 2")
  
  plot(mh_lambda_0.75_chain_3$rs$samples[,c],
       col=1, type='o', ylim=c(0,1), xlab=xlab,
       main=sprintf("Indicators on chain 3 (sparse) for %s\nhaving pooled ESS=%.2f (%d-th lowest pooled ESS)", c, bottom_ess_value, bottom_ess),
       ylab="Chain 3")
  
  plot(mh_lambda_0.75_chain_4$rs$samples[,c],
       col=1, type='o', ylim=c(0,1), xlab=xlab,
       main=sprintf("Indicators on chain 4 (random) for %s\nhaving pooled ESS=%.2f (%d-th lowest pooled ESS)", c, bottom_ess_value, bottom_ess),
       ylab="Chain 4")
}

vi_trace_plot()

```


```{r}
inclusion_summary_lam_0.75_asy_CI = inclusion_summary_asy_CI(
    rbind(mh_lambda_0.75_chain_1$rs$samples,
          mh_lambda_0.75_chain_2$rs$samples,
          mh_lambda_0.75_chain_3$rs$samples,
          mh_lambda_0.75_chain_4$rs$samples),
                  top_k=149,
    epsilon=1e-20,
    alpha=0.1)

plot_inclusion(inclusion_summary_lam_0.75_asy_CI)
```


```{r}
incl.summary.v2$variable
```



# Attempted empirical rule for choosing lambda.


```{r}

bic_max_min_diff = function(K=1000, bic_burnin=1000,
                            chain_id=3, max_min_diff=FALSE){
  
  
  ### note burn-in has not been applied to BIC traces 

  # extract all RS results
  pattern = sprintf("^mh_lambda_[0-9\\.]+_chain_%d_",chain_id)
  files <- list.files(
    here("output", "mh_runs_4_lambda_run_2"),
    pattern = pattern, full.names = TRUE
  )
  
  # extract lambda values
  lambdas <- na.omit(str_extract(list.files(here("output", "mh_runs_4_lambda_run_2")), ".*mh_lambda_(.+)_chain_3", group=1))
  lambdas = sort(unique(as.numeric(lambdas)))
  #print(lambdas)
  
  empirical_criteria = c()
  
  
  i=1
  for (f in files){
    obj <- readRDS(f)
    
    #cat(sprintf("\nRead file: %s \n", f))
    # Read BIC traces, remove unaccounted burnin, and after that
    # take first K  BIC values for empirical rule
    
    bt  <- obj$rs$bic_trace[bic_burnin : (K+bic_burnin)]
    min_bic = min(bt)
    max_bic = max(bt)
    
    if (max_min_diff){
      cat(sprintf("For lam=%.2f, max_bic-min_bic=%.3f\n",
                lambdas[i], max_bic-min_bic))
      # obtain normalized difference between smallest and greatest BIC
      # exp(-lambda*BIC_min) / (exp(-lambda*BIC_min) + exp(-lambda*BIC_max))
      # = 1 / (1 + exp(-lambda*BIC_max + lambda*BIC_min))
      # = 1 / (1 + exp(-lambda * (BIC_max - BIC_min))
      # = plogis(lambda*(BIC_max - BIC_min))
      empirical_criteria[i] =  stats::plogis(lambdas[i]*(max_bic - min_bic))
    } else {
      empirical_criteria[i] =  min_bic
    }
    
    i = i + 1
    
  }
  names(empirical_criteria) = lambdas
  sort(empirical_criteria)

     
}
  
#bic_max_min_diff()

```

```{r}
bic_max_min_diff(K=3000, max_min_diff=TRUE)
```

```{r}
bic_max_min_diff(K=2000)
```

```{r}
bic_max_min_diff(K=3000)
```


```{r}
bic_max_min_diff(K=5000)
```


# RS on core vars?

```{r}
Gibbs.on.core.vars = Gibbs.random.search.lambda(
  v0=rep(1, length(varnames_elites[[1]])),
  data=data.frame(X[,varnames_elites[[1]]], y),
  lambda=3,
  debug=TRUE
)
```


```{r}
MH.on.elite.vars = MH.random.search(
  #v0=rep(1,length(varnames_elites[[1]])),
  v0=tail(mh_lambda_0.75_chain_3$rs$samples, n=1),
  data=data.frame(X, y),
  lambda=0.75,
  debug=TRUE
)
```


```{r}
Gibbs.on.elite.vars = Gibbs.random.search.lambda(
  #v0=rep(1,length(varnames_elites[[1]])),
  v0=rbinom(length(varnames_elites[[1]]), size=1, prob=0.5),
  data=data.frame(X[,varnames_elites[[1]]], y),
  lambda=1.5,
  debug=TRUE
)
```


# Try compute BIC for marginals > 0.5?


```{r}

vars_lam0.75_gt_0.5 = (inclusion_summary_lam_0.75 %>% filter(p_hat >= 0.5))$variable

for (k in 1:length(vars_lam0.75_gt_0.5)){
  bic_first_k_rel_freqs = compute.BIC.auto(data.frame(X, y),
                 colnames(X) %in% vars_lam0.75_gt_0.5[1:k]
                 )
  cat(sprintf("BIC for first %d vars with highest rel freqs: %.3f\n", k, bic_first_k_rel_freqs))
}




```
# Proportion of top k marginal freqs containing core vars

```{r}

for (k in length(core_vars_among_elites):nrow(inclusion_summary_lam_0.75)){
     how_many_top_ks_match_cores = length(
              intersect(
                (inclusion_summary_lam_0.75$variable)[1:k],
                core_vars_among_elites
              )
     )
     
    cat(sprintf("Proportion of Top %d marginal freqs matching core vars: %.3f (i.e. %d out of %d) \n",
                k, how_many_top_ks_match_cores/length(core_vars_among_elites),
                how_many_top_ks_match_cores, length(core_vars_among_elites))
        )

    if (how_many_top_ks_match_cores == length(core_vars_among_elites)){
      print(k)
      print((inclusion_summary_lam_0.75$variable)[1:k])
      break
    }
}



```

```{r}

setdiff(
    core_vars_among_elites, 
    (inclusion_summary_lam_0.75$variable)[1:41]
)
```


# Stepwise?
```{r}
basemod = glm(y ~ 1, data=data.frame(X,y), family=binomial)
stepmod = step(basemod,
               scope=as.formula(paste("~ ", paste(colnames(X), collapse= "+"))),
               k=log(nrow(X)))
```

## Compute its BIC value

```{r}
#saveRDS(stepmod,
#        here("output", "validation_outputs", "stepmod.rds"))
```


```{r}

compute.BIC.auto(data.frame(X,y),
                 colnames(X) %in% names(stepmod$coefficients))


```
## Compute CV log loss, AUCPR

```{r}
cv_logloss(X, y, colnames(X) %in% names(stepmod$coefficients))
```
```{r}
cv_aucpr(X, y, colnames(X) %in% names(stepmod$coefficients))
```

```{r}
cv_deviance_stepwise = cv_deviance_rs(X, y,
                     vars = names(stepmod$coefficients)[!names(stepmod$coefficients) %in% c("(Intercept)")],
                     foldid = foldid)
cv_deviance_stepwise
```


# Proportion of top k marginal freqs containing core vars

```{r}



for (k in length(varnames_elites[[25]]):nrow(inclusion_summary_lam_0.75)){
     how_many_top_ks_match_best_elites = length(
              intersect(
                (inclusion_summary_lam_0.75$variable)[1:k],
                varnames_elites[[25]]
              )
     )
     
    cat(sprintf("Proportion of Top %d marginal freqs matching best elite vars: %.3f (i.e. %d out of %d) \n",
                k, how_many_top_ks_match_best_elites/length(varnames_elites[[25]]),
                how_many_top_ks_match_best_elites, length(varnames_elites[[25]]))
        )

    if (how_many_top_ks_match_best_elites == length(varnames_elites[[25]])){
      print(k)
      print((inclusion_summary_lam_0.75$variable)[1:k])
      break
    }
}



```

```{r}

compute.BIC.auto(
  data.frame(X,y),
  colnames(X) %in% (inclusion_summary_lam_0.75 %>% filter(p_hat >= 0.5))$variable
) / 2


```

```{r}

compute.BIC.auto(
  data.frame(X,y),
  core_vars_among_elites
) /2


```



```{r}
p_chart_slowest_converging_vars(mh_lambda_0.75_chain_1$rs$samples,
                                metric='MCSE')
```


# Top pressure-level core variables

```{r}
compute.BIC.auto(data.frame(X,y),
                 colnames(X) %in%
                   c("Coriolis_parameter_absolute",
                     "vorticity_relative_at_700_hPa",
                     "medium_cloud_cover",
                     "wind_u_at_1000_hPa",
                     "divergence",
                     "vertical_velocity_at_700_hPa",
                     "potential_vorticity_at_850_hPa",
                     "temperature",
                     "geopotential_at_500_hPa",
                     "vertical_wind_shear_between_925_hPa_and_1000_hPa",
                     "potential_vorticity_at_400_hPa",
                     "wind_v_at_200_hPa",
                     "sea_surface_temperature",
                     "vorticity_relative_at_200_hPa",
                     "vertical_wind_shear_between_200_hPa_and_850_hPa",
                     "wind_u_at_925_hPa",
                     "vorticity_relative_at_850_hPa",
                     "specific_humidity_at_200_hPa",
                     "divergence_at_700_hPa"))
```



```{r}
cv_logloss(X, y,
           vars=c("Coriolis_parameter_absolute",
                     "vorticity_relative_at_700_hPa",
                     "medium_cloud_cover",
                     "wind_u_at_1000_hPa",
                     "divergence",
                     "vertical_velocity_at_700_hPa",
                     "potential_vorticity_at_850_hPa",
                     "temperature",
                     "geopotential_at_500_hPa",
                     "vertical_wind_shear_between_925_hPa_and_1000_hPa",
                     "potential_vorticity_at_400_hPa",
                     "wind_v_at_200_hPa",
                     "sea_surface_temperature",
                     "vorticity_relative_at_200_hPa",
                     "vertical_wind_shear_between_200_hPa_and_850_hPa",
                     "wind_u_at_925_hPa",
                     "vorticity_relative_at_850_hPa",
                     "specific_humidity_at_200_hPa",
                     "divergence_at_700_hPa"))
```



# Claim: non-core variables having p_hat >= 0.5 are just included in suboptimal models.

```{r}
non_core_vars_phat_gt_0.5 = (inclusion_summary_lam_0.75 %>%
  filter(!variable %in% core_vars_among_elites) %>%
  filter(p_hat >= 0.5))$variable
```

```{r}
(inclusion_summary_lam_0.75 %>%
  filter(!variable %in% core_vars_among_elites) %>%
  filter(p_hat >= 0.5))
```







```{r}

#non_core_vars_phat_gt_0.5 = (inclusion_summary_lam_0.75 %>%
#  filter(!variable %in% core_vars_among_elites) %>%
#  filter(p_hat >= 0.5))$variable


#mh_lambda_0.75_chains = list(
#  mh_lambda_0.75_chain_1,
#  mh_lambda_0.75_chain_2,
#  mh_lambda_0.75_chain_3,
#  mh_lambda_0.75_chain_4
#)

n_chains <- length(mh_lambda_0.75_chains)
burn     <- 1000L

#chain_idxs = seq_len(n_chains)
chain_idxs = c(3)

for (non_core_var in non_core_vars_phat_gt_0.5) {
  for (chain_idx in chain_idxs) {
    chains <- mh_lambda_0.75_chains[[chain_idx]]
    cat(sprintf("\nInvestigating %s in chain %d\n", non_core_var, chain_idx))

    # post-burn BIC; check alignment
    bic_post <- chains$rs$bic_trace[(burn+1):length(chains$rs$bic_trace)]
    stopifnot(nrow(chains$rs$samples) == length(bic_post))

    # obtain indices
    sel_idx <- which(chains$rs$samples[, non_core_var] == 1L)

    if (length(sel_idx) == 0L) {
      prop_near <- 0
    } else {
      # choose your baseline: global best or best at this λ
      delta <- bic_post[sel_idx] - best_BIC_minimizer_all_lambda$best$BIC
      
      # HOW many distinct near optimal models among all models containing non_core_var
      n_distinct_nearoptimal_mods = length(
        unique(
          apply(chains$rs$samples[which(delta <= 2), ],
            MARGIN=1,
            function(x)paste0(x,collapse=""))
        )
      )
      # HOW many distinct models containing non_core_var
      n_distinct_models = length(
        unique(
          apply(chains$rs$samples[sel_idx, ],
            MARGIN=1,
            function(x)paste0(x,collapse=""))
        )
      )
      
      prop_near <- n_distinct_nearoptimal_mods / n_distinct_models
    }
    cat(sprintf("Proportion near-optimal (ΔBIC ≤ 2): %.4f (i.e. %d DISTINCT near-optimal models out of %d DISTINCT models)\n",
                prop_near, n_distinct_nearoptimal_mods, n_distinct_models))
  }
  cat("----------------------------------------------------------\n")
}

```
# Plot core vars for presentation

```{r}
inclusion_summary_lam_0.75 %>%
  filter(variable %in% core_vars_among_elites)
```


## Sort by MAXIMUM p_hat for @ family

```{r}
library(dplyr)
library(stringr)

# 1. Filter for your core variables
core_tbl <- inclusion_summary_lam_0.75 %>%
  filter(variable %in% core_vars_among_elites)

# 2. Extract base variable (physical field) and pressure level label
core_tbl <- core_tbl %>%
  mutate(
    base = case_when(
      str_detect(variable, "vertical_wind_shear_between_") ~ "vertical_wind_shear",
      str_detect(variable, "_at_") ~ str_replace(variable, "_at_.*$", ""),
      TRUE ~ variable
    ),
    level = case_when(
      str_detect(variable, "_between_") ~ str_replace(variable, ".*_between_", "") |> str_replace_all("_", " "),
      str_detect(variable, "_at_") ~ str_replace(variable, ".*_at_", "") |> str_replace_all("_", " "),
      TRUE ~ "physical"
    )
  )

# 3. Compute max p_hat for each base variable
leaders <- core_tbl %>%
  group_by(base) %>%
  summarise(max_phat_base = max(p_hat, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(max_phat_base), base) %>%
  mutate(base_rank = row_number())

# 4. Join base ranking back to the full table
ordered_tbl <- core_tbl %>%
  left_join(leaders, by = "base") %>%
  arrange(base_rank, desc(p_hat), level) %>%
  select(base, level, p_hat, ess, se, lower, upper, variable, base_rank)


```

```{r}
ordered_tbl
```

## Take maxs only 
```{r}
winners_only <- core_tbl %>%
  group_by(base) %>%
  filter(p_hat == max(p_hat, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(leaders %>% mutate(base_rank = row_number()), by = c("base","p_hat"="max_phat_base")) %>%
  arrange(base_rank, level) %>%
  select(base, level, p_hat, variable)

```

```{r}
winners_only
```

## Begin plotting!


```{r}

#install.packages('ggtext')

# ---- packages ----
library(dplyr)
library(stringr)
library(ggplot2)
library(ggtext)      # to render bold labels on the y-axis
# ordered_tbl, winners_only, core_coefs_df are assumed to exist (as in your last message)

# ---- 1) decide which winners to emphasize (or pass your own subset) ----
emphasize_vars <- winners_only$variable       # <- or provide a character vector yourself

# ---- 2) sign summary for winners from coefficient draws across elite models ----
sign_map <- core_coefs_df %>%
  filter(term %in% winners_only$variable) %>%
  group_by(variable = term) %>%
  summarise(min_est = min(estimate, na.rm = TRUE),
            max_est = max(estimate, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(sign_cat = case_when(
    min_est > 0 ~ "positive",
    max_est < 0 ~ "negative",
    TRUE        ~ "mixed"
  ))

# nice, desaturated palette
pal <- c(positive = "#78A9CF",   # muted blue
         negative = "#E07B73",   # muted red
         mixed    = "#B8A56A")   # muted brown/olive
gray_bar <- "#D0D0D0"

# ---- 3) mark winners, attach sign category, and make plotting labels ----
plot_df <- ordered_tbl %>%
  mutate(is_winner = variable %in% emphasize_vars) %>%
  left_join(sign_map, by = c("variable")) %>%
  mutate(fill_col = ifelse(is_winner, pal[sign_cat], gray_bar),
         # pretty left labels: base (+ level if not "physical")
         label = ifelse(level == "physical",
                        base,
                        paste0(base, " (", level, ")"))
  )

# ---- 4) boldface Gray’s six canonical predictors on the y-axis ----
grays_vars <- c(
  "Coriolis_parameter_absolute",
  "vorticity_relative_at_850_hPa",      # low-level rel. vorticity (choose the one you want)
  "vertical_wind_shear_between_200_hPa_and_850_hPa",  # deep-layer shear; adjust if you prefer 200–950
  "sea_surface_temperature",
  "equiv_pot_temperature_diff_between_surface_and_500_hPa",
  "relative_humidity_at_700_hPa"        # mid-tropospheric humidity
)

plot_df <- plot_df %>%
  mutate(label_md = ifelse(variable %in% grays_vars,
                           paste0("**", label, "**"),
                           label))

# ---- 5) y ordering: follow your base_rank then within-base p_hat desc ----
plot_df <- plot_df %>%
  arrange(base_rank, desc(p_hat)) %>%
  mutate(label_md = factor(label_md, levels = rev(unique(label_md))))  # rev for horizontal bars

# ---- 6) build plot ----
p <- ggplot(plot_df, aes(x = p_hat, y = label_md)) +
  geom_col(aes(fill = I(fill_col)), width = 0.8) +
  scale_x_continuous("Inclusion probability (p̂)", limits = c(0, 1)) +
  labs(
    y = NULL,
    title = "Core variable importance by physical family & pressure level",
    subtitle = "Colored bars = dominant level(s) per family; color shows coefficient sign across elite models"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.y = element_markdown(lineheight = 1.2),   # enables bold labels
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

p


```




# DEBUGGIN PACKAGE: need >= 0.52 of xfun

```{r}
    all_packages <- installed.packages()
    package_names <- rownames(all_packages)

    packages_using_xfun <- character()

    for (pkg in package_names) {
      desc_file <- file.path(find.package(pkg), "DESCRIPTION")
      if (file.exists(desc_file)) {
        desc_content <- readLines(desc_file)
        if (any(grepl("xfun", desc_content, ignore.case = TRUE))) {
          packages_using_xfun <- c(packages_using_xfun, pkg)
        }
      }
    }

    print(packages_using_xfun)


```

```{r}
## 0) Restart R (in RStudio: Session > Restart R)

## 1) See where xfun is coming from
.libPaths()
find.package("xfun", quiet = TRUE)
packageVersion("xfun")

## 2) Remove ANY installed copies across all lib paths
for (p in .libPaths()) {
  ip <- rownames(installed.packages(lib.loc = p))
  if ("xfun" %in% ip) remove.packages("xfun", lib = p)
}

## 3) Install fresh (pick one)
install.packages("xfun", repos = "https://cloud.r-project.org", dependencies = TRUE)
# Windows/macOS binary explicitly:
# install.packages("xfun", type = "binary", repos = "https://cloud.r-project.org")

## 4) Verify
packageVersion("xfun")   # should be >= "0.52"

```

## consider multi-level dimensions

```{r}
library(dplyr)
library(ggplot2)
library(ggtext)   # for bold strip labels

# --- 1) Mark winners (dominant level per family) and attach sign category ----
sign_by_var <- core_coefs_df %>%
  group_by(term) %>%
  summarize(
    pos = all(estimate > 0),
    neg = all(estimate < 0),
    .groups = "drop"
  ) %>%
  transmute(
    variable = term,
    sign_cat = case_when(
      pos ~ "pos",
      neg ~ "neg",
      TRUE ~ "mix"
    )
  )

plot_df <- ordered_tbl %>%
  left_join(sign_by_var, by = "variable") %>%
  mutate(
    winner = variable %in% winners_only$variable,
    sign_cat = ifelse(winner, sign_cat, "other")   # gray for non-winners
  )

# --- 2) Bold Gray’s 6 canonical predictors in the facet strips ---------------
grays_six <- c(
  "Coriolis_parameter_absolute",
  "vorticity_relative",                        # family label (low-level: 200–850 hPa shows up)
  "vertical_wind_shear",
  "sea_surface_temperature",
  "equiv_pot_temperature_diff_between_surface_and_500_hPa",
  "relative_humidity"
)

# Build a custom labeller for the facet strips
strip_labs <- unique(plot_df$base)
strip_map  <- setNames(
  ifelse(strip_labs %in% grays_six,
         paste0("<b>", strip_labs, "</b>"),
         strip_labs),
  strip_labs
)

# --- 3) Factor order: families by your base_rank, levels in the order they appear
plot_df <- plot_df %>% 
  mutate(
    base  = reorder(base, base_rank),
    level = factor(level, levels = rev(unique(level)))  # reversed so top-to-bottom matches table
  )


```



```{r}
library(scales)

label_num <- function(x) sprintf("%.2f", x)   # 0.00–1.00 format

SCALE_DOWN_FACTOR=0.5
p <- ggplot(plot_df, aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 3.1, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, SCALE_DOWN_FACTOR*max(plot_df$p_hat)),
                     breaks = seq(0, 1, 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 11) +
  theme(
    # soften vertical guides; remove x labels/ticks
    panel.grid.major.x = element_line(color = "grey85", linewidth = 0.2),
    panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.2),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 10)
    ),
    legend.position = "left"
  )

p

```

```{r}

SCALE_DOWN_FACTOR=0.05
p <- ggplot(plot_df %>% filter(winner==TRUE),
            aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 3.1, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, SCALE_DOWN_FACTOR*max(plot_df$p_hat)),
                     breaks = seq(0, SCALE_DOWN_FACTOR*max(plot_df$p_hat), 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 11) +
  theme(
    # soften vertical guides; remove x labels/ticks
    panel.grid.major.x = element_line(color = "grey85", linewidth = 0.2),
    panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.2),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 5)
    ),
    legend.position = "left"
  )

p
```



```{r}

SCALE_DOWN_FACTOR=0.05
p <- ggplot(plot_df %>% filter(winner==TRUE & sign_cat=="pos"),
            aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 3.1, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, SCALE_DOWN_FACTOR*max(plot_df$p_hat)),
                     breaks = seq(0, SCALE_DOWN_FACTOR*max(plot_df$p_hat), 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 11) +
  theme(
    # soften vertical guides; remove x labels/ticks
    panel.grid.major.x = element_line(color = "grey85", linewidth = 0.2),
    panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.2),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 5)
    ),
    legend.position = "right"
  )

p
```



```{r}

SCALE_DOWN_FACTOR=0.15
p <- ggplot(plot_df %>% filter(winner==TRUE & sign_cat=="neg"),
            aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 5, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, 0.5),
                     breaks = seq(0, 0.5, 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 20) +
  theme(
    # soften vertical guides; remove x labels/ticks
    #panel.grid.major.x = element_line(color = "grey85", linewidth = 0.05),
    #panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.05),
    
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 5)
    ),
    legend.position.inside = c(0.55, 0.5),
    legend.position = "inside"
  )

p

ggsave(here("output", "mh_runs_4_lambda_run_2_plots",
            "neg_core_vars.png"),
       p, width = 10, height = 9, units = "in", bg='white',
       scale=1.3 # Scale all plot elements (text, points, spacing) with the canvas
       )
```

```{r}

SCALE_DOWN_FACTOR=0.15
p <- ggplot(plot_df %>% filter(winner==TRUE & sign_cat=="pos"),
            aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 5, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, 0.5),
                     breaks = seq(0, 0.5, 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 20) +
  theme(
    # soften vertical guides; remove x labels/ticks
    #panel.grid.major.x = element_line(color = "grey85", linewidth = 0.05),
    #panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.05),
    
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 5)
    ),
    legend.position.inside = c(0.55, 0.5),
    legend.position = "inside"
  )

p

ggsave(here("output", "mh_runs_4_lambda_run_2_plots",
            "pos_core_vars.png"),
       p, width = 10, height = 9, units = "in", bg='white',
       scale=1.3 # Scale all plot elements (text, points, spacing) with the canvas
       )
```

```{r}

SCALE_DOWN_FACTOR=0.15
p <- ggplot(plot_df %>% filter(winner==TRUE & sign_cat=="mix"),
            aes(x = SCALE_DOWN_FACTOR*p_hat, y = level, fill = sign_cat)) +
  # thinner bars
  geom_col(width = 0.38) +
  # value labels just outside the right edge of each bar
  geom_text(aes(label = label_num(p_hat)),
            hjust = -0.15, size = 5, color = "grey20") +
  facet_grid(
    rows = vars(base),
    scales = "free_y",
    space  = "free_y",
    switch = "y",
    labeller = labeller(base = strip_map)
  ) +
  scale_fill_manual(
    values = c(pos="#5BA4D0", neg="#D06A6A", mix="#B69569", other="#D9D9D9"),
    breaks = c("pos","neg","mix"), labels = c("positive","negative","mixed"), name = NULL
  ) +
  # give extra space at right so labels fit; keep subtle vertical guides
  scale_x_continuous(limits = c(0, 0.5),
                     breaks = seq(0, 0.5, 0.1),
                     expand = expansion(mult = c(0.01, 0.12))) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 20) +
  theme(
    # soften vertical guides; remove x labels/ticks
    #panel.grid.major.x = element_line(color = "grey85", linewidth = 0.05),
    #panel.grid.minor.x = element_line(color = "grey92", linewidth = 0.05),
    
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

    panel.spacing.y    = unit(10, "pt"),
    strip.placement    = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left  = element_markdown(
      angle = 0, hjust = 0, vjust = 0.5, lineheight = 1.05, margin = margin(r = 5)
    ),
    legend.position.inside = c(0.55, 0.5),
    legend.position = "inside"
  )

p

ggsave(here("output", "mh_runs_4_lambda_run_2_plots",
            "mixed_core_vars.png"),
       p, width = 10, height = 9, units = "in", bg='white',
       scale=1.3 # Scale all plot elements (text, points, spacing) with the canvas
       )
```






```{r}

```



# Investigate coef for equiv_pot?

```{r}
core_coefs_df %>%
  filter(str_detect(term, "equiv")) %>% 
  arrange(desc(estimate))
```


```{r}
best_BIC_minimizer_all_lambda$elite$BIC
```


```{r}

mod_idx_equiv_pot_coefs_gt_0 = (core_coefs_df %>%
  filter(str_detect(term, "equiv") & estimate > 0) %>% 
  arrange(desc(estimate)) )$model_id


vars_in_mod_equiv_pot_coefs_gt_0 = Reduce(intersect,varnames_elites[mod_idx_equiv_pot_coefs_gt_0])


mod_idx_equiv_pot_coefs_le_0 = (core_coefs_df %>%
  filter(str_detect(term, "equiv") & estimate <= 0) %>% 
  arrange(desc(estimate)) )$model_id

vars_in_mod_idx_equiv_pot_coefs_le_0 = Reduce(intersect,varnames_elites[mod_idx_equiv_pot_coefs_le_0])

```

```{r}
sort(
  best_BIC_minimizer_all_lambda$elite$BIC[mod_idx_equiv_pot_coefs_gt_0]
)

```

```{r}
sort(
  best_BIC_minimizer_all_lambda$elite$BIC[mod_idx_equiv_pot_coefs_le_0]
)
```
```{r}
sort(
  best_BIC_minimizer_all_lambda$elite$k[mod_idx_equiv_pot_coefs_gt_0]
)
```

```{r}
sort(
  best_BIC_minimizer_all_lambda$elite$k[mod_idx_equiv_pot_coefs_le_0]
)
```

Test: is each model on the >0 coef *nested* on any model for the >0 coef?





# BIC density plot for presentaion
```{r}
ggplot(data.frame(BIC=unlist(S_list_MH)[unlist(S_list_MH) <= 9800]),
       aes(x=BIC)) +
  geom_density(fill='red', bw=1.2, alpha=0.1) +
  theme_classic() + 
  xlim(9650, 9740) +

  labs(title="Illustration of BIC distributions \n in high dimensional model spaces",
       caption = "Illustration of 160,000 BIC values from sampled models in a 149-variable space.\nEach peak corresponds to clusters of models with similar quality.") + 
  theme(
    axis.ticks = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    plot.title = element_text(hjust = 0.5, # Center the title
                                  color = "black",
                                  size = 16,
                                  face = "bold"),
    plot.caption = element_text(size = 7.5, colour = "grey30", hjust = 0),
    plot.caption.position = "plot" )    # puts caption below panel
```


```{r}
ggplot(data.frame(BIC=unlist(S_list_MH)[unlist(S_list_MH) <= 9800]),
       aes(x=BIC)) +
  geom_density(fill='green', bw=4, alpha=0.1) +
  theme_classic() + 
  xlim(9660, 9740) +

  labs(title="Illustration of BIC distributions \n in low dimensional model spaces",
       caption = "Illustration of 160,000 BIC values from sampled models in a 149-variable space.") + 
  theme(
    axis.ticks = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    plot.title = element_text(hjust = 0.5, # Center the title
                                  color = "black",
                                  size = 16,
                                  face = "bold"),
    plot.caption = element_text(size = 7.5, colour = "grey30", hjust = 0),
    plot.caption.position = "plot" )    # puts caption below panel
```











```{r}
library(ggplot2)

make_bic_panel <- function(x, title, fill_col, bw = NULL,
                           bic_min = NULL, delta2 = 2,
                           xlim = range(x, na.rm = TRUE)) {
  # clean
  x <- x[is.finite(x)]
  stopifnot(is.numeric(x), length(x) > 1)

  if (is.null(bic_min)) bic_min <- min(x, na.rm = TRUE)

  df <- data.frame(BIC = x)

  ggplot(df, aes(BIC)) +
    # ΔBIC band as annotate() to avoid “length 1” warnings
    annotate("rect",
             xmin = bic_min - delta2, xmax = bic_min,
             ymin = -Inf, ymax = Inf,
             fill = "grey90", alpha = 0.5) +
    geom_density(fill = fill_col, colour = "black",
                 bw = bw, na.rm = TRUE) +
    geom_vline(xintercept = bic_min, linetype = 2, linewidth = 0.4) +
    scale_x_continuous(limits = xlim, expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), breaks = NULL) +
    labs(title = title, x = "BIC", y = NULL) +
    theme_classic() +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          plot.title = element_text(face = "bold", size = 14, hjust = 0.5))
}


# Example usage (replace x_easy/x_hard with your two BIC vectors or the same vector with different bw)
x <- unlist(S_list_MH)                     # your 160k BICs
#rng <- quantile(X, c(.01, .99), na.rm = TRUE)  # common x-range

rng <- c(9650, 9740)
p_easy <- make_bic_panel(x, "…low-dimensional feel", "#cfeecf",
                         bw = 4, xlim = rng)
p_hard <- make_bic_panel(x, "…high-dimensional feel", "#f6c7c7",
                         bw = 0.8, xlim = rng)


cowplot::plot_grid(p_easy, p_hard, ncol = 1, rel_heights = c(1,1))

```




```{r}
plot(X[,'geopotential_at_200_hPa'], X[,'geopotential_at_500_hPa'])
```


# Check mode


```{r}
library(dplyr)
library(tibble)
library(purrr)

# --- helpers ---------------------------------------------------------------
check_mode = function(){
      # turn a 0/1 vector into a compact key (for counting)
    .bin_key <- function(v) paste0(v, collapse = "")
    
    # compute Hamming distance between two 0/1 model vectors
    ham_dist <- function(a, b){
      cat(sprintf("calculating hamming distance of a,b with lengths %d and %d\n",length(a),length(b)))
      dist = sum(a != b)
      cat(sprintf("done calculating hamming distance\n"))
      dist
    }
    
    # return the mode model of a chain (post burn-in)
    mode_model <- function(chain, burn = 1000) {
      #S   <- chain$rs$samples[(burn+1):nrow(chain$rs$samples), , drop = FALSE]
      S   <- chain$rs$samples
      bic <- chain$rs$bic_trace[(burn+1):length(chain$rs$bic_trace)]  # aligns to S rows
      cat(sprintf("nrow(S)=%d, length(bic)=%d\n", nrow(S), length(bic)))
    
      keys <- apply(S, 1, .bin_key)
      tab  <- sort(table(keys), decreasing = TRUE)
      topk <- names(tab)[1]
      idx  <- which(keys == topk)
    
      v_mode <- as.integer(strsplit(topk, "")[[1]])
    
      list(
        key        = topk,
        v          = v_mode,
        count      = as.integer(tab[1]),
        prop       = as.integer(tab[1]) / nrow(S),
        bic_min    = min(bic[idx]),
        bic_median = median(bic[idx]),
        idx        = idx
      )
    }
    
    # convenience: summarize a chain against the global best
    summarize_chain_mode <- function(chain, chain_id, best) {
      mm <- mode_model(chain)
      tibble(
        chain   = chain_id,
        visits  = mm$count,
        prop    = mm$prop,
        bic_min = mm$bic_min,
        dBIC    = mm$bic_min - best$BIC,
        hamming_to_global_best = ham_dist(mm$v, colnames(X) %in% best$v),
        key     = mm$key
      )
    }
    
    # --- data you already have -------------------------------------------------
    # mh_lambda_0.75_chain_1 .. _4
    # best_BIC_minimizer_all_lambda$best$v and $BIC
    
    chains_075 <- list(
      `1: full`   = mh_lambda_0.75_chain_1,
      `2: empty`  = mh_lambda_0.75_chain_2,
      `3: sparse` = mh_lambda_0.75_chain_3,
      `4: random` = mh_lambda_0.75_chain_4
    )
    
    # --- main summaries --------------------------------------------------------
    
    mode_summaries <- imap_dfr(chains_075, ~ summarize_chain_mode(.x, .y, best_BIC_minimizer_all_lambda$best))
    
    mode_summaries
    # Columns:
    # chain | visits | prop | bic_min | dBIC | hamming_to_global_best | key
    
    # do all chains share the same modal model?
    all_same_mode <- n_distinct(mode_summaries$key) == 1
    all_same_mode
    
    # if not, show which chains agree
    split(mode_summaries$chain, mode_summaries$key)
    
    # --- optional: top-K modes per chain --------------------------------------
    
    top_k_modes <- function(chain, K = 5, burn = 1000) {
      #S   <- chain$rs$samples[(burn+1):nrow(chain$rs$samples), , drop = FALSE]
      S   <- chain$rs$samples
      bic <- chain$rs$bic_trace[(burn+1):length(chain$rs$bic_trace)]
      cat(sprintf("nrow(S)=%d, length(bic)=%d\n", nrow(S), length(bic)))
      keys <- apply(S, 1, .bin_key)
      tab  <- sort(table(keys), decreasing = TRUE)
      out_keys <- names(tab)[seq_len(min(K, length(tab)))]
      map_dfr(out_keys, function(k) {
        idx  <- which(keys == k)
        v    <- as.integer(strsplit(k, "")[[1]])
        tibble(
          key = k,
          visits = as.integer(tab[k]),
          prop = as.integer(tab[k]) / nrow(S),
          bic_min = min(bic[idx]),
          bic_median = median(bic[idx]),
          hamming_to_global_best = ham_dist(v,
                                            colnames(X) %in% best_BIC_minimizer_all_lambda$best$v)
        )
      })
    }
    
    top5_by_chain <- imap(chains_075, ~ top_k_modes(.x, K = 5) %>% mutate(chain = .y)) %>% bind_rows()
    top5_by_chain %>% arrange(chain, desc(prop))
}

check_mode_0.75 = check_mode()
check_mode_0.75
```
```{r}
check_mode_0.75 %>% filter(chain == "3: sparse")
```

## IT TURNS OUT SECOND MOST FREQUENT MODEL EXACTLY COINCIDES WITH GLOBAL MINIMIZER!!

```{r}

check_mode_0.75 %>%
  filter(chain == "3: sparse") %>%
  select(visits, prop, bic_min, hamming_to_global_best, chain)

```
```{r}
xtable(
  check_mode_0.75 %>%
  filter(chain == "3: sparse") %>%
  select(visits, prop, bic_min, hamming_to_global_best, chain)
)
```



IT TURNS OUT THE SECOND MOST FREQUENT MODEL MATCHES THE BIC MINMIZER!!!!

```{r}
setdiff(varnames_elites[[1]],
          colnames(X)[sapply(str_split((check_mode_0.75 %>% filter(chain == "3: sparse"))$key[2], "")[[1]], function(x)as.logical(as.integer(x)))])
```
```{r}
setdiff(colnames(X)[sapply(str_split((check_mode_0.75 %>% filter(chain == "3: sparse"))$key[2], "")[[1]], function(x)as.logical(as.integer(x)))],varnames_elites[[1]]
          )
```



