---
title: "analyze_results_n_10000_p_160"
author: "Keith Howen"
date: "2025-08-28"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))

#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("cowplot")
#library(cowplot)

#install.packages("mvnTest")

#library(splitstackshape)

#install.packages("glmnet")
library(glmnet)

```

# Read stuff

```{r}
list.files(here("output", "simulation_factorial_analysis_n_10000_p_160_outputs"))
```


## Read synthetic datasets & mcmc samples
```{r}

res = readRDS(here("output", "simulation_factorial_analysis_n_10000_p_160_outputs", "results_n_10000_p_160.rds"))[["0.9"]][["moderate"]]

X = res$X.syn
y = res$y.syn


mcmc.samples = readRDS(here("output", "simulation_factorial_analysis_n_10000_p_160_outputs", "gibbs_samples_n_10000_p_160.rds"))

```


# Plot

## rel freq with ESS-adjusted CI
```{r}
inc_df <- inclusion_summary(mcmc.samples, top_k = 12, gt_idx = 1:4, ensure_gt = TRUE)
plot_inclusion(inc_df,
               title="Inclusion probabilities with ESS 95% CI\n(Highly Correlated GTs, Moderate SNR)")
```


## Trace of number of variables
```{r}
# high corr case number of variables in every sampled model
plot_num_vars_trace(mcmc.samples,
                    window = 50,
                    title = "Trace: number of variables in every sampled model",
                    dimension_BIC_minim = 3)
```
## Heatmap of model path

```{r}
plot_model_heatmap(mcmc.samples,
                   thin = 10,
                   order = "by_mean",
                   exclude_unimportant_vars=152,
                   put_gts_on_top=FALSE,
                   title = "Model path heatmap (highly correlated GTs, moderate SNR)")
```


# LASSO comparison
```{r}
lasso.variable.selection = function(X, y,
                                    how='lambda.1se'){
  
  # default is LASSO regression (alpha=1), so we good here
  lasso.fit = glmnet(X, y, family = "binomial")
  
  # note best lambda is not chosen yet -- one way is via cross validation.
  cvfit = cv.glmnet(X, y)
  
  # now, get model coefficients at that value of lambda
  coefs.best.lambda = coef(cvfit, s=how)
  
  # extract non-zero coefficients
  
  vars.non.zero.coefs = names(
      coefs.best.lambda[,how][
        coefs.best.lambda[,how] != 0
        ]
    )
    

  coefs.best.lambda.nonzero =
    coefs.best.lambda[vars.non.zero.coefs, ]
  
  # return everything!
  return(
    list(
      lasso.fit=lasso.fit,
      cvfit=cvfit,
      coefs.best.lambda.nonzero=coefs.best.lambda.nonzero
    )
  )
  
}
```



```{r}
lasso = lasso.variable.selection(X, y)
lasso$coefs.best.lambda.nonzero
```


```{r}
lasso.variable.selection(X, y, how='lambda.min')$coefs.best.lambda.nonzero
```



```{r}
dim(X)


summary(
  glm(y ~ X1+X2+X3+X4, data=data.frame(X,y), family='binomial')
)

anova(glm(y ~ X1+X3+X4, data=data.frame(X,y), family='binomial'),
      glm(y ~ X1+X2+X3+X4, data=data.frame(X,y), family='binomial'))


```





