---
title: "TC_sim_factorial_analysis_RS_1_results"
author: "Keith Howen"
date: "2025-07-15"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}

source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(tidyverse)
library(here)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")
#install.packages("tidyverse")

#library(splitstackshape)
```


# Path for results
```{r}
BASE_DIR = "simulation_factorial_analysis_n_2000_p_16_outputs"
list.files(
  here("output", BASE_DIR)
)
```

# Read MCMC samples for 15 combs

The 3 design matrices for each of the 5 correlation level are the SAME.
⦁	For example, the X matrix for c1, c6, and c11 are the same
✅Combination 1/c1: corr(X1,X2)=0, high s/n
✅Combination 2/c2: corr(X1,X2)=+0.5, high s/n
✅Combination 3/c3: corr(X1,X2)=+0.9, high s/n
✅Combination 4/c4: corr(X1,X2)=-0.5, high s/n
✅Combination 5/c5: corr(X1,X2)=-0.9, high s/n
✅Combination 6/c6: corr(X1,X2)=0, moderate s/n
✅Combination 7/c7: corr(X1,X2)=+0.5, moderate s/n
✅Combination 8/c8: corr(X1,X2)=+0.9, moderate s/n
✅Combination 9/c9: corr(X1,X2)=-0.5, moderate s/n
✅Combination 10/c10: corr(X1,X2)=-0.9, moderate s/n
✅Combination 11/c11: corr(X1,X2)=0, weak s/n
✅Combination 12/c12: corr(X1,X2)=+0.5, weak s/n
✅Combination 13/c13: corr(X1,X2)=+0.9, weak s/n
✅Combination 14/c14: corr(X1,X2)=-0.5, weak s/n
✅Combination 15/c15: corr(X1,X2)=-0.9, weak s/n

Results are prefixed by `gibbs`. 

```{r}
sort(
  grep("gibbs",
     list.files(here("output", BASE_DIR)),
     value=TRUE)
)
```

```{r}

CORR_LEVELS = c("0", "+0.5", "+0.9", "-0.5", "-0.9")
SN_LEVELS = c("strong", "moderate", "weak")

meta <- tibble(
    combination = 1:15,
    correlation = rep(CORR_LEVELS, length(SN_LEVELS)),
    s_n = rep(SN_LEVELS, each = length(CORR_LEVELS))
  )


read_all_MCMC_samples = function(){
  
  ## 1. INITS 
  # Metadata for combinations
  meta <- tibble(
    combination = 1:15,
    correlation = rep(CORR_LEVELS, length(SN_LEVELS)),
    s_n = rep(SN_LEVELS, each = length(CORR_LEVELS))
  )
  
  # dict with 2 keys essentially
  mcmc.samples <- list()
  for (rho in CORR_LEVELS) {
    mcmc.samples[[rho]] <- list()
    for (s in SN_LEVELS) {
      mcmc.samples[[rho]][[s]] <- "empty"  
    }
  }
  
  ## 2. store MCMC samples
  for (i in 1:15) {
    
    combo_id <- paste0("c", i)
    path_base <- here("output", BASE_DIR)
    
    row = readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".rds")))

    # retrieve corresponding (corr_level, sn_level) from
    # the 'meta' metadata -- have to use redirection like operator
    # to filter tibbles
    corr_level = (meta %>% filter(combination == i))$correlation
    sn_level = (meta %>% filter(combination == i))$s_n
    
    mcmc.samples[[corr_level]][[sn_level]] = row
  }
  
  return(mcmc.samples)
  
}



# Final summary table
mcmc.samples = read_all_MCMC_samples()
```


# Start Analyzing

```{r}
names(mcmc.samples)

dim(mcmc.samples[["0"]][["strong"]])
```

## 🖼️ Heatmap

```{r}

plot_grid(
  
  
  plot_model_heatmap(mcmc.samples[["+0.9"]][["moderate"]],
                     thin = 5,
                     order = "by_mean",
                     exclude_unimportant_vars=12,
                     put_gts_on_top=FALSE,
                     transpose=FALSE,
                     title = "Model path heatmap (highly correlated GTs, moderate SNR)"),
    
  plot_model_heatmap(mcmc.samples[["+0.9"]][["weak"]],
                     thin = 5,
                     order = "by_mean",
                     exclude_unimportant_vars=12,
                     put_gts_on_top=FALSE,
                     transpose=FALSE,
                     title = "Model path heatmap (highly correlated GTs, weak SNR)"), 
    
    
    nrow=2
  
)

```



## 🖼️ Trace plot of number of variables

```{r}


plot_grid(
  

  plot_num_vars_trace(mcmc.samples[["+0.9"]][["moderate"]],
                    window = 10,
                    dimension_BIC_minim = 4,
                    title = "Trace: number of variables \n in sampled models \n (highly correlated GTs, \n moderate SNR)"), 
    
  plot_num_vars_trace(mcmc.samples[["+0.9"]][["weak"]],
                    window = 10,
                    dimension_BIC_minim = 2,
                    title = "Trace: number of variables \n in sampled models \n (highly correlated GTs, \n weak SNR)"), 
    
    
  ncol=2
  
)


```
```{r}
plot_num_vars_trace(mcmc.samples[["+0.9"]][["weak"]],
                    window = 10,
                    dimension_BIC_minim = 4)
```

## Rel freqs with ESS-adjusted CI
```{r}

plot_grid(
  
  plot_inclusion(inclusion_summary(mcmc.samples[["+0.9"]][["moderate"]],
                            top_k = 16,
                            ensure_gt = TRUE,
                            gt_idx=1:4),
                 title = "Inclusion probabilities\nwith ESS 95% CI\n(Highly correlated GTs, moderate SNR)"),
  
  plot_inclusion(inclusion_summary(mcmc.samples[["+0.9"]][["weak"]],
                            top_k = 16,
                            ensure_gt = TRUE,
                            gt_idx=1:4),
                 title = "Inclusion probabilities\nwith ESS 95% CI\n(Highly correlated GTs, weak SNR)"),
  ncol=2
  
  
)





```



# BIC Minimizers
(moderate, +0.5)
and all levels of weak

```{r}
print(meta)
```


```{r}
investigate_imperfect_combos = function(){
  
  
  # manually enumerate scnearios where correct model is not obtained
  
  imperfect_combs_id = c(7,11,12,13,14,15)
  
  path_base <- here("output", BASE_DIR)
  
  for (i in imperfect_combs_id){
    
    corr_level = (meta %>% filter(combination == i))$correlation
    sn_level = (meta %>% filter(combination == i))$s_n
    
    cat("Top 5 BIC Minimizers for ", corr_level, ",",  sn_level, "\n")
    
    data_comb_i = readRDS(file.path(path_base, paste0("result.","c",i,".rds")))
    
    X = data_comb_i$X.syn
    y = data_comb_i$y.syn
    
    print(
      find_BIC_minimizer(data.frame(X,y),mcmc.samples[[corr_level]][[sn_level]])$models_by_bic
    )
  }
  
}


investigate_imperfect_combos()


```

# Obtain BIC of true models for imperfect results
✅Combination 1/c1: corr(X1,X2)=0, high s/n
✅Combination 2/c2: corr(X1,X2)=+0.5, high s/n
✅Combination 3/c3: corr(X1,X2)=+0.9, high s/n
✅Combination 4/c4: corr(X1,X2)=-0.5, high s/n
✅Combination 5/c5: corr(X1,X2)=-0.9, high s/n
✅Combination 6/c6: corr(X1,X2)=0, moderate s/n
✅Combination 7/c7: corr(X1,X2)=+0.5, moderate s/n
✅Combination 8/c8: corr(X1,X2)=+0.9, moderate s/n
✅Combination 9/c9: corr(X1,X2)=-0.5, moderate s/n
✅Combination 10/c10: corr(X1,X2)=-0.9, moderate s/n
✅Combination 11/c11: corr(X1,X2)=0, weak s/n
✅Combination 12/c12: corr(X1,X2)=+0.5, weak s/n
✅Combination 13/c13: corr(X1,X2)=+0.9, weak s/n
✅Combination 14/c14: corr(X1,X2)=-0.5, weak s/n
✅Combination 15/c15: corr(X1,X2)=-0.9, weak s/n
```{r}

# 11, 14
compute.BIC.true.model = function(i){
  
    data_comb_i = readRDS(file.path(path_base, paste0("result.","c",i,".rds")))
    X = data_comb_i$X.syn
    y = data_comb_i$y.syn
    
    return(
      compute.BIC(data.frame(X,y), colnames(X) %in% c("X1","X2","X3","X4"))
    )
  
}

compute.BIC.true.model(14)

```




# 🧹Summarize results in a table

## raw code

```{r}


# Metadata for combinations
meta <- tibble(
  combination = 1:15,
  correlation = rep(c("0", "+0.5", "+0.9", "-0.5", "-0.9"), 3),
  s_n = rep(c("high", "moderate", "weak"), each = 5)
)

# Ground truth variables
ground_truth_vars <- paste0("X", 1:4)

# Initialize empty list to collect rows
all_results <- vector("list", length = 15)

for (i in 1:15) {
  combo_id <- paste0("c", i)
  path_base <- here("output", "simulation_factorial_analysis_outputs")

  # Read relative frequencies
  rel_freq_obj <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".rel.freqs.rds")))
  rel_freqs <- rel_freq_obj$rel.freqs.non.basin

  # Read empirical mode (most common model in MCMC)
  empirical_mode <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".best.Xs.rds")))

  # Read BIC-selected model
  bic_model_obj <- readRDS(file.path(path_base, paste0("gibbs.RS.", combo_id, ".best.Xs.BIC.rds")))
  bic_model <- names(which(bic_model_obj$best.model == 1))
  min_bic <- bic_model_obj$min.BIC

  # Logical check: were all GT variables recovered?
  recovered_all_gt_emp <- all(ground_truth_vars %in% empirical_mode)
  recovered_all_gt_bic <- all(ground_truth_vars %in% bic_model)

  # Combine into one row (as tibble)
  row <- tibble(
    Combination = i,
    Correlation = meta$correlation[i],
    Signal_to_Noise = meta$s_n[i],
    Empirical_Mode = paste(empirical_mode, collapse = ", "),
    BIC_Model = paste(bic_model, collapse = ", "),
    Min_BIC = min_bic,
    Recovered_GT_Empirical = recovered_all_gt_emp,
    Recovered_GT_BIC = recovered_all_gt_bic
  ) %>%
    bind_cols(as_tibble(t(rel_freqs)))  # Add relative freq columns

  all_results[[i]] <- row
}

# Final summary table
summary_table <- bind_rows(all_results)

# Show in viewer
print(summary_table)

```



```{r}
SUMMARY_TABLE_COLS_TO_EXCLUDE = c(sapply(5:16, function(i) paste0("X",i)), "Combination")
summary_table_filtered = summary_table[,!colnames(summary_table) %in% SUMMARY_TABLE_COLS_TO_EXCLUDE]
```


```{r}
summary_table_filtered
```


## package it into a function

```{r}

```



# 🧹Automatically generate latex code from result!.

```{r}

# install.packages(c("dplyr","knitr","kableExtra"))
# a little helper to turn one SNR‐subset into a LaTeX table
make_snr_table <- function(df, snr_level) {
  sub <- df %>%
    filter(Signal_to_Noise == snr_level) %>%
    select(
      Correlation,
      `v1 freq` = X1,
      `v2 freq` = X2,
      `v3 freq` = X3,
      `v4 freq` = X4,
      `Empirical mode`    = Empirical_Mode,
      `BIC model`         = BIC_Model,
      `Min. BIC`          = Min_BIC
    )
  
  cap <- paste0("Summary statistics for \\textbf{", snr_level,
                "} signal‐to‐noise across correlation levels")
  
  sub %>%
    kable(
      format       = "latex",
      booktabs     = TRUE,
      caption      = cap,
      label        = paste0("tab:", snr_level),
      align        = c("l", rep("c", ncol(sub)-1)),
      escape       = FALSE
    ) %>%
    kable_styling(latex_options = c("hold_position","scale_down")) %>%
    column_spec(1, bold = TRUE)
}

# generate result tables for latex report

make_sim_study_tables = function(summary_table_filtered){
  # now loop over each SNR level and cat() the tables
  snrs <- unique(summary_table_filtered$Signal_to_Noise)
  for (snr in snrs) {
    cat("\n\n", make_snr_table(summary_table_filtered, snr), "\n\n")
  }
}

make_sim_study_tables(summary_table_filtered)



```





