---
title: "TC_sim_factorial_analysis_generate_data"
author: "Keith Howen"
date: "2025-05-29"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))

source(here("lib", "probit_gibbs.R"))
source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
#library(splitstackshape)
```

## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
```

# Import sampled data

```{r}

X.sim.balanced.class <- readRDS(here("output", "simulation_v2_outputs", "X.sim.balanced.class.rds"))

covariates.mwr.final.mod <- readRDS(here("output", "simulation_v2_outputs", "covariates.mwr.final.mod.rds"))
COVARIATES.TABLE4 = readRDS(here("output", "simulation_v2_outputs", "COVARIATES.TABLE4.rds"))

y.sim <- readRDS(here("output", "simulation_v2_outputs", "y.sim.rds"))

X.sim.table4.vars = X.sim.balanced.class[,COVARIATES.TABLE4]

```

# Fit 2000 rows first

# 📕 Propose algorithm to tweak signal given correlation

*Ground truth* covariates: covariates that were involved in generating
the response.

Obtain a synthetic dataset comprised of ground truth covariates (+
redundant covariates) and synthetic responses in a way that **ALL ground
truth covariates are significant**.

Steps (for a **fixed** correlation structure between $X_1, \cdots, X_4$:

0.  Set $t=0$, where $y^{(0)}$ corresponds to binary responses either
    from the TC genesis data or used to fit a model to
    $X_1, \cdots, X_4$.

    For $t \ge 1$,

1.  Fit a a model to the 2000 observations of $X_1, \cdots, X_4$ with
    the responses $y^{(t-1)}$. Obtain estimates
    $\hat{\beta_1}^{(t-1)},\hat{\beta_2}^{(t-1)},\hat{\beta_3}^{(t-1)},\hat{\beta_4}^{(t-1)}$
    .

2.  If any $\hat{\beta_i}^{(t-1)}$ not significant, scale
    $\hat{\beta_i}^{(t)}=c_i \cdot \hat{\beta_i}^{(t-1)},\ |c_i|>1$ to
    **increase their magnitude (hoping to make them significant
    enough)**.

    -   Otherwise stop and return
        $\hat{\beta_1}^{(t-1)},\hat{\beta_2}^{(t-1)},\hat{\beta_3}^{(t-1)},\hat{\beta_4}^{(t-1)}$
        as well as $y^{(t-1)}$.

3.  Generate another set of responses $y^{(t)}$ by
    $Ber\left(\sigma(\sum_{i=1}^4\hat{\beta_i}^{(t)}X_i)\right)$

4.  Loop to next iteration.

Then include the rest of 12 variables as candidate variables in the
variable selection process.

# Obtain the original responses of the 2000 covariates

```{r}
table(data.NA.fixed[as.integer(rownames(X.sim.balanced.class)), ]$TC_genesis)
```

```{r}
y.balanced.class = data.NA.fixed[as.integer(rownames(X.sim.balanced.class)), ]$TC_genesis
```

## Examine correlation structure

Examine pairwise correlation coefficient in ground truth variables on
the 2000 sampled covariates.

```{r}
cor(X.sim.balanced.class[,covariates.mwr.final.mod])[1:4,1:4]
```

## Reflection on previous wrong methodology

Recall `potential_vorticity_at_600_hPa` and
`vertical_wind_shear_between_200_hPa_and_700_hPa` are picked from the RS
but not the other 2.

Recall the significance of the ground truth variables w.r.t to the
**simulated response**

```{r}
summary(
  glm(y.sim ~ 0 + basinEP + basinNA + basinNI + basinSA +
        basinSI + basinSP + basinWP + Coriolis_parameter + potential_vorticity_at_600_hPa +
        vertical_wind_shear_between_200_hPa_and_700_hPa + vorticity_relative_at_925_hPa,
    family=binomial,
    data=data.frame(cbind(X.sim.table4.vars, y.sim)))
)
```

`Coriolis_parameter` is relatively significant but does not get picked
-- since it is highly correlated with `potential_vorticity_at_600_hPa`.

Possible explanation: Once `potential_vorticity_at_600_hPa` is in the
model, adding `Coriolis_parameter` won’t substantially improve the fit.

But in terms of `vorticity_relative_at_925_hPa` it is *just* because its
not significant.

Hence why the 2 factors determining variable selection performance is
collinearity AND significance.

# PREFACE ON SYNTHETIC COVARIATE GENERATION

Since we are not inferring the distirbution of covariates (in linear
models they are assumed to be fixed), we can use any method to generate
covariates based on their correlation.

-   After all, the thing that impacts variable selection is their
    RELATIONSHIP to the response as well as their CORRELATION (since we
    are using a linear model)

# ✅Comb 1: rho=0, high s/n

## Generate uncorrelated X1,..,X4 covariates with 12 redundants

```{r}
#mod.1 = glm(y.balanced.class ~ 0 + basinEP + basinNA + basinNI + basinSA +
#        basinSI + basinSP + basinWP + Coriolis_parameter + potential_vorticity_at_600_hPa +
#        vertical_wind_shear_between_200_hPa_and_700_hPa + vorticity_relative_at_925_hPa,
#    family=binomial,
#    data=data.frame(cbind(X.sim.balanced.class, y.balanced.class)))

#summary(mod.1)
```

Idea: use 1 real covariate realization, and from there generate 3
covariates such that they are mutually uncorrelated.

-   NOPE. Its easier to generate synthetic covariates entirely to ensure
    any form of mutual uncorrelation.

### ✏️ Helper function

NOTE: its on the `factorial_analysis_routines` module.

```{r}

set.seed(123)

#resid_out <- function(x, ref) {
#  return(lm(x ~ ref)$residuals)
#}

#gen_3_uncorr_vars = function(X_real) {
#  
#  SIZE = length(X_real)
#  
#  X_syn1 <- resid_out(rnorm(SIZE), X_real)
#  X_syn2 <- resid_out(rnorm(SIZE), X_real)
#  X_syn3 <- resid_out(rnorm(SIZE), X_real)
#  
#  # bunch them up into a n x 4 matrix
#  mat = cbind(X_real, X_syn1, X_syn2, X_syn3)
#  colnames(mat) = c("X1", "X2", "X3", "X4")
#  
#  return(mat)
#}


# generates MVNorm synthetic variables
# given a covariance (or correlation) matrix r
# NOTE: by convention, X1 to X4 are ground truths
gen_synthetic_vars = function(r, SIZE=2000){
  

  # generates SIZE random vectors of dimension nrow(r) or ncol(r)
  syn.vars <- mvrnorm(n=SIZE,
                mu=rep(0, nrow(r)),
                Sigma=r,
                empirical=TRUE )
  
  # include an intercept column?
  
  # fix column names ("X0" = intercept)
  varnames = 1:nrow(r)
  varnames = sapply(varnames, FUN=function(i){paste("X",i,sep='')})
  
  
  colnames(syn.vars) = varnames
  
  return(syn.vars)

}



```

```{r}

set.seed(123)

NUM_GROUND_TRUTHS = 4
NUM_REDUNDANT = 12

r.c1 = diag(NUM_GROUND_TRUTHS + NUM_REDUNDANT)

X.c1 = gen_synthetic_vars(r.c1)

round(cor(X.c1)[1:NUM_GROUND_TRUTHS, 1:NUM_GROUND_TRUTHS], 10)

```

## Start tweaking beta values

We might automate it in a function

```{r}
set.seed(123)
# GOAL: generate dataset for combination 1 (c1)

# 'guess' a beta value
# perhaps have some variables with a mix of + and - strong signal?
beta.c1 = c(3, 3.5, -3, -4.5)

# generate responses based on the beta coefficients (all strong)
X.true.c1 = X.c1[,1:NUM_GROUND_TRUTHS]
eta.c1 = X.true.c1 %*% beta.c1
p.c1 = 1 / (1 + exp(-eta.c1))
y.c1 = rbinom(nrow(X.c1), size=1, prob=p.c1)


# see how significant each variable is
mod.c1 = glm(formula= y.c1 ~ .,
    family='binomial',
    data=data.frame(X.true.c1, y.c1))

summary(mod.c1)

```

```{r}
table(y.c1)
```



All the ground truth variables involved in generating the response are
**highly significant**.

Misc remarks - Estimates are fairly close, but not that close. This is
most likely due to the inherent noise present in the fairly modest
sample size, though I think 2000 is good enough for stable coefficient
estimation.

## Saving the dataset for combination 1

```{r}

# save the true beta coeffs used to generate the responses
saveRDS(beta.c1,
        here("output",
             "simulation_factorial_analysis_outputs",
             "beta.c1.rds"))

# save the synthetic response
saveRDS(y.c1,
        here("output",
             "simulation_factorial_analysis_outputs",
             "y.c1.rds"))


# save design matrix of candidate variables
saveRDS(X.c1,
        here("output",
             "simulation_factorial_analysis_outputs",
             "X.c1.rds"))


# save the true probabilities 
saveRDS(p.c1,
        here("output",
             "simulation_factorial_analysis_outputs",
             "p.c1.rds"))

```

# ✏️ Automate significance checking in a function

See `lib/factorial_analysis_routines.R`.

# ✅Comb 2: rho=+0.5, high s/n

4 ground truth variables, 2 of which are moderately correlated (ρ =
0.5), and 12 redundant variables that are uncorrelated with all ground
truths and each other.

## 💡 TIP: construct block diagonal matrices using `Matrix.bdiag()`

NOTE: you should have the `Matrix` library installed and imported first.

## 💡 Math'al setup

$$
\text{Assume we have 4 ground truths and 12 redundant variables} \\
\text{(redundant meaning NOT involved in generating the response)} \\
\text{Let }\mathbf{x_{g1}}= \begin{bmatrix} X_1 \\ X_2  \end{bmatrix},
\text{ ground truths with controlled correlation} \\
\text{Let }\mathbf{x_{g2}}= \begin{bmatrix} X_3 \\ X_4  \end{bmatrix},
\text{ uncorrelated ground truths}\\
\text{Let }\mathbf{x_{r}}= \begin{bmatrix} X_5 \\ \cdots \\ X_{16}  \end{bmatrix},
\text{ the redundant variables} \\
\Sigma = \begin{bmatrix}
\mathbf{Var(x_{g1})} = \begin{bmatrix} 1 && \rho \\ \rho && 1  \end{bmatrix}  && \mathbf{O} && \mathbf{O} \\ 
\mathbf{O} && \mathbf{Var(x_{g1})}=I_2 && \mathbf{O} \\
\mathbf{O} && \mathbf{O} && \mathbf{Var(x_{r})}=I_{12} \\
\end{bmatrix} \\
\text{NOTE: for now set variables to have unit variance for simplicity,} \\ 
\text{and set the redundant variables to be uncorrelated with e.o. and the ground truths}
$$

## Generate the data

```{r}


r.c2 = as.matrix(
  bdiag( # Ensuring that (X1,X2), (X3,X4) and (X5,..,X16) are mutually uncorrelated.
    matrix(c(1, 0.5, 0.5, 1),c(2,2)), # Block 1: 2 correlated ground truths (ρ = 0.5)
    diag(2), # Block 2: 2 uncorrelated ground truths
    diag(12) # Block 3: 12 uncorrelated redundant variables
  )
)

X.c2 = generate_synthetic_vars(r.c2,SIZE=2000)

head(X.c2)

round(
  cor(X.c2),
  10
)
```

## Start tweaking beta values

```{r}

result.c2 = check_significance(c(3, 3.5, -3, -4.5), X.c2)

```

```{r}

cat("Distribution of synthetic response below:")
print(table(result.c2$y.syn))

#cat("\nCoefficient estimates, p-values, and significance of each ground truth variable below:")
#print(result.c2$sig)

#hist(
#  1 / (1 + exp(-result.c2$X.syn[,1:4] %*% result.c2$beta))
#)

```

## Saving the results for combination 2

```{r}
names(result.c2)
```



```{r}
# save the results for simulation scenario 2 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c2,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c2.rds"))
```



# ✅ Comb 3: rho=+0.9, high s/n

4 ground truth variables, 2 of which are highly correlated (ρ = 0.9),
and 12 redundant variables that are uncorrelated with all ground truths
and each other.

## Generate the data

```{r}
r.c3 = as.matrix(
  bdiag( # Ensuring that (X1,X2), (X3,X4) and (X5,..,X16) are mutually uncorrelated.
    matrix(c(1, 0.9, 0.9, 1),c(2,2)), # Block 1: 2 correlated ground truths (ρ = 0.9)
    diag(2), # Block 2: 2 uncorrelated ground truths
    diag(12) # Block 3: 12 uncorrelated redundant variables
  )
)

X.c3 = generate_synthetic_vars(r.c3,SIZE=2000)

head(X.c3)

round(
  cor(X.c3),
  10
)

```

## Start tweaking beta values

```{r}
result.c3 = check_significance(c(3, 3.5, -3, -4.5), X.c3)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c3$y.syn))
```


Standard errors for X1 and X2 are larger (presumably due to the strong
correlation between X1 and X2), and overall z-statistics decrease
(though given the effect sizes, these are considered large since the
p-values are still very small)

## Saving the results

```{r}
# save the results for simulation scenario 3 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c3,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c3.rds"))
```

# ✅Comb 4: rho=-0.5, high s/n

## Generate the data

```{r}
r.c4 = as.matrix(
  bdiag( # Ensuring that (X1,X2), (X3,X4) and (X5,..,X16) are mutually uncorrelated.
    matrix(c(1, -0.5, -0.5, 1),c(2,2)), # Block 1: 2 correlated ground truths (ρ = -0.5)
    diag(2), # Block 2: 2 uncorrelated ground truths
    diag(12) # Block 3: 12 uncorrelated redundant variables
  )
)

X.c4 = generate_synthetic_vars(r.c4,SIZE=2000)

head(X.c4)

round(
  cor(X.c4),
  10
)

```

## Start tweaking beta values

(3,3.5,-3,-4.5) should be more than sufficient for a strong s/n, despite
the -0.5 correlation coeff between X1 and X2

```{r}
result.c4 = check_significance(c(3, 3.5, -3, -4.5), X.c4)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c4$y.syn))
```



## Saving the results

```{r}
# save the results for simulation scenario 3 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c4,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c4.rds"))
```



# ✅Comb 5: rho=-0.9, high s/n

## Generate the data

```{r}
r.c5 = as.matrix(
  bdiag( # Ensuring that (X1,X2), (X3,X4) and (X5,..,X16) are mutually uncorrelated.
    matrix(c(1, -0.9, -0.9, 1),c(2,2)), # Block 1: 2 correlated ground truths (ρ = -0.9)
    diag(2), # Block 2: 2 uncorrelated ground truths
    diag(12) # Block 3: 12 uncorrelated redundant variables
  )
)

X.c5 = generate_synthetic_vars(r.c5,SIZE=2000)

head(X.c5)

round(
  cor(X.c5),
  10
)

```

## Start tweaking beta values

(3,3.5,-3,-4.5) should be more than sufficient for a strong s/n, despite
the -0.9 correlation coeff between X1 and X2

```{r}
result.c5 = check_significance(c(3, 3.5, -3, -4.5), X.c5)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c5$y.syn))
```



```{r}
# save the results for simulation scenario 3 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c5,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c5.rds"))
```


# 💡 DEFINING STRENGTH OF s/n

(From the meeting) You have several choices to decide the extent of s/n:

1.  Moderate

-   Reduce **all 4** coefficients to have moderate signal, or
-   Make 2 strong, 2 moderate (⚠️only problem for this is the
    overshadowing phenomenon present in the previous incorrect sim
    study-- or is this expected?)

2.  Weak

-   Do **NOT** reduce **all** of them to be weak
-   Maybe make 1 or 2 of them to be weak.

How about we define a rule of thumb: - Strong s/n = all are Strong -
Moderate s/n = 2 Moderate, other 2 is Strong - Weak s/n = 2 weak, other
2 is strong

Strong, moderate, and weak defined here:

$$
\text{Strong }\texttt{(***): }0 < \text{p-value} \le 0.001 \\
\text{Moderate }\texttt{(**): }0.001 < \text{p-value} \le 0.01 \\
\text{Weak }\texttt{(*): }0.01 < \text{p-value} \le 0.05
$$

# ✅ Comb 6: rho=0, moderate s/n

4 **uncorrelated** ground truth variables, and 12 redundant variables
that are uncorrelated with all ground truths and each other.

Also ONLY 2/4 ground truth variables have strong signals (i.e. not all
magnitude of $\beta_j$s are large to be statistically significant)

## Reuse Data From C1

We can use the same dataset (mutually uncorrelated) from combination 1.

```{r}
X.c6 <- X.c1
dim(X.c6)
round(cor(X.c6), 10)
```

## Start tweaking beta values


Last time (rho=0 and high s/n) we got $\beta = [3, 3.5, -3, -4.5]^T$

Perhaps we can reduce the magnitudes of 2 directly

$\beta = [3, \mathbf{1.5}, -3, \mathbf{-1.5}]^T$

After some trial and error, we had to reduce the magnitudes to **below
1**.

```{r}
result.c6 = check_significance(c(3, 0.35, -3, -0.18), X.c6)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c6$y.syn))
```


Here X1 and X3 strong, X2 and X4 moderate

```{r}
names(result.c6)
result.c6$sig


```

## Saving the results

```{r}
# save the results for simulation scenario 6 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c6,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c6.rds"))
```

# ✅ Comb 7: rho=+0.5, moderate s/n

## Reuse Data From C2

We can use the same design matrix X from Combination 2.

```{r}

X.c7 <- X.c2
dim(X.c7)
round(cor(X.c7), 10)

```

## Start Tweaking Beta Values

To reduce to moderate s/n (i.e. 2 strong sig, 2 moderate sig), - perhaps
we can start with the same $\beta$ from Combination 6 (rho=0, moderate
s/n), - that is $\beta = [3, 0.35, -3, -0.18]^T$

Initially X1,X2,X3 Strong but X4 is very weak (p-value is 0.45). So I
tried to increase the MAGNITUDE actual beta for X4 (-0.18 to -0.34),
whilst bringing down the acual beta for X2 (0.35 to 0.28).

```{r}
result.c7 = check_significance(c(3, 0.28, -3, -0.32), X.c7)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c7$y.syn))
```



## Saving the results

```{r}
# save the results for simulation scenario 6 (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c7,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c7.rds"))
```

# ✅ Comb 8: rho=+0.9, moderate s/n

## Reuse Data From C3

```{r}

X.c8 <- X.c3
dim(X.c8)
round(cor(X.c8), 10)

```

## Start tweaking beta values (from C7)

To reduce to moderate s/n (i.e. 2 strong sig, 2 moderate sig), - perhaps
we can start with the same $\beta$ from Combination 7 (rho=+0.5,
moderate s/n), - that is $\beta = [3, 0.28, -3, -0.32]^T$ -- here is the
first result

```{r}
result.c8.init = check_significance(c(3, 0.28, -3, -0.32), X.c8)
```

So for some reason I had to reduce the magnitude for X4 first (as
opposed to raising the magnitude for X2 first), and then raise the
magnitude (quite substantially, from 0.28 to 0.6) for X2.

(🤔 it would be much easier if we started from Combination 3 -- rho=0.9
with high s/n)

```{r}
result.c8 = check_significance(c(3, 0.6, -3, -0.29), X.c8)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c8$y.syn))
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c8,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c8.rds"))
```

# ✅Comb 9: rho= -0.5, moderate s/n


## Reuse dataset from C4
```{r}
X.c9 <- X.c4
dim(X.c9)
round(cor(X.c9), 10)
```
## Start tweaking beta values (from C4, C7)

Goal is to get 2 significant results, 2 moderate results

To reduce to moderate s/n (i.e. 2 strong sig, 2 moderate sig), - perhaps
we can start with the same $\beta$ from Combination 4 (rho=-0.5,
high s/n), - that is $\beta = [3, 3.5, -3, -4.5]^T$.

Reduce magnitude of X2, X4 for consistency.
- Actually, we can try to use the beta values from C7 -- exact same thing
except correlation coeff differ by a sign. Let's try (3, 0.28, -3, -0.32) first.

Yep, only had to reduce magnitude of X4 from -0.32 to -0.3.

```{r}
result.c9 = check_significance(c(3, 0.28, -3, -0.3), X.c9)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c9$y.syn))
```



## Saving the results.
```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c9,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c9.rds"))
```



# ✅Comb 10: rho=-0.9, moderate s/n

## Reuse dataset from C5

```{r}
X.c10 <- X.c5
dim(X.c10)
round(cor(X.c10), 10)
```

## Start tweaking beta values (from C8)

This is because dataset in C8 has rho=+0.9 with moderate s/n (only differ
by correlation sign)
We can use the resulting beta values from C8 as a starting point.
- I.e. start from (3, 0.6, -3, -0.29).

Only had to decrease magnitude of beta for X2 from 0.6 to 0.58.

```{r}
result.c10 = check_significance(c(3, 0.58, -3, -0.29), X.c10)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c10$y.syn))
```


## Saving the results.
```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c10,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c10.rds"))
```


# ✅Comb 11: rho=0, weak s/n

## Reuse Data From C1

We can use the same dataset (mutually uncorrelated) from combination 1.

```{r}
X.c11 <- X.c1
dim(X.c11)
round(cor(X.c11), 10)
```

## Start tweaking beta values (from C6)


Last time (rho=0 and moderate s/n) we got $\beta = [3, 0.35, -3, -0.18]^T$

Perhaps we can even further reduce the magnitude for X2 and X4.
Let's say beta for X2: 0.3 (from 0.35)
Let's say beta for X4: -0.15 (from -0.18)

```{r}
result.c11 = check_significance(c(3, 0.3, -3, -0.15), X.c11)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c11$y.syn))
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c11,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c11.rds"))
```



# ✅Comb 12: rho=+0.5, weak s/n

## Reuse data from C2

We can use the same dataset from combination 2.

```{r}
X.c12 <- X.c2
dim(X.c12)
round(cor(X.c12), 10)
```

## Start Tweaking Beta Values (from C7)

To reduce to weak s/n (i.e. 2 strong sig, 2 weak sig), - perhaps
we can start with the same $\beta$ from Combination 7 (rho=+0.5, moderate
s/n), - that is $\beta = [3, 0.28, -3, -0.32]^T$

We can reduce magnitude of X2 and X4 to yield weak statistical significance

Approach: decrease magnitude of X4 VERY SLIGHTLY then proceed with X2.
Apparently this works.


```{r}
result.c12 = check_significance(c(3, 0.23, -3, -0.31), X.c12)
```
```{r}
cat("Distribution of synthetic response below:")
print(table(result.c12$y.syn))
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c12,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c12.rds"))
```


# ✅Comb 13: rho=+0.9, weak s/n

## Reuse data from C3

We can use the same dataset from combination 3.

```{r}
X.c13 <- X.c3
dim(X.c13)
round(cor(X.c13), 10)
```

## Start tweaking beta values (from C8)
To reduce to weak s/n (i.e. 2 strong sig, 2 weak sig), - perhaps
we can start with the same $\beta$ from Combination 8 (rho=+0.9,
moderate s/n), - that is $\beta = [3, 0.6, -3, -0.29]^T$ -- here is the
first result

We try to reduce magnitude for X2 first then X4.
- coefficient for X2 altered from 0.6 to 0.5, 
- coefficient for X4 altered from -0.29 to -0.26

```{r}
result.c13 = check_significance(c(3, 0.5, -3, -0.26), X.c13)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c13$y.syn))
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c13,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c13.rds"))
```



# ✅Comb 14: rho=-0.5, weak s/n

## Reuse data from C4
```{r}
X.c14 <- X.c4
dim(X.c14)
round(cor(X.c14), 10)
```

## Start tweaking beta values (from C12)

Since data in C12 have rho=+0.5 and weak s/n (2 strong 2 weak), the respective beta value
could be a good start (since only differ by correlation coeff sign)
- namely from (3, 0.23, -3, -0.31).

Decreasing magnitude of beta for X4 doesn't work (instead it became significant).
So tried to decrease magnitude of beta for X2 just a bit (since already weak).

Nope. doesn't work either.

What ended up working is to decrease magnitude of X2,X4 concurrently.
- Arrived at 0.22 and -0.22 for X2 and X4 respectively. X2 was already
weak by then so just needed to decrease magnitude for X4 to -0.19.


```{r}
result.c14 = check_significance(c(3, 0.2, -3, -0.19), X.c14)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c14$y.syn))
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c14,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c14.rds"))
```

# ✅Comb 15: rho=-0.9, weak s/n

## Reuse data from C5

```{r}
X.c15 <- X.c5
dim(X.c15)
round(cor(X.c15), 10)
```

## Start tweaking beta values (from C13)
As the data from C13 has rho=+0.9 and weak s/n, using its resulting
beta values can be a starting point (as here only difference is the sign of the
correlation)
- To recap, the resulting beta values from C13 are (3, 0.5, -3, -0.26)

Well, only had to decrease magnitude of beta for X4 from -0.26 to -0.25.

```{r}
result.c15 = check_significance(c(3, 0.5, -3, -0.25), X.c15)
```

```{r}
cat("Distribution of synthetic response below:")
print(table(result.c15$y.syn)/ 2000)
```


## Saving the results

```{r}
# save the results for simulation scenario (design matrix, response, 
# and other metadata that will not be used for the RS procedure)
saveRDS(result.c15,
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c15.rds"))
```







