---
title: "TC_sim_factorial_analysis_RS"
author: "Keith Howen"
date: "2025-07-14"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))

source(here("lib", "probit_gibbs.R"))
source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)

library(mvnTest)

#install.packages("mvnTest")

#library(splitstackshape)
```

## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
```



# ____
# ✓ C1: Corr(X1,X2)=0, high s/n

## Read synthetic dataset

```{r}

y.c1 = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "y.c1.rds"))


# save design matrix of candidate variables
X.c1 = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "X.c1.rds"))

```


```{r}
round(cor(X.c1), 5)
```
## Perform the MCMC RS

```{r}
gibbs.RS.c1 =
              Gibbs.random.search.basin(
                       rep(1, ncol(X.c1)),
                       #v02,
                       data.frame(cbind(X.c1, y.c1)),
                       burnin=200,
                       iter=1000,
                       penalty='BIC')


rel.freq.inference(gibbs.RS.c1, CUTOFF=0.5)
```

```{r}
gibbs.RS.c1.best.Xs = colnames(X.c1)[
  as.logical(find_MAP(gibbs.RS.c1))
]

gibbs.RS.c1.best.Xs
```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c1.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c1, y.c1)), gibbs.RS.c1)
)

gibbs.RS.c1.best.Xs.BIC

saveRDS(gibbs.RS.c1.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c1.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c1.rel.freqs = rel.freq.inference(gibbs.RS.c1, CUTOFF=0.5)

saveRDS(gibbs.RS.c1,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c1.rds"))


saveRDS(gibbs.RS.c1.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c1.rel.freqs.rds"))

saveRDS(gibbs.RS.c1.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c1.best.Xs.rds"))

```

# ____
# ✓ C2: Corr(X1,X2)=0.5, high s/n

## Read synthetic dataset

```{r}

result.c2.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c2.rds"))

names(result.c2.read)

X.c2 = result.c2.read$X.syn
y.c2 = result.c2.read$y.syn

# verifying synthetic dataset properties
dim(X.c2)
table(y.c2)
round(cor(X.c2), 5)

```
## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c2 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c2)),
    data.frame(cbind(X.c2, y.c2)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c2, CUTOFF=0.5)

```
```{r}
gibbs.RS.c2.best.Xs = colnames(X.c2)[
  as.logical(find_MAP(gibbs.RS.c2))
]

gibbs.RS.c2.best.Xs
```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c2.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c2, y.c2)), gibbs.RS.c2)
)

gibbs.RS.c2.best.Xs.BIC

saveRDS(gibbs.RS.c2.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c2.best.Xs.BIC.rds"))

```





## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c2.rel.freqs = rel.freq.inference(gibbs.RS.c2, CUTOFF=0.5)

saveRDS(gibbs.RS.c2,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c2.rds"))


saveRDS(gibbs.RS.c2.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c2.rel.freqs.rds"))

saveRDS(gibbs.RS.c2.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c2.best.Xs.rds"))

```


# ____
# ✓ C3: Corr(X1,X2)=0.9, high s/n

## Read synthetic dataset

```{r}

result.c3.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c3.rds"))

names(result.c3.read)

X.c3 = result.c3.read$X.syn
y.c3 = result.c3.read$y.syn

# verifying synthetic dataset properties
dim(X.c3)
table(y.c3)
round(cor(X.c3), 5)

```

## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c3 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c3)),
    data.frame(cbind(X.c3, y.c3)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c3, CUTOFF=0.5)

```

```{r}

gibbs.RS.c3.best.Xs = colnames(X.c3)[
  as.logical(find_MAP(gibbs.RS.c3))
]

gibbs.RS.c3.best.Xs

```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c3.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c3, y.c3)), gibbs.RS.c3)
)

gibbs.RS.c3.best.Xs.BIC

saveRDS(gibbs.RS.c3.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c3.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c3.rel.freqs = rel.freq.inference(gibbs.RS.c3, CUTOFF=0.5)

saveRDS(gibbs.RS.c3,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c3.rds"))


saveRDS(gibbs.RS.c3.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c3.rel.freqs.rds"))

saveRDS(gibbs.RS.c3.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c3.best.Xs.rds"))

```


# ____
# ✓ C4: Corr(X1,X2)= -0.5, high s/n

## Read Synthetic Dataset

```{r}

result.c4.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c4.rds"))

names(result.c4.read)

X.c4 = result.c4.read$X.syn
y.c4 = result.c4.read$y.syn

# verifying synthetic dataset properties
dim(X.c4)
table(y.c4)
round(cor(X.c4), 5)

summary(result.c4.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c4 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c4)),
    data.frame(cbind(X.c4, y.c4)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c4, CUTOFF=0.5)



colnames(X.c4)[
  as.logical(find_MAP(gibbs.RS.c4))
]


```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c4.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c4, y.c4)), gibbs.RS.c4)
)

gibbs.RS.c4.best.Xs.BIC

saveRDS(gibbs.RS.c4.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c4.best.Xs.BIC.rds"))

```


## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c4.rel.freqs = rel.freq.inference(gibbs.RS.c4, CUTOFF=0.5)

gibbs.RS.c4.best.Xs = colnames(X.c4)[
  as.logical(find_MAP(gibbs.RS.c4))
]

saveRDS(gibbs.RS.c4,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c4.rds"))


saveRDS(gibbs.RS.c4.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c4.rel.freqs.rds"))

saveRDS(gibbs.RS.c4.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c4.best.Xs.rds"))


```

# ____
# ✓  C5: Corr(X1,X2)= -0.9, high s/n

## Read synthetic dataset

```{r}

result.c5.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c5.rds"))

names(result.c5.read)

X.c5 = result.c5.read$X.syn
y.c5 = result.c5.read$y.syn

# verifying synthetic dataset properties
dim(X.c5)
table(y.c5)
round(cor(X.c5), 5)

summary(result.c5.read$mod)

```
## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c5 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c5)),
    data.frame(cbind(X.c5, y.c5)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c5, CUTOFF=0.5)



colnames(X.c5)[
  as.logical(find_MAP(gibbs.RS.c5))
]


```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c5.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c5, y.c5)), gibbs.RS.c5)
)

gibbs.RS.c5.best.Xs.BIC

saveRDS(gibbs.RS.c5.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c5.best.Xs.BIC.rds"))

```



## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c5.rel.freqs = rel.freq.inference(gibbs.RS.c5, CUTOFF=0.5)

gibbs.RS.c5.best.Xs = colnames(X.c5)[
  as.logical(find_MAP(gibbs.RS.c5))
]

saveRDS(gibbs.RS.c5,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c5.rds"))


saveRDS(gibbs.RS.c5.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c5.rel.freqs.rds"))

saveRDS(gibbs.RS.c5.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c5.best.Xs.rds"))


```

# ____
# ✔ C6: Corr(X1,X2)=0, moderate s/n

## Read synthetic dataset

```{r}

result.c6.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c6.rds"))

names(result.c6.read)

X.c6 = result.c6.read$X.syn
y.c6 = result.c6.read$y.syn

# verifying synthetic dataset properties
dim(X.c6)
table(y.c6)
round(cor(X.c6), 5)
summary(result.c6.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c6 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c6)),
    data.frame(cbind(X.c6, y.c6)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c6, CUTOFF=0.5)



colnames(X.c6)[
  as.logical(find_MAP(gibbs.RS.c6))
]


```

```{r}

gibbs.RS.c6.best.Xs = colnames(X.c6)[
  as.logical(find_MAP(gibbs.RS.c6))
]

gibbs.RS.c6.best.Xs

```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c6.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c6, y.c6)), gibbs.RS.c6)
)

gibbs.RS.c6.best.Xs.BIC

saveRDS(gibbs.RS.c6.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c6.best.Xs.BIC.rds"))

```



## Save samples & relative frequencies & best model

```{r}

gibbs.RS.c6.rel.freqs = rel.freq.inference(gibbs.RS.c6, CUTOFF=0.5)

saveRDS(gibbs.RS.c6,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c6.rds"))


saveRDS(gibbs.RS.c6.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c6.rel.freqs.rds"))

saveRDS(gibbs.RS.c6.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c6.best.Xs.rds"))


```



# ____
# ✔ C7: Corr(X1,X2)=0.5, moderate s/n

## Read synthetic dataset

```{r}

result.c7.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c7.rds"))

names(result.c7.read)

X.c7 = result.c7.read$X.syn
y.c7 = result.c7.read$y.syn

# verifying synthetic dataset properties
dim(X.c7)
table(y.c7)
round(cor(X.c7), 5)
summary(result.c7.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c7 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c7)),
    data.frame(cbind(X.c7, y.c7)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c7, CUTOFF=0.5)

```


```{r}

gibbs.RS.c7.best.Xs = colnames(X.c7)[
  as.logical(find_MAP(gibbs.RS.c7))
]

gibbs.RS.c7.best.Xs

```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c7.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c7, y.c7)), gibbs.RS.c7)
)

gibbs.RS.c7.best.Xs.BIC

saveRDS(gibbs.RS.c7.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c7.best.Xs.BIC.rds"))

```


## Save samples & relative frequencies

```{r}

gibbs.RS.c7.rel.freqs = rel.freq.inference(gibbs.RS.c7, CUTOFF=0.5)

saveRDS(gibbs.RS.c7,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c7.rds"))


saveRDS(gibbs.RS.c7.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c7.rel.freqs.rds"))

saveRDS(gibbs.RS.c7.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c7.best.Xs.rds"))


```

# ____
# ✔ C8: Corr(X1,X2)=0.9, moderate s/n

## Read synthetic dataset
```{r}

result.c8.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c8.rds"))

names(result.c8.read)

X.c8 = result.c8.read$X.syn
y.c8 = result.c8.read$y.syn

# verifying synthetic dataset properties
dim(X.c8)
table(y.c8)
round(cor(X.c8), 5)
summary(result.c8.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c8 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c8)),
    data.frame(cbind(X.c8, y.c8)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c8, CUTOFF=0.5)

```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c8.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c8, y.c8)), gibbs.RS.c8)
)

gibbs.RS.c8.best.Xs.BIC

saveRDS(gibbs.RS.c8.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c8.best.Xs.BIC.rds"))

```


## Save samples & relative frequencies

```{r}

gibbs.RS.c8.rel.freqs = rel.freq.inference(gibbs.RS.c8, CUTOFF=0.5)


gibbs.RS.c8.best.Xs = colnames(X.c8)[
  as.logical(find_MAP(gibbs.RS.c8))
]

gibbs.RS.c8.best.Xs


saveRDS(gibbs.RS.c8,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c8.rds"))


saveRDS(gibbs.RS.c8.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c8.rel.freqs.rds"))

saveRDS(gibbs.RS.c8.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c8.best.Xs.rds"))


```


# ____
# ✔ C9: Corr(X1,X2)= -0.5, moderate s/n


## Read synthetic dataset

```{r}
result.c9.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c9.rds"))

names(result.c9.read)

X.c9 = result.c9.read$X.syn
y.c9 = result.c9.read$y.syn

# verifying synthetic dataset properties
dim(X.c9)
table(y.c9)
round(cor(X.c9), 5)
summary(result.c9.read$mod)
```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c9 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c9)),
    data.frame(cbind(X.c9, y.c9)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c9, CUTOFF=0.5)

```

## Find BIC Minimizer && Save it

```{r}

suppressWarnings(
  gibbs.RS.c9.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c9, y.c9)), gibbs.RS.c9)
)

gibbs.RS.c9.best.Xs.BIC

saveRDS(gibbs.RS.c9.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c9.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies

```{r}

gibbs.RS.c9.rel.freqs = rel.freq.inference(gibbs.RS.c9, CUTOFF=0.5)


gibbs.RS.c9.best.Xs = colnames(X.c9)[
  as.logical(find_MAP(gibbs.RS.c9))
]

gibbs.RS.c9.best.Xs


saveRDS(gibbs.RS.c9,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c9.rds"))


saveRDS(gibbs.RS.c9.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c9.rel.freqs.rds"))

saveRDS(gibbs.RS.c9.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c9.best.Xs.rds"))


```



# ____
# ✔ C10: Corr(X1,X2)= -0.9, moderate s/n


## Read synthetic dataset
```{r}

result.c10.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c10.rds"))

names(result.c10.read)

X.c10 = result.c10.read$X.syn
y.c10 = result.c10.read$y.syn

# verifying synthetic dataset properties
dim(X.c10)
table(y.c10)
round(cor(X.c10), 5)
summary(result.c10.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c10 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c10)),
    data.frame(cbind(X.c10, y.c10)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c10, CUTOFF=0.5)

```


## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c10.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c10, y.c10)), gibbs.RS.c10)
)

gibbs.RS.c10.best.Xs.BIC

saveRDS(gibbs.RS.c10.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c10.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies

```{r}

gibbs.RS.c10.rel.freqs = rel.freq.inference(gibbs.RS.c10, CUTOFF=0.5)


gibbs.RS.c10.best.Xs = colnames(X.c10)[
  as.logical(find_MAP(gibbs.RS.c10))
]

gibbs.RS.c10.best.Xs


saveRDS(gibbs.RS.c10,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c10.rds"))


saveRDS(gibbs.RS.c10.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c10.rel.freqs.rds"))

saveRDS(gibbs.RS.c10.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c10.best.Xs.rds"))


```

# ____
# ✔  C11: Corr(X1,X2)=0, weak s/n


## Read synthetic dataset

```{r}

result.c11.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c11.rds"))

names(result.c11.read)

X.c11 = result.c11.read$X.syn
y.c11 = result.c11.read$y.syn

# verifying synthetic dataset properties
dim(X.c11)
table(y.c11)
round(cor(X.c11), 5)
summary(result.c11.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c11 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c11)),
    data.frame(cbind(X.c11, y.c11)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c11, CUTOFF=0.5)

```


## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c11.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c11, y.c11)), gibbs.RS.c11)
)

gibbs.RS.c11.best.Xs.BIC

saveRDS(gibbs.RS.c11.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c11.best.Xs.BIC.rds"))

```



## Save samples & relative frequencies

```{r}

gibbs.RS.c11.rel.freqs = rel.freq.inference(gibbs.RS.c11, CUTOFF=0.5)


gibbs.RS.c11.best.Xs = colnames(X.c11)[
  as.logical(find_MAP(gibbs.RS.c11))
]

gibbs.RS.c11.best.Xs


saveRDS(gibbs.RS.c11,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c11.rds"))


saveRDS(gibbs.RS.c11.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c11.rel.freqs.rds"))

saveRDS(gibbs.RS.c11.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c11.best.Xs.rds"))


```

# ____
# ✔ C12: Corr(X1,X2)=0.5, weak s/n

## Read synthetic dataset

```{r}

result.c12.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c12.rds"))

names(result.c12.read)

X.c12 = result.c12.read$X.syn
y.c12 = result.c12.read$y.syn

# verifying synthetic dataset properties
dim(X.c12)
table(y.c12)
round(cor(X.c12), 5)
summary(result.c12.read$mod)

```


## Perform the MCMC RS
```{r}

suppressWarnings(
  gibbs.RS.c12 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c12)),
    data.frame(cbind(X.c12, y.c12)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c12, CUTOFF=0.5)

```


## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c12.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c12, y.c12)), gibbs.RS.c12)
)

gibbs.RS.c12.best.Xs.BIC

saveRDS(gibbs.RS.c12.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c12.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies

```{r}

gibbs.RS.c12.rel.freqs = rel.freq.inference(gibbs.RS.c12, CUTOFF=0.5)


gibbs.RS.c12.best.Xs = colnames(X.c12)[
  as.logical(find_MAP(gibbs.RS.c12))
]

gibbs.RS.c12.best.Xs


saveRDS(gibbs.RS.c12,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c12.rds"))


saveRDS(gibbs.RS.c12.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c12.rel.freqs.rds"))

saveRDS(gibbs.RS.c12.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c12.best.Xs.rds"))


```

# ____
# ✔ C13: Corr(X1,X2)=0.9, weak s/n

## Read synthetic dataset

```{r}

result.c13.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c13.rds"))

names(result.c13.read)

X.c13 = result.c13.read$X.syn
y.c13 = result.c13.read$y.syn

# verifying synthetic dataset properties
dim(X.c13)
table(y.c13)
round(cor(X.c13), 5)
summary(result.c13.read$mod)

```


## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c13 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c13)),
    data.frame(cbind(X.c13, y.c13)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c13, CUTOFF=0.5)

```


## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c13.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c13, y.c13)), gibbs.RS.c13)
)

gibbs.RS.c13.best.Xs.BIC

saveRDS(gibbs.RS.c13.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c13.best.Xs.BIC.rds"))

```


## Save samples & relative frequencies

```{r}

gibbs.RS.c13.rel.freqs = rel.freq.inference(gibbs.RS.c13, CUTOFF=0.5)


gibbs.RS.c13.best.Xs = colnames(X.c13)[
  as.logical(find_MAP(gibbs.RS.c13))
]

gibbs.RS.c13.best.Xs


saveRDS(gibbs.RS.c13,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c13.rds"))


saveRDS(gibbs.RS.c13.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c13.rel.freqs.rds"))

saveRDS(gibbs.RS.c13.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c13.best.Xs.rds"))


```


# ____
# ✔  C14: Corr(X1,X2)= -0.5, weak s/n

## Read synthetic dataset

```{r}

result.c14.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c14.rds"))

names(result.c14.read)

X.c14 = result.c14.read$X.syn
y.c14 = result.c14.read$y.syn

# verifying synthetic dataset properties
dim(X.c14)
table(y.c14)
round(cor(X.c14), 5)
summary(result.c14.read$mod)

```


## Perform the MCMC RS
```{r}

suppressWarnings(
  gibbs.RS.c14 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c14)),
    data.frame(cbind(X.c14, y.c14)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c14, CUTOFF=0.5)

```
## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c14.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c14, y.c14)), gibbs.RS.c14)
)

gibbs.RS.c14.best.Xs.BIC

saveRDS(gibbs.RS.c14.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c14.best.Xs.BIC.rds"))

```


## Save samples & relative frequencies

```{r}

gibbs.RS.c14.rel.freqs = rel.freq.inference(gibbs.RS.c14, CUTOFF=0.5)


gibbs.RS.c14.best.Xs = colnames(X.c14)[
  as.logical(find_MAP(gibbs.RS.c14))
]

gibbs.RS.c14.best.Xs


saveRDS(gibbs.RS.c14,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c14.rds"))


saveRDS(gibbs.RS.c14.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c14.rel.freqs.rds"))

saveRDS(gibbs.RS.c14.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c14.best.Xs.rds"))


```

# ____
# ✔ C15: Corr(X1,X2)= -0.9, weak s/n

## Read synthetic dataset
```{r}

result.c15.read = readRDS(
        here("output",
             "simulation_factorial_analysis_outputs",
             "result.c15.rds"))

names(result.c15.read)

X.c15 = result.c15.read$X.syn
y.c15 = result.c15.read$y.syn

# verifying synthetic dataset properties
dim(X.c15)
table(y.c15)
round(cor(X.c15), 5)
summary(result.c15.read$mod)

```

## Perform the MCMC RS

```{r}

suppressWarnings(
  gibbs.RS.c15 <- Gibbs.random.search.basin(
    rep(1, ncol(X.c15)),
    data.frame(cbind(X.c15, y.c15)),
    burnin = 200,
    iter = 1000,
    penalty = 'BIC'
  )
)


rel.freq.inference(gibbs.RS.c15, CUTOFF=0.5)

```
## Find BIC Minimizer && Save It

```{r}

suppressWarnings(
  gibbs.RS.c15.best.Xs.BIC <- find_BIC_minimizer(data.frame(cbind(X.c15, y.c15)), gibbs.RS.c15)
)

gibbs.RS.c15.best.Xs.BIC

saveRDS(gibbs.RS.c15.best.Xs.BIC,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c15.best.Xs.BIC.rds"))

```

## Save samples & relative frequencies


```{r}

gibbs.RS.c15.rel.freqs = rel.freq.inference(gibbs.RS.c15, CUTOFF=0.5)


gibbs.RS.c15.best.Xs = colnames(X.c15)[
  as.logical(find_MAP(gibbs.RS.c15))
]

gibbs.RS.c15.best.Xs


saveRDS(gibbs.RS.c15,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c15.rds"))


saveRDS(gibbs.RS.c15.rel.freqs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c15.rel.freqs.rds"))

saveRDS(gibbs.RS.c15.best.Xs,
        here("output",
             "simulation_factorial_analysis_outputs",
             "gibbs.RS.c15.best.Xs.rds"))


```



# Misc exploration of TC genesis dataset.

```{r}

DH.test(
  data[,colnames(data)[!colnames(data) %in% c("basin", "TC_genesis")]],
  qqplot = TRUE
)

```








