---
title: "record_simulation_setups"
author: "Keith Howen"
date: "2025-09-23"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(NA)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))

#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
#source(here("lib", "factorial_analysis_routines.R"))

#library(mvtnorm)
#library(truncnorm)
#library(Matrix)
#library(MASS)
#library(splitstackshape)

library(knitr)
library(kableExtra)
```


```{r}
list.files(here("output"))
```


```{r}
list.files(here("output", "simulation_factorial_analysis_n_2000_p_160_outputs" ))
```

# n=200,p=16

## reading sim quantities

```{r}

results_n_200_p_16 = readRDS(
  here("output",
       "simulation_factorial_analysis_n_200_p_16_outputs",
       "results.rds")
)

```

```{r}
names(results_n_200_p_16)
names(results_n_200_p_16[["0"]])
```

```{r}
names(results_n_200_p_16[["0"]][["strong"]])
```



```{r}
CORR_LEVELS = c("0", "0.5", "0.9", "-0.5", "-0.9")
SN_LEVELS = c("strong", "moderate", "weak")


for (c_level in CORR_LEVELS){
  for (sn_level in SN_LEVELS){
    #cat(sprintf("corr_level: %s, sn_level: %s\n", c_level, sn_level))
    
    xtable_obj = xtable(
                    summary(results_n_200_p_16[[c_level]][[sn_level]]$mod),
                    digits=c(0,9,9,9,9),
                    caption=sprintf("$Corr(X1,X2)=%s$, SNR=%s", c_level, sn_level)
        )
    
    print(xtable_obj, type='latex', table.placement="H")

    
    #cat("--------------------------\n")
    
  }
}

```



# n=2000,p=16
```{r}
list.files(here("output", "simulation_factorial_analysis_n_2000_p_16_outputs" ))
```

```{r}

names(
  readRDS(
  here("output", "simulation_factorial_analysis_n_2000_p_16_outputs",
       "result.c10.rds")
)
)




```





✅Combination 1/c1: corr(X1,X2)=0, high s/n
✅Combination 2/c2: corr(X1,X2)=+0.5, high s/n
✅Combination 3/c3: corr(X1,X2)=+0.9, high s/n
✅Combination 4/c4: corr(X1,X2)=-0.5, high s/n
✅Combination 5/c5: corr(X1,X2)=-0.9, high s/n
✅Combination 6/c6: corr(X1,X2)=0, moderate s/n
✅Combination 7/c7: corr(X1,X2)=+0.5, moderate s/n
✅Combination 8/c8: corr(X1,X2)=+0.9, moderate s/n
✅Combination 9/c9: corr(X1,X2)=-0.5, moderate s/n
✅Combination 10/c10: corr(X1,X2)=-0.9, moderate s/n
✅Combination 11/c11: corr(X1,X2)=0, weak s/n
✅Combination 12/c12: corr(X1,X2)=+0.5, weak s/n
✅Combination 13/c13: corr(X1,X2)=+0.9, weak s/n
✅Combination 14/c14: corr(X1,X2)=-0.5, weak s/n
✅Combination 15/c15: corr(X1,X2)=-0.9, weak s/n


```{r}
CORR_LEVELS = c("0", "+0.5", "+0.9", "-0.5", "-0.9")
SN_LEVELS = c("strong", "moderate", "weak")

meta <- tibble(
    combination = 1:15,
    correlation = rep(CORR_LEVELS, length(SN_LEVELS)),
    s_n = rep(SN_LEVELS, each = length(CORR_LEVELS))
  )
meta
```


```{r}

read_all_beta_coefs = function(){
  
  BASE_DIR = "simulation_factorial_analysis_n_2000_p_16_outputs"
  
  ## 1. INITS 
  # Metadata for combinations
  meta <- tibble(
    combination = 1:15,
    correlation = rep(CORR_LEVELS, length(SN_LEVELS)),
    s_n = rep(SN_LEVELS, each = length(CORR_LEVELS))
  )
  
  # dict with 2 keys essentially
  results <- list()
  for (rho in CORR_LEVELS) {
    results[[rho]] <- list()
    for (s in SN_LEVELS) {
      results[[rho]][[s]] <- "empty"  
    }
  }
  
  ## 2. store MCMC samples
  for (i in 2:15) {
    
    combo_id <- paste0("c", i)
    path_base <- here("output", BASE_DIR)
    
    row = readRDS(file.path(path_base, paste0("result.", combo_id, ".rds")))

    # retrieve corresponding (corr_level, sn_level) from
    # the 'meta' metadata -- have to use redirection like operator
    # to filter tibbles
    corr_level = (meta %>% filter(combination == i))$correlation
    sn_level = (meta %>% filter(combination == i))$s_n
    
    results[[corr_level]][[sn_level]] = row
  }
  
  return(results)
  
}



# Final summary table
results_n_2000_p_16 = read_all_beta_coefs()
```


```{r}
results_n_2000_p_16[["0"]][["strong"]] = list()

X.c1 = readRDS(here("output", "simulation_factorial_analysis_n_2000_p_16_outputs",
       "X.c1.rds"))

y.c1 = readRDS(here("output", "simulation_factorial_analysis_n_2000_p_16_outputs",
       "y.c1.rds"))


results_n_2000_p_16[["0"]][["strong"]]$mod = glm(y.c1 ~ .,
                                                 data=data.frame(X.c1[,1:4],y.c1),
                                                 family=binomial)
```



```{r}
names(results_n_2000_p_16)
```
```{r}
names(results_n_2000_p_16[["0"]])
```
```{r}
names(results_n_2000_p_16[["0"]][["strong"]])
```
```{r}
results_n_2000_p_16[["+0.5"]][["strong"]]$mod
```
```{r}
xtable(
  summary(results_n_2000_p_16[["+0.5"]][["strong"]]$mod,
          digits=c(0,9,9,9,9))
)
```


```{r}
CORR_LEVELS = c("0", "+0.5", "+0.9", "-0.5", "-0.9")
SN_LEVELS = c("strong", "moderate", "weak")


for (c_level in CORR_LEVELS){
  for (sn_level in SN_LEVELS){
    cat(sprintf("corr_level: %s, sn_level: %s\n", c_level, sn_level))
    
    print(
      xtable(
          summary(results_n_2000_p_16[[c_level]][[sn_level]]$mod,
                  digits=c(0,9,9,9,9))
        )
    )
    cat("--------------------------\n")
    
  }
}

```



# n=2000,p=160
```{r}
CORR_LEVELS = c("0","0.9")
SN_LEVELS = c("moderate")

results_n_2000_p_160 = readRDS(
  here("output",
       "simulation_factorial_analysis_n_2000_p_160_outputs",
       "results_n_2000_p_160.rds")
)


for (c_level in CORR_LEVELS){
  for (sn_level in SN_LEVELS){
    #cat(sprintf("corr_level: %s, sn_level: %s\n", c_level, sn_level))
    
    xtable_obj = xtable(
                    summary(results_n_2000_p_160[[c_level]][[sn_level]]$mod),
                    digits=c(0,9,9,9,9),
                    caption=sprintf("$Corr(X1,X2)=%s$, SNR=%s", c_level, sn_level)
        )
    
    print(xtable_obj, type='latex', table.placement="H")

    
    #cat("--------------------------\n")
    
  }
}

```



# n=10000, p=160
```{r}
CORR_LEVELS = c("0.9")
SN_LEVELS = c("moderate")

results_n_2000_p_160 = readRDS(
  here("output",
       "simulation_factorial_analysis_n_10000_p_160_outputs",
       "results_n_10000_p_160.rds")
)


for (c_level in CORR_LEVELS){
  for (sn_level in SN_LEVELS){
    #cat(sprintf("corr_level: %s, sn_level: %s\n", c_level, sn_level))
    
    xtable_obj = xtable(
                    summary(results_n_2000_p_160[[c_level]][[sn_level]]$mod),
                    digits=c(0,9,9,9,9),
                    caption=sprintf("$Corr(X1,X2)=%s$, SNR=%s", c_level, sn_level)
        )
    
    print(xtable_obj, type='latex', table.placement="H")

    
    #cat("--------------------------\n")
    
  }
}

```


