---
title: "analyze_potential_outliers"
author: "Keith Howen"
date: "2025-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Check working directory

```{r}
library(here)
here()
```


# Rationale
logistic regression CAN be affected by outliers as it tries to align predicted
probabilities with class labels as close as possible.

Concern is that this could affect model fittings which in turns affect the procedure.


# Import all libraries (incl. custom)

```{r}
source(here("lib", "load_and_preprocess_TC.R"))
source(here("lib", "gibbs_sampler_random_search.R"))


install.packages('robustbase')
library(robustbase)
```



# Detect high leverage observations

```{r}

detect_high_leverage <- function(X,
                                 method = c("chisq","iqr"),
                                 alpha = 0.975,
                                 return = c("indices","distances"),
                                 iqr_scaling=1.5) {
  method  <- match.arg(method)
  return  <- match.arg(return)
  
  # standardize and reduce dimension
  X_scaled <- scale(X)
  pca_fit  <- prcomp(X_scaled, center = TRUE, scale. = TRUE)
  
  # keep PCs that explain 95% variance
  var_expl <- cumsum(pca_fit$sdev^2) / sum(pca_fit$sdev^2)
  k <- which(var_expl >= 0.95)[1]
  X_pca <- pca_fit$x[, 1:k]
  
  # robust covariance + Mahalanobis distance
  cov_rob <- robustbase::covMcd(X_pca) 
  mdist   <- mahalanobis(X_pca, center = cov_rob$center, cov = cov_rob$cov)
  
  if (method == "chisq") {
    cutoff   <- qchisq(alpha, df = k)
    outliers <- which(mdist > cutoff)
    title <- sprintf("Robust Mahalanobis distances\n(Flagged %d obs, α=%.3f)", length(outliers), alpha)
    abline_h <- cutoff
  } else if (method == "iqr") {
    Q1 <- quantile(mdist, 0.25)
    Q3 <- quantile(mdist, 0.75)
    IQR_val <- Q3 - Q1
    cutoff <- Q3 + iqr_scaling * IQR_val
    outliers <- which(mdist > cutoff)
    title <- sprintf("Robust Mahalanobis distances\n(Flagged %d obs by dists > Q3 + %.1f * IQR)",
                     length(outliers),
                     iqr_scaling)
    abline_h <- cutoff
  }
  
  # plot
  plot(mdist, pch=20, col = ifelse(seq_along(mdist) %in% outliers, "red","black"),
       main=title, ylab="Distance", xlab="Obs index")
  abline(h=abline_h, col="blue", lty=2)
  
  if (return == "indices") {
    return(outliers)
  } else {
    return(data.frame(index=seq_along(mdist), mdist=mdist, flagged=seq_along(mdist) %in% outliers))
  }
}


high_leverage_idx = detect_high_leverage(X,
                                         method="iqr",
                                         return="indices",
                                         iqr_scaling=55)

```

```{r}
X[high_leverage_idx,]
```

```{r}
xvar = "absolute_vorticity_averaged"

dummy_y = rep(0, nrow(X))
emphasize_high_leverages = 0.000000005
dummy_y[high_leverage_idx]=emphasize_high_leverages

plot(X[,xvar],
     dummy_y,
     pch=20,
     col = ifelse(1:nrow(X) %in% high_leverage_idx, "red","black"),
       main=sprintf("Dot plot for %s", xvar),
       xlab=xvar,
       yaxt = "n",
       ylab="",
     ylim=c(0, emphasize_high_leverages + 0.00000000005))
```

# Try to run RS without high leverage observations
```{r}
mh.no.high.leverages = MH.random.search(
                             rep(1,ncol(X)),
                             data=data.frame(X[-high_leverage_idx,],y[-high_leverage_idx]),
                             burnin = 200,
                             iter   = 5200,
                             debug  = TRUE,
                             lambda = 1)
```





