---
title: "discuss_TC_genesis_RS_results"
author: "Keith Howen"
date: "2025-09-06"
output: html_document
---

# 💻 set seed for reproducibility
```{r}
set.seed(123)
```



# 💻 Working Directory Setup using `here()`
Use `here()` to *standardize the working directory between R markdown and
R console* as opposed to using `getwd()` (and `setwd()`), at the slight
inconveniences of importing the `here` library, using `here()` to write file
paths, and manually specify paths for files outside the working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports
## Check working directory
```{r}
here()
```

## Import all libraries (incl. custom)
```{r}

# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))
source(here("lib", "load_and_preprocess_TC.R"))
#library(mvtnorm)
#library(truncnorm)
#library(Matrix)
library(ggplot2)
#library(splitstackshape)
#install.packages('reshape2')
library(reshape2)

library(glmnet)
#install.packages('caret')
library(caret)

```


## Set path for data

```{r}
list.files(here("data"))
```

# 💻 Import raw data & fix NA basin value

```{r}
#data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
# import the data
#data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
#data.NA.fixed = fix_NA_value_and_refactor_basin(data)
#data.NA.fixed.no.basin = data.NA.fixed[colnames(data.NA.fixed) != "basin"]
#data.NA.fixed.no.TC = data.NA.fixed.no.basin[colnames(data.NA.fixed.no.basin) != "TC_genesis"]

#X.colnames = colnames(data.NA.fixed.no.TC)

#X = data.NA.fixed.no.TC
#y = data.NA.fixed.no.basin["TC_genesis"]

```


# Import result quantities

## Results from first run

```{r}
# set dir names
TC_RESULTS_DIR = here("output", "random_search_TC_dataset_outputs", "RCP_outputs")
VALIDATION_OUTPUTS_DIR = here("output", "validation_outputs")


# relative frequencies
TC.rel.freqs = readRDS(file.path(TC_RESULTS_DIR, "rel.freqs.rds"))$rel.freqs.non.basin
TC.rel.freqs.sorted = sort(TC.rel.freqs, decreasing=TRUE)


# best var set from BIC minimizer in RS
TC.BIC.minimizer.rds = readRDS(file.path(TC_RESULTS_DIR, "BIC.minimizer.rds"))
rs_bic_min_vars = TC.BIC.minimizer.rds$models[[1]]$selected.variables


mcmc.samples.tc.genesis = readRDS(here("output", "random_search_TC_dataset_outputs", "RCP_outputs", "gibbs_TC_dataset_results_iter_1.rds")) 

# lasso object
tc.genesis.lasso.lm = readRDS(file.path(VALIDATION_OUTPUTS_DIR, "cvfit_lasso.rds"))
  

```




## Results from multiple chains run

```{r}

mcmc.samples.TC.multiple.chains = readRDS(file.path(TC_RESULTS_DIR, "gibbs_TC_dataset_multiple_chains.rds"))

```

```{r}

cat("Number of chains being run: ", length(mcmc.samples.TC.multiple.chains), "\n")
cat("Dim of samples for chain 1: ", dim(mcmc.samples.TC.multiple.chains[[1]]), "\n")
cat("Dim of samples for chain 2: ", dim(mcmc.samples.TC.multiple.chains[[2]]), "\n")
cat("Dim of samples for chain 3: ", dim(mcmc.samples.TC.multiple.chains[[3]]), "\n")
cat("Dim of samples for chain 4: ", dim(mcmc.samples.TC.multiple.chains[[4]]), "\n")


```

# Check diagnostics (using GR)

```{r}

GR_diagnostics <- function(results,
                           drop_degenerate = TRUE,
                           trim_last_half = FALSE,
                           cum_sum = FALSE) {
  if (!requireNamespace("coda", quietly = TRUE)) install.packages("coda")
  library(coda)
  
  # relative inclusion frequency per var across *all* chains
  rel_freqs <- colMeans(do.call(rbind, results))
  
  keep_idx <- seq_len(ncol(results[[1]]))
  if (drop_degenerate) {
    keep_idx <- which(rel_freqs > 0 & rel_freqs < 1)
  }
  
  # compute running inclusion prob per var
  if (cum_sum){
    chains_ip <- lapply(results, function(S) {
      apply(S[, keep_idx, drop = FALSE], 2, function(col) cumsum(col) / seq_along(col))
    })
  } else{
    chains_ip <- lapply(results, function(S)S[, keep_idx])
  }

  
  if (trim_last_half){
    chains_ip <- lapply(chains_ip, function(M) M[floor(nrow(M)/2):nrow(M), , drop = FALSE])
  }
  
  
  mcmc_list <- mcmc.list(lapply(chains_ip, function(M) mcmc(M)))
  gr <- gelman.diag(mcmc_list, multivariate = FALSE)$psrf
  
  ess <- sapply(1:length(keep_idx), function(j)
    sum(sapply(results, function(S) coda::effectiveSize(S[, keep_idx[j]]))))
  
  list(gr = gr,
       ess = ess,
       kept_vars = colnames(results[[1]])[keep_idx],
       dropped_vars = if (drop_degenerate) colnames(results[[1]])[-keep_idx] else NULL)
}


GR_diagnostics_v2 <- function(results,
                           split_chains = TRUE,
                           drop_rule = c("per_chain","global","none"),
                           rhat_cut = 1.05) {
  stopifnot(is.list(results), length(results) >= 2)
  if (!requireNamespace("coda", quietly = TRUE)) install.packages("coda")
  drop_rule <- match.arg(drop_rule)

  # 1) Collect raw indicator matrices per chain
  mats <- lapply(results, function(S) {
    stopifnot(is.matrix(S))
    S
  })

  # 2) Split each chain into two halves (BDA3 recommendation)
  if (split_chains) {
    mats <- unlist(lapply(mats, function(S) {
      n <- nrow(S); h <- floor(n/2)
      list(S[1:h, , drop = FALSE], S[(h+1):n, , drop = FALSE])
    }), recursive = FALSE)
  }

  p <- ncol(mats[[1]])
  vnames <- colnames(mats[[1]])

  # 3) Choose variables to keep (avoid degeneracy)
  keep <- rep(TRUE, p); why_drop <- rep(NA_character_, p)
  if (drop_rule != "none") {
    if (drop_rule == "per_chain") {
      wvar <- sapply(mats, function(S) apply(S, 2, var))
      keep <- apply(wvar > 0, 1, all)
      why_drop[!keep] <- "degenerate_in_some_chain"
    } else if (drop_rule == "global") {
      pooled <- do.call(rbind, mats)
      keep <- apply(pooled, 2, var) > 0
      why_drop[!keep] <- "degenerate_globally"
    }
  }
  if (!any(keep)) stop("No non-degenerate variables to diagnose.")

  # 4) Build coda mcmc.list from subchains
  mlist <- coda::mcmc.list(lapply(mats, function(S) coda::mcmc(S[, keep, drop = FALSE])))

  # 5) Univariate GR and ESS (ESS summed across subchains)
  gr   <- coda::gelman.diag(mlist, multivariate = FALSE)$psrf
  ess  <- sapply(seq_len(sum(keep)), function(j)
           sum(sapply(mats, function(S) coda::effectiveSize(S[, j]))))

  out <- data.frame(
    variable   = vnames[keep],
    Rhat_point = gr[, "Point est."],
    Rhat_upper = gr[, "Upper C.I."],
    ESS_sum    = ess,
    converged  = gr[, "Point est."] < rhat_cut
  )
  rownames(out) <- NULL

  list(
    table   = out[order(out$Rhat_point, decreasing = TRUE), ],
    dropped = data.frame(variable = vnames[!keep], reason = why_drop[!keep]),
    params  = list(split_chains = split_chains, drop_rule = drop_rule, rhat_cut = rhat_cut)
  )
}


```

## V1: no splitting

```{r}
mcmc.samples.TC.multiple.chains.diag = GR_diagnostics(mcmc.samples.TC.multiple.chains)
```


## V2: split

```{r}
mcmc.samples.TC.multiple.chains.diag2 = GR_diagnostics_v2(mcmc.samples.TC.multiple.chains,
                                                          rhat_cut = 1.1)


```


```{r}
names(mcmc.samples.TC.multiple.chains.diag2)
```
```{r}
mcmc.samples.TC.multiple.chains.diag2$dropped
```

```{r}
mcmc.samples.TC.multiple.chains.diag2$table
```


```{r}
mcmc.samples.TC.multiple.chains.diag2$table %>% filter(converged==FALSE)
```


## V2.1: further burn-in, then split

### How many more burn-in samples?

```{r}


for(extra_burn in 10*(5:40)){
  
  trimmed <- lapply(mcmc.samples.TC.multiple.chains, function(S)
    S[(extra_burn + 1):nrow(S), , drop = FALSE]
  )
  
  mcmc.samples.TC.multiple.chains.diag2.trimmed = 
    GR_diagnostics_v2(
      trimmed,
      rhat_cut = 1.1)
  
  prop_vars_converged = nrow(mcmc.samples.TC.multiple.chains.diag2.trimmed$table %>% filter(converged == TRUE)) /
    nrow(mcmc.samples.TC.multiple.chains.diag2.trimmed$table)
  
  prop_vars_converged = round(prop_vars_converged,2)
  
  cat("Proportion of variables converged if burn-in is extended by", extra_burn, ":", prop_vars_converged, "\n")
  
}



```


### Perform the diagnostic check.

```{r}

extra_burn <- 60

trimmed <- lapply(mcmc.samples.TC.multiple.chains, function(S)
  S[(extra_burn + 1):nrow(S), , drop = FALSE]
)

cat("Dimension of each chain: ", dim(trimmed[[1]]), "\n")


mcmc.samples.TC.multiple.chains.diag2.trimmed = 
  GR_diagnostics_v2(
    trimmed,
    rhat_cut = 1.1)


mcmc.samples.TC.multiple.chains.diag2.trimmed$table


prop_vars_converged = nrow(mcmc.samples.TC.multiple.chains.diag2.trimmed$table %>% filter(converged == TRUE)) /
  nrow(mcmc.samples.TC.multiple.chains.diag2.trimmed$table)

prop_vars_converged = round(prop_vars_converged,2)

cat("Proportion of variables converged if burn-in is extended by", extra_burn, ":", prop_vars_converged)


```
# Check diagnostics (using contingency tables)
```{r}
stationarity_chisq <- function(
  S,                      # numeric vector (one long chain) OR list of numeric vectors (q chains)
  k0 = 0,                 # extra burn-in to drop per chain
  q  = NULL,              # number of sub-sequences (if NULL and S is a vector, chosen automatically)
  r  = NULL,              # number of intervals (if NULL, chosen automatically)
  breaks = c("quantile", "equalwidth"),  # how to make r bins
  min_expected = 5,       # target >= 5 expected counts/cell
  simulate_p = FALSE,     # if some cells < 5, use Monte Carlo p-value
  B = 10000               # replicates for Monte Carlo p-value (chisq.test)
) {
  breaks <- match.arg(breaks)

  # --- helper: coerce input to list of q sub-sequences, each length K ---
  to_blocks <- function(S, k0, q) {
    if (is.list(S)) {
      # q chains provided
      chains <- lapply(S, function(x) {
        stopifnot(is.numeric(x))
        if (k0 >= length(x)) stop("k0 >= chain length.")
        x[(k0 + 1):length(x)]
      })
      K <- min(sapply(chains, length))
      if (K < 10) stop("Not enough post-burn draws per chain for splitting.")
      blocks <- lapply(chains, function(x) x[(length(x) - K + 1):length(x)])
      list(blocks = blocks, q = length(blocks), K = K)
    } else {
      # single long chain: split into q blocks after burn-in
      x <- as.numeric(S)
      if (k0 >= length(x)) stop("k0 >= chain length.")
      x <- x[(k0 + 1):length(x)]
      if (is.null(q)) {
        # aim for at least ~100 per block, but ≥ 2 blocks
        q <- max(2L, min(10L, floor(length(x) / 100)))
      }
      K <- floor(length(x) / q)
      if (K < 10) stop("Not enough draws per block; decrease q or k0.")
      # take the last q*K draws so blocks are equal length
      x <- tail(x, q * K)
      blocks <- split(x, rep(seq_len(q), each = K))
      list(blocks = blocks, q = q, K = K)
    }
  }

  blk <- to_blocks(S, k0 = k0, q = q)
  blocks <- blk$blocks; q <- blk$q; K <- blk$K

  # --- choose r if missing to satisfy ~>= min_expected expected per cell ---
  N_total <- q * K
  if (is.null(r)) {
    r <- max(2L, min(10L, floor(N_total / (min_expected * q))))
  }

  # --- define bin edges over pooled S ---
  pooled <- unlist(blocks, use.names = FALSE)
  if (breaks == "quantile") {
    # quantile bins give near-balanced row totals -> helps the 5-per-cell rule
    probs <- seq(0, 1, length.out = r + 1)
    # jitter duplicates in extremes to avoid identical breaks
    qs <- as.numeric(stats::quantile(pooled, probs, type = 7, na.rm = TRUE))
    eps <- .Machine$double.eps^0.25
    qs <- unique(qs)  # guard: if r is too large for unique values, reduce r
    if (length(qs) < r + 1) {
      r <- length(qs) - 1L
      if (r < 2) stop("Too few unique values in S to form >=2 bins.")
    }
    brks <- if (length(qs) == r + 1) qs else stats::quantile(pooled, seq(0, 1, length.out = r + 1))
    brks[1]  <- brks[1]  - eps
    brks[r+1] <- brks[r+1] + eps
  } else {
    brks <- seq(min(pooled), max(pooled), length.out = r + 1)
  }

  # --- build r x q contingency table ---
  binlab <- function(x) cut(x, breaks = brks, include.lowest = TRUE, right = TRUE)
  tab <- sapply(blocks, function(b) table(binlab(b)))
  # ensure matrix with rownames (intervals) and colnames (block index)
  if (is.null(dim(tab))) tab <- matrix(tab, nrow = r)
  rownames(tab) <- levels(binlab(pooled))
  colnames(tab) <- paste0("block_", seq_len(q))

  # --- chi-squared test of independence ---
  # check expected counts
  exp_counts <- suppressWarnings(chisq.test(tab)$expected)
  small <- any(exp_counts < min_expected)

  test <- suppressWarnings(
    chisq.test(tab, simulate.p.value = simulate_p || small, B = B)
  )

  list(
    p_value = unname(test$p.value),
    statistic = unname(test$statistic),
    df = unname(test$parameter),
    small_expected = small,
    method = if (simulate_p || small) "Chi-squared with simulated p-value" else "Pearson Chi-squared",
    r = r, q = q, K = K, k0 = k0,
    breaks = brks,
    table = tab
  )
}

```

## Compute ALL BICs and perform chisq test
```{r}
start_time = Sys.time()

S_list <- lapply(trimmed, function(S) {
  apply(S, 1, function(v) compute.BIC(data.frame(X, y), v))
})

# CONSIDER SAVING S_List!!!

# no need burn-in -- already calibrated
out <- stationarity_chisq(S_list, k0 = 0)

end_time = Sys.time()
cat("Overall time taken (minutes): ",
      as.numeric(end_time - start_time, units = "mins"),
      "\n")

```

```{r}

saveRDS(S_list,
        here("output", "random_search_TC_dataset_outputs", "all_BICs_4_chains.rds"))

```




```{r}


#length(S_list[[1]])

plot(S_list[[1]], col=1, type='l', xlab='Iteration', ylab='BIC', lwd=4)

for(i in 2:length(S_list)){
  lines(S_list[[i]], col=i, xlab='Iteration', ylab='BIC', lwd=4)
}


```



```{r}
raftery.diag(as.mcmc(unlist(S_list)))
```
## Leaving only 75 samples per chain

```{r}
stationarity_chisq(S_list, k0 = 665)
```



## Thinning to every k-th sample per chain
CAUTION: this is mostly loss of power (each block has K=106
K=106 vsK=740), and thinning can hide slow drifts.

It doesn’t prove stationarity.


```{r}
stationarity_chisq(lapply(S_list, function(S){S[seq(1,length(S),by=7)]}), 
                   k0 = 0)
```



```{r}
stationarity_chisq_v2 <- function(
  S, k0 = 0, r = NULL,
  breaks = c("quantile","equalwidth"),
  blocks_per_chain = 1,   # <— NEW
  min_expected = 5, simulate_p = FALSE, B = 10000
){
  breaks <- match.arg(breaks)

  # Coerce to list of chains
  chains <- if (is.list(S)) S else list(S)
  chains <- lapply(chains, function(x){
    x <- as.numeric(x); if (k0 >= length(x)) stop("k0 >= chain length")
    x[(k0+1):length(x)]
  })

  # Split each chain into 'blocks_per_chain' equal blocks
  split_blocks <- function(x, a){
    n <- length(x); K <- floor(n / a); if (K < 10) stop("block too short")
    x <- tail(x, a*K)                                # equalize
    split(x, rep(seq_len(a), each = K))
  }
  blocks_list <- lapply(chains, split_blocks, a = blocks_per_chain)
  blocks <- do.call(c, blocks_list)                  # concatenated blocks
  q <- length(blocks)
  K <- length(blocks[[1]])

  # Choose r to keep expected counts >= min_expected
  N_total <- q*K
  if (is.null(r)) r <- max(2L, min(10L, floor(N_total / (min_expected*q))))

  pooled <- unlist(blocks, use.names = FALSE)

  if (breaks == "quantile") {
    probs <- seq(0, 1, length.out = r + 1)
    brks <- as.numeric(quantile(pooled, probs, na.rm = TRUE))
    brks[1] <- brks[1] - 1e-8; brks[length(brks)] <- brks[length(brks)] + 1e-8
  } else {
    brks <- seq(min(pooled), max(pooled), length.out = r + 1)
  }

  binlab <- function(x) cut(x, breaks = brks, include.lowest = TRUE)
  tab <- sapply(blocks, function(b) table(binlab(b)))
  if (is.null(dim(tab))) tab <- matrix(tab, nrow = r)
  rownames(tab) <- levels(binlab(pooled))
  colnames(tab) <- paste0("block_", seq_len(q))

  tst <- suppressWarnings(chisq.test(tab, simulate.p.value = simulate_p, B = B))
  small <- any(tst$expected < min_expected)

  list(
    p_value = unname(tst$p.value),
    statistic = unname(tst$statistic),
    df = unname(tst$parameter),
    small_expected = small,
    method = if (simulate_p) "Chi-squared (MC p)" else "Pearson Chi-squared",
    r = r, q = q, K = K, k0 = k0, blocks_per_chain = blocks_per_chain,
    breaks = brks, table = tab
  )
}

out2 <- stationarity_chisq_v2(S_list, k0 = 0, blocks_per_chain = 2)
out4 <- stationarity_chisq_v2(S_list, k0 = 0, blocks_per_chain = 4)

out2$p_value
out4$p_value

```


# BIC Minimizers

## for all chains

```{r}

mcmc.samples.all = rbind(trimmed[[1]],
                         trimmed[[2]],
                        trimmed[[3]],
                         trimmed[[4]])

bic.minimizers.multiple.chains.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               mcmc.samples.all)
bic.minimizers.multiple.chains.tc.genesis

```
```{r}
saveRDS(bic.minimizers.multiple.chains.tc.genesis,
        file.path(TC_RESULTS_DIR, "bic.minimizers.pooled.chains.rds"))
```

## for chain 1

```{r}
bic.minimizers.chain1.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               trimmed[[1]])
bic.minimizers.chain1.tc.genesis

#saveRDS(bic.minimizers.chain1.tc.genesis,
#        file.path(TC_RESULTS_DIR,
#                  "bic.minimizers.chain1.tc.genesis.rds"))
```


## for chain 2
```{r}
bic.minimizers.chain2.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               trimmed[[2]])
bic.minimizers.chain2.tc.genesis

saveRDS(bic.minimizers.chain2.tc.genesis,
        file.path(TC_RESULTS_DIR,
                  "bic.minimizers.chain2.tc.genesis.rds"))
```


## for chain 3
```{r}
bic.minimizers.chain3.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               trimmed[[3]])
bic.minimizers.chain3.tc.genesis

saveRDS(bic.minimizers.chain2.tc.genesis,
        file.path(TC_RESULTS_DIR,
                  "bic.minimizers.chain3.tc.genesis.rds"))
```



## for chain 4 
```{r}
bic.minimizers.chain4.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               trimmed[[4]])
bic.minimizers.chain4.tc.genesis

saveRDS(bic.minimizers.chain4.tc.genesis,
        file.path(TC_RESULTS_DIR,
                  "bic.minimizers.chain4.tc.genesis.rds"))
```

# Plots
```{r, fig.width=8, fig.height=7}


inc_df <- inclusion_summary(mcmc.samples.tc.genesis,
                            top_k = 50
                            #ensure_gt = TRUE
                            )
plot_inclusion(inc_df,
               title="Inclusion probabilities with ESS 95% CI")


```


# Functions

## Standardize TC dataset, fit glm after
```{r}

std_and_examine_coefs = function(X, y, rs_best_vars, verbose=FALSE){
  
  # standardize design matrix
  X_std = scale(X)
  X_std = X_std[,rs_best_vars]
  
  # only take selected vars into account
  
  
  lm = glm(as.formula(paste(names(y), "~ .")),
           data=data.frame(X_std,y),
           family='binomial')
  
  
  if (verbose){
    print(summary(lm))  
  }
  
  return(lm)
  
}

```


### Analyze Result
```{r}

fit = std_and_examine_coefs(X, y, rs_bic_min_vars)


coefs <- summary(fit)$coefficients

coefs[names(sort(coefs[,"Estimate"], decreasing=TRUE)),]

```

```{r}
install.packages("statisticalModeling")
library(statisticalModeling)
effect_size(fit)
```



## Boxplot of frequently picked vars, grouped by y

```{r}

rs_bic_min_vars

```
```{r}

boxplot_var_by_TC_genesis = function(data, col_var, std = TRUE) {
  library(dplyr)
  library(ggplot2)
  
  # standardize data 
  df <- data
  if (std) {
    df[, colnames(df) != "TC_genesis"] <- scale(
      df[, colnames(df) != "TC_genesis"]
    )
  }
  
  # set TC_genesis to be a factor, NOT continuous
  df$TC_genesis <- as.factor(df$TC_genesis)
  
  # Compute group medians
  medians <- df %>%
    group_by(TC_genesis) %>%
    summarise(med = median(.data[[col_var]], na.rm = TRUE), .groups = "drop")
  
  
  
  # col_var is not a literal, so need to use .data[[..]]
  ggplot(df, aes(y = .data[[col_var]], x = TC_genesis)) +
    
    geom_boxplot() +
    
    # its hline not vline because we flip: (1) set horizontal first
    # (2) then flip
    geom_hline(data = medians, aes(yintercept = med, color = TC_genesis),
               linetype = "solid", size = 0.8, show.legend = FALSE,
               alpha=0.7) +
    
    coord_flip() +
    
    labs(y = col_var, x = "TC_genesis") +
    
    theme_minimal(base_size = 13)
}

```



```{r}
boxplot_var_by_TC_genesis <- function(data, col_var, std = TRUE,
                                      cols = c("0" = "#0B7D0B",  # dark green
                                               "1" = "#D62728"), # red
                                      show_trend = TRUE,
                                      median_width = 0.55,
                                      show_legend=TRUE) {
  library(dplyr); library(ggplot2)

  df <- data
  if (std) {
    df[, colnames(df) != "TC_genesis"] <- scale(df[, colnames(df) != "TC_genesis"])
  }
  
  
  # set TC_genesis to be a factor, NOT continuous
  df$TC_genesis <- factor(df$TC_genesis, levels = c(0, 1))

  # group medians
  meds <- df %>%
    group_by(TC_genesis) %>%
    summarise(med = median(.data[[col_var]], na.rm = TRUE), .groups = "drop")

  # for optional trend line between the two medians
  seg <- if (nrow(meds) == 2) data.frame(
    x = meds$TC_genesis, y = meds$med, grp = 1
  ) else NULL

  # col_var is not a literal, so need to use .data[[..]]
  ggplot(df, aes(x = TC_genesis, y = .data[[col_var]])) +
    
    # colored boxes by class
    geom_boxplot(aes(fill = TC_genesis), outlier.alpha = 0.25) +
    
    # short horizontal median bars (cleaner than full hlines)
    
    #geom_crossbar(data = meds,
    #              aes(x = TC_genesis, y = med, ymin = med-0.8, ymax = med+0.8,
    #                  color = TC_genesis),
    #              width = median_width, fatten = 1.1, inherit.aes = FALSE,
    #              show.legend = show_legend) +
    
    #geom_hline(data = meds, aes(yintercept = med, color = cols),
    #           linetype = "solid", size = 0.8, show.legend = FALSE,
    #           alpha=0.7) +
    
    # optional: show a “trend” line connecting medians
    { if (show_trend && !is.null(seg))
        geom_path(data = seg, aes(x = x, y = y, group = grp),
                  inherit.aes = FALSE, linetype = "solid", linewidth = 0.7,
                  color = "blue",
                  size=0.2,
                  lineend = "round"
                  ) } +
    
    coord_flip() +
    
    scale_fill_manual(values = cols, name = "TC_genesis",
                      labels = c("0 = no genesis", "1 = genesis")) +
    
    scale_color_manual(values = cols, guide = "none") +
    
    labs(x = "TC_genesis", y = col_var) +
    
    theme_minimal(base_size = 13) +
    
    theme(legend.position = ifelse(show_legend, "top", "none"))
}

```




```{r}
plist = list(
  p1 = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,"geopotential"),
  p2 = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,"geopotential_at_200_hPa", show_legend=FALSE),
  p3 = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,"geopotential_at_300_hPa", show_legend=FALSE),
  p4 = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,"geopotential_at_400_hPa", show_legend=FALSE),
  p5 = boxplot_var_by_TC_genesis(data.NA.fixed.no.basin,"geopotential_at_600_hPa", show_legend=FALSE)
)


plot_grid(plotlist = plist,
          ncol = 3, nrow=ceiling(length(plist)/3),
          rel_widths = c(2, 2, 2),
          rel_heights=c(2,2,2))



```


```{r}

cor(data.NA.fixed.no.basin[,grep("geopotential", rs_bic_min_vars, value=TRUE)])

```


# Test

```{r}
MH.random.search(rep(1,ncol(X)),
                 data.frame(X,y),
                 iter=1000)
```




