---
title: "random_search_RCP_TC_dataset"
author: "Keith Howen"
date: "2025-08-03"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)

```{r}
# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))

# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))

source(here("lib", "probit_gibbs.R"))
source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
#library(splitstackshape)
```

## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import TC genesis data & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
```



```{r}
dim(data.NA.fixed)
str(data.NA.fixed)
```
# Prep the data

```{r}



```

