---
title: "validate_RS_with_LASSO_TC_genesis"
author: "Keith Howen"
date: "2025-08-17"
output: html_document
---

# 💻 set seed for reproducibility
```{r}
set.seed(123)
```



# 💻 Working Directory Setup using `here()`
Use `here()` to *standardize the working directory between R markdown and
R console* as opposed to using `getwd()` (and `setwd()`), at the slight
inconveniences of importing the `here` library, using `here()` to write file
paths, and manually specify paths for files outside the working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports
## Check working directory
```{r}
here()
```

## Import all libraries (incl. custom)
```{r}

# one hot encoding with intercept
source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
#source(here("lib", "simulation_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))
#library(mvtnorm)
#library(truncnorm)
#library(Matrix)
library(ggplot2)
#library(splitstackshape)
#install.packages('reshape2')
library(reshape2)

library(glmnet)
install.packages('caret')
library(caret)

```


## Set path for data

```{r}
list.files(here("data"))
```

```{r}
data_path = here("data", "envDataset_12h_10x10_with_mask.csv")
```

# 💻 Import raw data & fix NA basin value

```{r}
# import the data
data = read.csv(data_path)

# fix basin NA category first (NA missing to "NA") and
# convert it into a factor
# (but not yet encoded yet)
data.NA.fixed = fix_NA_value_and_refactor_basin(data)
data.NA.fixed.no.basin = data.NA.fixed[colnames(data.NA.fixed) != "basin"]
data.NA.fixed.no.TC = data.NA.fixed.no.basin[colnames(data.NA.fixed.no.basin) != "TC_genesis"]

X.colnames = colnames(data.NA.fixed.no.TC)

X = data.NA.fixed.no.TC
y = data.NA.fixed.no.basin["TC_genesis"]

```




# Compute BIC Minmizers for RS

```{r}
list.files(here("output", "random_search_TC_dataset_outputs", "RCP_outputs")) 
```

```{r}

mcmc.samples.tc.genesis = readRDS(here("output", "random_search_TC_dataset_outputs", "RCP_outputs", "gibbs_TC_dataset_results_iter_1.rds")) 

bic.minimizers.tc.genesis = find_BIC_minimizer(data.frame(X,y),
                                               mcmc.samples.tc.genesis)

bic.minimizers.tc.genesis

```

```{r}

saveRDS(bic.minimizers.tc.genesis,
        here("output", "random_search_TC_dataset_outputs", "RCP_outputs", "top.bic.minimizers.tc.genesis.rds"))

```



```{r}

for(i in 1:5){
  cat(
    sprintf("Dimension for top %d smallest BIC minimizer: %d\n",
            i,
            length(bic.minimizers.tc.genesis$models_by_bic[[as.character(i)]]$models[[1]]))
  )
}





```


# Inspect relative frequencies
```{r}
rel.freqs.tc.genesis = readRDS(here("output",
                "random_search_TC_dataset_outputs",
                "RCP_outputs",
                "rel.freqs.rds")) 

sort(rel.freqs.tc.genesis$rel.freqs.non.basin, decreasing=TRUE)

```
## Plot rel freqs

```{r}
plot(sort(rel.freqs.tc.genesis$rel.freqs.non.basin, decreasing=TRUE))
abline(h=0.5, col='red')

```
```{r}

# 33 variables
sort(
  rel.freqs.tc.genesis$rel.freqs.non.basin[
  rel.freqs.tc.genesis$rel.freqs.non.basin > 0.75
]
)


```

## BIC for vars with v>0.85
```{r}
vars.with.top.rel.freqs = names(rel.freqs.tc.genesis$rel.freqs.non.basin[
  rel.freqs.tc.genesis$rel.freqs.non.basin > 0.5
])


submod.TC.genesis = function(cols, X, y){
  
  return(
    glm(TC_genesis ~ . ,
        data=data.frame(X[,vars.with.top.rel.freqs],y),
        family='binomial')
  )
}

mod.vars.with.top.rel.freqs = submod.TC.genesis(vars.with.top.rel.freqs,
                                                X,
                                                y)
summary(mod.vars.with.top.rel.freqs)

compute.BIC(data.frame(X,y),
            colnames(X) %in% vars.with.top.rel.freqs
            )
```



# Diagnostic plots

## Rel freqs with ESS-adjusted CI
```{r, fig.width=10, fig.height=10}
inc_df <- inclusion_summary(mcmc.samples.tc.genesis,
                            top_k = 50
                            #ensure_gt = TRUE
                            )
plot_inclusion(inc_df,
               title="Inclusion probabilities with ESS 95% CI")
```


## Trace plots of num vars
```{r}
plot_num_vars_trace(mcmc.samples.tc.genesis,
                    window = 50,
                    title = "Trace: number of variables in every sampled model",
                    dimension_BIC_minim = 50)
```


## Model path heatmap
```{r}
plot_model_heatmap(mcmc.samples.tc.genesis,
                   thin = 10,
                   order = "by_mean",
                   exclude_unimportant_vars=100,
                   #put_gts_on_top=FALSE,
                   title = "Model path heatmap")
```

## P-charts
```{r, fig.width=10, fig.height=30}

p_chart_slowest_converging_vars(mcmc.samples.tc.genesis,
                                "MCSE",
                                top_n=30)
```

## CONSIDER THINNING???!!!!! UNIVARIATE ACF PLOTS
```{r}
acf(mcmc.samples.tc.genesis[,2])
```

## COMPUTE ESS. SEE IF TOP K LOWEST IS AMONG BIC MINIMZIER OR NOT



# Import best variable set from RS
```{r}
BIC.minimizer.TC.genesis =
  readRDS(here("output", "random_search_TC_dataset_outputs", "RCP_outputs", "BIC.minimizer.rds")) 
rs_vars = BIC.minimizer.TC.genesis$models[[1]]$selected.variables
rs_vars
```

# Functions for CV Deviance

```{r}

## 1) Build stratified folds once
make_folds <- function(y, k = 5, seed = 123) {
  set.seed(seed)
  # caret returns a list of indices per fold; we’ll convert to a foldid vector
  idx_list <- createFolds(y, k = k, list = TRUE, returnTrain = FALSE)
  foldid <- integer(length(y))
  for (i in seq_along(idx_list)) foldid[idx_list[[i]]] <- i
  foldid
}

## 2) Binomial deviance on a test set given probs (per-observation)
binom_deviance_per_obs <- function(y_true, p_hat, eps = 1e-15) {
  p_hat <- pmin(pmax(p_hat, eps), 1 - eps)
  # deviance = -2 * sum(y log p + (1-y) log (1-p))
  -2 * mean(y_true * log(p_hat) + (1 - y_true) * log(1 - p_hat))
}

## 3) CV for a fixed predictor set (RS) using GLM
cv_deviance_rs <- function(X, y, vars, foldid) {
  k <- max(foldid)
  scores <- numeric(k)
  for (f in 1:k) {
    test_idx  <- which(foldid == f)
    train_idx <- setdiff(seq_along(y), test_idx)

    df_tr <- data.frame(X[train_idx, vars, drop=FALSE], y = y[train_idx])
    df_te <- data.frame(X[test_idx,  vars, drop=FALSE], y = y[test_idx])

    fit <- glm(y ~ ., data = df_tr, family = binomial)
    p_hat <- predict(fit, newdata = df_te, type = "response")
    scores[f] <- binom_deviance_per_obs(df_te$y, p_hat)
  }
  list(mean = mean(scores), sd = sd(scores), per_fold = scores)
}

## 4) CV for LASSO on the same folds (uses *all* predictors by design)
cv_deviance_lasso <- function(X, y, foldid) {
  # glmnet expects matrix X and vector y; foldid aligns with rows of X
  fit_cv <- cv.glmnet(
    x = as.matrix(X), y = y,
    family = "binomial",
    type.measure = "deviance",
    foldid = foldid,
    keep = TRUE # keep per-fold predictions if you want to inspect later
  )
  # glmnet's cvm is mean deviance per observation
  list(mean = as.numeric(min(fit_cv$cvm)),
       sd   = as.numeric(fit_cv$cvsd[which.min(fit_cv$cvm)]),
       model = fit_cv)
}


```


# Validate RS


## against LASSO


```{r}
## ------- Example usage -------
# X: data.frame of predictors (no response column)
# y: 0/1 numeric or factor with levels c(0,1)
# rs_vars: character vector of selected columns from X (your RS set)

X = data.NA.fixed.no.TC
y = data.NA.fixed$TC_genesis
k=10
# 1) folds
foldid <- make_folds(y, k = k, seed = 42)

# 2) RS model CV deviance
rs_out <- cv_deviance_rs(X, y, vars = rs_vars, foldid = foldid)

# 3) LASSO CV deviance (on all X)
lasso_out <- cv_deviance_lasso(X, y, foldid = foldid)

cat(sprintf("RS (fixed vars) CV deviance:   mean=%.4f (sd=%.4f)\n", rs_out$mean, rs_out$sd))
cat(sprintf("LASSO (glmnet) CV deviance:    mean=%.4f (sd=%.4f)\n", lasso_out$mean, lasso_out$sd))
```
Technically their performance, accounting for sd are indistinguishable
- Argue that RS is a fixed discrete method -- hence slightly higher variance.

```{r}
lasso_out
```



## against stepwise selection?
```{r}
#mod.stepwise = step(
#  glm(TC_genesis ~ ., 
#      family='binomial',
#      data=data.NA.fixed.no.basin),
#  steps=1,
#  trace=1
#)

#mod.stepwise
```



# Error bars
$$
\bar{x}\  \pm\ t_{0.975,4} \cdot \frac{s}{\sqrt{5}}
$$

```{r}
# Results
methods <- c("RS (fixed vars),\nWith 50 Variables", "LASSO (glmnet),\nWith 81 Variables")
means   <- c(rs_out$mean, lasso_out$mean)
sds     <- c(rs_out$sd, lasso_out$sd)
nfold   <- k
df      <- nfold - 1

# Critical t value for 95% CI
tval <- qt(0.975, df)

# Compute CI half-widths
errors <- tval * sds / sqrt(nfold)

# Put into a data.frame for plotting
df_plot <- data.frame(
  method = methods,
  mean   = means,
  lower  = means - errors,
  upper  = means + errors
)

df_plot$mean = round(df_plot$mean, 3)

#vars <- c("50 variables\npicked", "81 variables\npicked")  # text annotations
#df_plot$vars <- vars




###
palette <- c("LASSO (glmnet),\nWith 81 Variables" = "#2563EB",  # blue
             "RS (fixed vars),\nWith 50 Variables" = "#DC2626") # red

ggplot(df_plot, aes(x = method, y = mean)) +
  
  # bars (colored)
  geom_col(aes(fill = method), width = 0.5, alpha = 0.35, color='grey') +
  
  # error bars (always black)
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.12, linewidth = 1, color = "black") +
  
  # points (colored)
  geom_point(aes(color = method), size = 2.5) +
  
  # mean labels (colored)
  geom_text(aes(label = mean, color = method), 
            vjust = 0.7, hjust = -0.5, size = 4.5, fontface = "bold") +
  
  # manual colors
  scale_fill_manual(values = palette, guide = "none") +
  scale_color_manual(values = palette, guide = "none") +
  
  labs(
    title = "Comparison of RS vs LASSO",
    y = "CV Deviance (mean ± 95% CI)", x = NULL
  ) +
  
  coord_cartesian(ylim = c(0, max(df_plot$upper) + 0.05)) +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(margin = margin(t = 6))
  ) 


```

# Final Model 


```{r}

final_TC_genesis_model = 
    glm(TC_genesis ~ .,
    family='binomial', 
    data=data.NA.fixed.no.basin[,c(rs_vars,"TC_genesis")]
  )

summary(final_TC_genesis_model)


drop1(final_TC_genesis_model)

```



# Plot PR-AUC curves

```{r}
library(glmnet)
install.packages("precrec")
library(precrec)   # for PR curves/PR-AUC
set.seed(1)

cvfit_lasso <- cv.glmnet(as.matrix(X), y,
                         family="binomial",
                         type.measure="deviance",
                         nfolds=length(unique(foldid)))
```



```{r}


# Inputs assumed already have
# X: n x p matrix/data.frame, y: 0/1 (numeric or logical), foldid: vector 1..K
# vars_rs: column indices (or names) selected by RS
# cvfit_lasso: optional cv.glmnet object you already ran on full data (bias note below)

cv_pr_compare <- function(X, y, foldid, vars_rs,
                          refit_lasso_within_fold = TRUE,
                          family = "binomial",
                          rs_use_ridge = FALSE) {
  stopifnot(length(foldid) == nrow(X))
  K <- length(unique(foldid)); n <- nrow(X)

  sc_rs    <- numeric(n)
  sc_lasso <- numeric(n)

  for (k in seq_len(K)) {
    cat("Fold ", k, " evaluated\n")
    te <- which(foldid == k) # te means TEsting fold
    tr <- setdiff(seq_len(n), te) # tr means TRaining folds

    ## standardize within fold (means/sds from training)
    mu  <- colMeans(X[tr, , drop = FALSE])
    sds <- apply(X[tr, , drop = FALSE], 2, sd)
    sds[sds == 0] <- 1

    Xtr <- sweep(X[tr, , drop = FALSE], 2, mu, "-")
    Xtr <- sweep(Xtr, 2, sds, "/")
    Xte <- sweep(X[te, , drop = FALSE], 2, mu, "-")
    Xte <- sweep(Xte, 2, sds, "/")

    ## RS model on the fixed variable set
    Xtr_rs <- as.data.frame(Xtr[, vars_rs, drop = FALSE])
    Xte_rs <- as.data.frame(Xte[, vars_rs, drop = FALSE])

    if (!rs_use_ridge) {
      fit_rs <- glm(y[tr] ~ ., data = cbind(Xtr_rs, y = y[tr]), family = family)
      sc_rs[te] <- drop(predict(fit_rs, newdata = Xte_rs, type = "response"))
    } else {
      # ridge (alpha=0) to stabilize if separation/collinearity is severe
      # (but we shouldn't need this, as fitting with the selected 50 variables
      # was okay, no NAs or anything)
      fit_rs <- glmnet::cv.glmnet(as.matrix(Xtr_rs), y[tr], family = "binomial",
                                  alpha = 0, type.measure = "deviance")
      sc_rs[te] <- drop(predict(fit_rs, newx = as.matrix(Xte_rs),
                                type = "response", s = "lambda.min"))
    }

    ## --- LASSO path
    if (refit_lasso_within_fold) {
      cvk <- glmnet::cv.glmnet(as.matrix(Xtr), y[tr], family = "binomial",
                               type.measure = "deviance")
      sc_lasso[te] <- drop(predict(cvk, newx = as.matrix(Xte),
                                   type = "response", s = "lambda.min"))
    } else {
      stop("For fair comparison, set refit_lasso_within_fold = TRUE.")
    }
  }

  # PR/ROC with precrec
  # collect all predictied probs (out of sample)
  mm <- precrec::mmdata(scores = list(RS = sc_rs, LASSO = sc_lasso),
                        labels = list(RS = y,  LASSO = y),
                        modnames = c("RS", "LASSO"))
  ev <- precrec::evalmod(mm)
  list(ev = ev,
       scores = data.frame(sc_rs = sc_rs, sc_lasso = sc_lasso, y = y))
}



out <- cv_pr_compare(X, y, foldid, rs_vars, refit_lasso_within_fold = TRUE)

# Pretty plot
#autoplot(out$pr_object) + ggplot2::labs(title="PR Curves (pooled out-of-fold)")

```

```{r}
mm <- mmdata(scores = list(RS = out$scores$sc_rs, LASSO = out$scores$sc_lasso),
               labels = list(RS = y,  LASSO = y),
               modnames = c("RS", "LASSO"))
ev <- evalmod(mm)                 # PR + ROC; we’ll read PR parts
ap <- auc(ev)                     # includes PR-AUC (Average Precision)
```


```{r}
autoplot(ev) + ggplot2::labs(title="PR Curves (pooled out-of-fold)")
```



```{r}
hist(out$scores$sc_rs, breaks = 30)
hist(out$scores$sc_lasso, breaks = 30)
```

```{r}
K <- length(unique(foldid))
for (k in seq_len(K)) {
  te <- which(foldid == k)
  tr <- setdiff(seq_len(n), te)
  
  # --- RS model: refit glm on training with the fixed RS variables
  
  Xtr_rs <- as.matrix(X[tr, rs_vars, drop = FALSE])
  Xte_rs <- as.matrix(X[te, rs_vars, drop = FALSE])
  fit_rs <- glm(y[tr] ~ Xtr_rs, family = 'binomial')
  summary(fit_rs)
    
}
```



# Save glmnet fit and ROC, PR curves results
```{r}



list.files(
  here("output", "validation_outputs" )
)

saveRDS(out,
        here("output", "validation_outputs", "pr_roc_results.rds"))

saveRDS(cvfit_lasso,
        here("output", "validation_outputs", "cvfit_lasso.rds"))

saveRDS(cvfit_lasso,
        here("output", "validation_outputs", "cvfit_lasso.rds"))



saveRDS(rs_out,
        here("output", "validation_outputs", "rs_out_deviance.rds"))

saveRDS(lasso_out,
        here("output", "validation_outputs", "lasso_out_deviance.rds"))




```







