---
title: "generate_dataset_n_200_p_160"
author: "Keith Howen"
date: "2025-07-29"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)
# -------

```{r}
# TC genesis libs
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")

#library(splitstackshape)
```

# Defining parameters & Inits for this study.
# -------

NOTE in favor of the sparsity problem we also choose X1,..,X4 as
the ground truths.

```{r}
p=16
n=200

# Initializing a 'dict' like structure to store results for each
# of the 15 factor combinations
# this is because we cannot directly do multi-level indexing -- 
# we have to initialize each value as a list first, then overwrite it.
rhos <- c('-0.9', '-0.5', '0', '0.5', '0.9')
signal_strengths <- c('weak', 'moderate', 'strong')

# storing result from check_significnance() -- 
# that is, the synthetic response, fitted true model, distribution of y, etc
results <- list()
for (rho in rhos) {
  
  results[[rho]] <- list()
  
  for (s in signal_strengths) {
    results[[rho]][[s]] <- "empty"  
  }
  
}

# covariance matrix for design matrices
Sigmas <- list()
for (rho in rhos) {
  
  Sigmas[[rho]] <- list()
  
  for (s in signal_strengths) {
    Sigmas[[rho]][[s]] <- "empty"  
  }
  
}

# the X data for each combination
X_data <- list()
for (rho in rhos) {
  
  X_data[[rho]] <- list()
  
  for (s in signal_strengths) {
    X_data[[rho]][[s]] <- "empty"  
  }
  
}

results

```

Want:
- Histogram of true probabilities (with true beta coef)
- Histogram of probabilities in model containing X1,..,X4
- LRT in addition to wald test to test each of $H_0: \beta_i=0,\ 1 \le i \le4$
  - Do we need only p-values? 
  


# Foreword on synthetic design matrix generation.
# -------
One way to enforce the SAME design matrix for a value of rho regardless
of the s/n of the ground truths to y is to
- Set the X matrix for rho, moderate & weak s/n to be the same as X matrix
for rho, high s/n. 
Analogy: if b=a and c=a, then a=b=c. 

In other words,
"high" S/N is the canonical design matrix for each rho,
from which other S/N datasets are derived.


# ✓ Comb 1: rho= 0, high (or strong) s/n
# -------

```{r}

# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0'
desired_sig='strong'


stopifnot(desired_rho %in% rhos)
stopifnot(desired_sig %in% signal_strengths)


# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['high']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['high']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))

# we can try using initial beta values from Comb 1, 
# see if correlation reduces significance of ground truths or not.

results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.3, 1.3, -1.3, -1.3),
                                  X_data[[desired_rho]][[desired_sig]])

```
Potential question: why the true beta values are so different? 
- if intention is to mitiage Hauck-DOnner effect why didnt you do it sooner

# ✓ Comb 2: rho= 0, moderate s/n
# -------

Intuition: sometimes altering significance of X2, X4 isn't enough since 
even decreasing a bit can lead to weak signifiance
- Which is why one probably should consider lowering the significance of X1,X3
(since partial tests mean that, in the presence of the strong ones, how can
including this variable improve the likelihood)

```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0'
desired_sig='moderate'


stopifnot(desired_rho %in% rhos)
stopifnot(desired_sig %in% signal_strengths)


# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1, 0.2, -1, -0.33),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 3: rho= 0, weak s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0'
desired_sig='weak'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))

# we can try using initial beta values from Comb 1, 
# see if correlation reduces significance of ground truths or not.

results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.3, 0.1, -1.3, -0.25),
                                  X_data[[desired_rho]][[desired_sig]])

```

# ✓ Comb 4: rho= +0.5, strong s/n
# -------


```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.5'
desired_sig='strong'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.3, 1.3, -1.3, -1.3),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 5: rho= +0.5, moderate s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.5'
desired_sig='moderate'


stopifnot(desired_rho %in% rhos)
stopifnot(desired_sig %in% signal_strengths)


# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1, 0.4, -1, -0.5),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 6: rho= +0.5, weak s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.5'
desired_sig='weak'


stopifnot(desired_rho %in% rhos)
stopifnot(desired_sig %in% signal_strengths)


# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1, 0.3, -1, -0.5135),
                                  X_data[[desired_rho]][[desired_sig]])

```

# ✓ Comb 7: rho= -0.5, strong s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.5'
desired_sig='strong'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.3, 1.3, -1.3, -1.3),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 8: rho= -0.5, moderate s/n
# -------

```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.5'
desired_sig='moderate'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))

# initial beta from 
# rho=+0.5, moderate s/n which is c(1, 0.4, -1, -0.5)

results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1, 0.7, -1, -0.35),
                                  X_data[[desired_rho]][[desired_sig]])

```

# ✓ Comb 9: rho= -0.5, weak s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.5'
desired_sig='weak'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))

# initial beta from 
# rho=+0.5, weak s/n which is c(1, 0.3, -1, -0.5135)

results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1, 0.69, -1, -0.2),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 10: rho= +0.9, strong s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.9'
desired_sig='strong'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.5, -1.7, 1, -1),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 11: rho= +0.9, moderate s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.9'
desired_sig='moderate'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.6, -1.47, 1, -0.27),
                                  X_data[[desired_rho]][[desired_sig]])

```

# ✓ Comb 12: rho= +0.9, weak s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='0.9'
desired_sig='weak'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.6, -1.43, 1, -0.23),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 13: rho= -0.9, strong s/n
# -------

```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.9'
desired_sig='strong'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.37, -3, 1.1, -0.8),
                                  X_data[[desired_rho]][[desired_sig]])

```


# ✓ Comb 14: rho= -0.9, moderate s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.9'
desired_sig='moderate'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.37, -2.5, 1.1, -0.64),
                                  X_data[[desired_rho]][[desired_sig]])

```

# ✓ Comb 15: rho= -0.9, weak s/n
# -------
```{r}
# we just need to input these 2 variables. No hard coding rho and s/n levels :)
desired_rho='-0.9'
desired_sig='weak'

stopifnot( desired_rho %in% rhos)
stopifnot( desired_sig %in% signal_strengths)

# if both are still empty, this means we need to generate a new X matrix,
# in which case the covariance matrix and design matrix will be generated.
#
# Upon the generation, for desired_sig != 'high', the if condition will be false
# in which we can just copy the synthetic design matrix from rho, high s/n for
# the other s/n levels.

# note we use identical() because we cannot compare matrices and strings directly.
want_new_design_matrix =
  ( identical(Sigmas[[desired_rho]][['strong']],'empty') &
      identical(X_data[[desired_rho]][['strong']],'empty') )

if (want_new_design_matrix) {
  Sigmas[[desired_rho]][['strong']] = construct_block_cov(p, as.numeric(desired_rho))
  X_data[[desired_rho]][['strong']] = generate_synthetic_vars(Sigmas[[desired_rho]][['strong']], SIZE=n)
} else {
  # Deep copy to ensure complete independence
  Sigmas[[desired_rho]][[desired_sig]] <- as.matrix(Sigmas[[desired_rho]][['strong']])
  X_data[[desired_rho]][[desired_sig]] <- as.matrix(X_data[[desired_rho]][['strong']])
  # OR for absolutely bulletproof deep copy (works for any R object)
  # Sigmas[[desired_rho]][[desired_sig]] <- deepcopy(Sigmas[[desired_rho]][['strong']])
  # X_data[[desired_rho]][[desired_sig]] <- deepcopy(X_data[[desired_rho]][['strong']])
}


dim(X_data[[desired_rho]][[desired_sig]])
dim(Sigmas[[desired_rho]][[desired_sig]])
Sigmas[[desired_rho]][[desired_sig]][1:6,1:6]
sum(Sigmas[[desired_rho]][[desired_sig]]) - sum(diag(Sigmas[[desired_rho]][[desired_sig]]))



results[[desired_rho]][[desired_sig]] = check_significance(
                                  c(1.37, -1.8, 1.1, -0.4),
                                  X_data[[desired_rho]][[desired_sig]])

```

# Finally, save all results!
```{r}
BASE_DIR = "simulation_factorial_analysis_n_200_p_16_outputs"
outdir = here("output", BASE_DIR)
saveRDS(results,
        file = file.path(outdir, "results.rds"))
```





