---
title: "random_search_and_analyze_results_n_200_p_16"
author: "Keith Howen"
date: "2025-08-01"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)
# -------

```{r}
# TC genesis libs
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")

#library(splitstackshape)
```


# Read synthetic datasets

```{r}
BASE_DIR = "simulation_factorial_analysis_n_200_p_16_outputs"
outdir = here("output", BASE_DIR)
results = readRDS(file.path(outdir, "results.rds"))
```

# Verify correctness of synthetic datasets

## rho=0
```{r}
CORR_TEST = "0"
SN_TEST = "strong"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0"
SN_TEST = "moderate"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0"
SN_TEST = "weak"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

## rho=+0.5

```{r}
CORR_TEST = "0.5"
SN_TEST = "strong"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0.5"
SN_TEST = "moderate"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0.5"
SN_TEST = "weak"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```



## rho=-0.5
```{r}
CORR_TEST = "-0.5"
SN_TEST = "strong"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "-0.5"
SN_TEST = "moderate"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "-0.5"
SN_TEST = "weak"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```



## rho=+0.9
```{r}
CORR_TEST = "0.9"
SN_TEST = "strong"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0.9"
SN_TEST = "moderate"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "0.9"
SN_TEST = "weak"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```


## rho=-0.9
```{r}
CORR_TEST = "-0.9"
SN_TEST = "strong"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "-0.9"
SN_TEST = "moderate"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```

```{r}
CORR_TEST = "-0.9"
SN_TEST = "weak"
round(cor(results[[CORR_TEST]][[SN_TEST]]$X.syn)[1:5,1:5], 5)
results[[CORR_TEST]][[SN_TEST]]$lrt_vs_wald_table
```


# Running simulation studies for all 15 datasets.

```{r}

# initialize data structure to hold the gibbs samples

mcmc.samples <- list()
rhos <- c('-0.9', '-0.5', '0', '0.5', '0.9')
signal_strengths <- c('weak', 'moderate', 'strong')

for (rho in rhos) {
  mcmc.samples[[rho]] <- list()
  for (s in signal_strengths) {
    mcmc.samples[[rho]][[s]] <- "empty"  
  }
}

run_sim_studies = function(results, burnin=300, iter=2300){
  # checkpointing!
  if (file.exists(file.path(outdir, "mcmc_samples.rds"))) {
    # <<- we want to overwrite the GLOBAL variable
    mcmc.samples <<- readRDS(file.path(outdir, "mcmc_samples.rds"))
  }
  for (rho in rhos) {
    for (s in signal_strengths) {
      
      if (identical(mcmc.samples[[rho]][[s]], "empty")) {
        cat(sprintf("[%s] Begin running RS for %s, %s\n", Sys.time(), rho, s))
        X = results[[rho]][[s]]$X.syn
        y = results[[rho]][[s]]$y.syn
        vinit = rep(1, ncol(X))
        suppressWarnings(
          mcmc.samples[[rho]][[s]] <<-
            Gibbs.random.search(vinit, data.frame(X, y),
                              burnin=burnin, iter=iter, penalty='BIC')
        )
        saveRDS(mcmc.samples, file.path(outdir, "mcmc_samples.rds"))
      } else {
        cat(sprintf("[%s] RS for %s, %s already done\n", Sys.time(), rho, s))
      }
      
    }
  }
}


```


```{r}
run_sim_studies(results)
```

# Analyze BIC Minimizers
```{r}
rhos <- c('-0.9', '-0.5', '0', '0.5', '0.9')
signal_strengths <- c('weak', 'moderate', 'strong')

for (rho in rhos) {
    for (s in signal_strengths) {
      cat(sprintf("[%s] Calculating top 5 BIC mnimizers for %s, %s\n", Sys.time(), rho, s))
      X = results[[rho]][[s]]$X.syn
      y = results[[rho]][[s]]$y.syn 
      suppressWarnings(
        print(
          find_BIC_minimizer(data.frame(cbind(X, y)), mcmc.samples[[rho]][[s]])$models
        )
        
      )
      
    }
  }


```


# Format results into latex-renderable code.
```{r}
summary_table <- make_summary_from_mcmc_samples(
  mcmc.samples,
  ground_truth_vars = paste0("X", 1:4),
  results)
```
```{r}
summary_table
```

```{r}
format_sim_study_tables(summary_table)
```







