---
title: "reanalyze_RS_n_200_p_16"
author: "Keith Howen"
date: "2025-08-17"
output: html_document
---

# 💻 set seed for reproducibility

```{r}
set.seed(123)
```

# 💻 Working Directory Setup using `here()`

Use `here()` to *standardize the working directory between R markdown
and R console* as opposed to using `getwd()` (and `setwd()`), at the
slight inconveniences of importing the `here` library, using `here()` to
write file paths, and manually specify paths for files outside the
working directory.

```{r setup, include=FALSE}
#install.packages('here')
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

# Consolidate working directory & Imports

## Check working directory

```{r}
here()
```

## Import all libraries (incl. custom)
# -------

```{r}
# TC genesis libs
# one hot encoding with intercept
#source(here("lib", "preprocessing_v3.R"))
# full rank encoding with NO intercept
#source(here("lib", "preprocessing_v2.R"))
#source(here("lib", "probit_gibbs.R"))
#source(here("lib", "MCMC_diagnostic_functions.R"))
source(here("lib", "simulation_routines.R"))
source(here("lib", "factorial_analysis_routines.R"))
source(here("lib", "gibbs_sampler_random_search.R"))

library(mvtnorm)
library(truncnorm)
library(Matrix)
library(MASS)
library(mvnTest)
library(dplyr)
library(knitr)
library(kableExtra)

#install.packages("mvnTest")

#library(splitstackshape)
```


# Find BIC Minimizers

```{r}
BASE_DIR = "simulation_factorial_analysis_n_200_p_16_outputs"
outdir = here("output", BASE_DIR)
mcmc.samples = readRDS(file.path(outdir, "mcmc_samples.rds"))
results = readRDS(file.path(outdir, "results.rds"))

rhos <- c('-0.9', '-0.5', '0', '0.5', '0.9')
#signal_strengths <- c('weak', 'moderate', 'strong')
signal_strengths <- c('weak')

for (rho in rhos) {
    for (s in signal_strengths) {
      
      cat(sprintf("[%s] Calculating BIC mnimizer for %s, %s\n", Sys.time(), rho, s))
      X = results[[rho]][[s]]$X.syn
      y = results[[rho]][[s]]$y.syn 
      suppressWarnings(
        print(
          find_BIC_minimizer(data.frame(cbind(X, y)), mcmc.samples[[rho]][[s]])$models
        )
      )
      
      
      #cat(sprintf("[%s] Top Rel freqs for %s, %s\n", Sys.time(), rho, s))
      #print(
      #  sort(rel.freq.inference(mcmc.samples[[rho]][[s]])$rel.freqs.non.basin, decreasing=TRUE)
      #)
      
      
      
    }
  }

```

```{r}
compute.BIC.true.model.small = function(rho, s){
  
    X = results[[rho]][[s]]$X.syn
    y = results[[rho]][[s]]$y.syn 
    return(
      compute.BIC(data.frame(X,y), 
                  colnames(X) %in% c("X1","X2","X3","X4"))
    )
}


compute.BIC.true.model.small("-0.5", "weak")
```

# Plots

## ESS adjusted CIs
```{r}
plot_grid(
  
  plot_inclusion(inclusion_summary(mcmc.samples[["0.9"]][["moderate"]],
                            top_k = 16,
                            ensure_gt = TRUE,
                            gt_idx=1:4),
                 title = "Inclusion probabilities\nwith ESS 95% CI\n(Highly correlated GTs, moderate SNR)"),
  
  plot_inclusion(inclusion_summary(mcmc.samples[["0.9"]][["weak"]],
                            top_k = 16,
                            ensure_gt = TRUE,
                            gt_idx=1:4),
                 title = "Inclusion probabilities\nwith ESS 95% CI\n(Highly correlated GTs, weak SNR)"),
  ncol=2
  
  
)
```



## Model Path Heatmaps
```{r}
plot_grid(
  
  
  plot_model_heatmap(mcmc.samples[["0.9"]][["moderate"]],
                     thin = 5,
                     order = "by_mean",
                     exclude_unimportant_vars=12,
                     put_gts_on_top=FALSE,
                     transpose=FALSE,
                     title = "Model path heatmap (highly correlated GTs, moderate SNR)"),
    
  plot_model_heatmap(mcmc.samples[["0.9"]][["weak"]],
                     thin = 5,
                     order = "by_mean",
                     exclude_unimportant_vars=12,
                     put_gts_on_top=FALSE,
                     transpose=FALSE,
                     title = "Model path heatmap (highly correlated GTs, weak SNR)"), 
    
    
    nrow=2
  
)
```



## Trace of number of variables

```{r}
plot_grid(
  

  plot_num_vars_trace(mcmc.samples[["0.9"]][["moderate"]],
                    window = 10,
                    dimension_BIC_minim = 4,
                    title = "Trace: number of variables \n in sampled models \n (highly correlated GTs, \n moderate SNR)"), 
    
  plot_num_vars_trace(mcmc.samples[["0.9"]][["weak"]],
                    window = 10,
                    dimension_BIC_minim = 2,
                    title = "Trace: number of variables \n in sampled models \n (highly correlated GTs, \n weak SNR)"), 
    
    
  ncol=2
  
)
```




